---
title: "GBV Detection Analysis"
subtitle: "*Document structure and formatting were assisted with Cursor.*"
author: "Dr Tom Moore - Statistical Scientist"
date: "`r Sys.Date()`"
format:
  html:
    include-in-header:
      text: |
        <style>
        .quarto-title .subtitle,
        p.subtitle,
        .subtitle.lead {
          font-size: 0.88rem;
          font-weight: normal;
        }
        </style>
    df-print: paged
    code-fold: false
    toc: true
    toc-float: true
    toc-depth: 5
    theme: united
    highlight-style: tango
    fig-width: 9
    fig-height: 6
    fig-dpi: 600
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup-packages, message=FALSE, warning=FALSE}
# Install packages and functions (based on existing Rmd setups)
ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  invisible(sapply(pkg, require, character.only = TRUE))
}

packages <- c(
  "tidyverse", "readxl", "janitor", "here", "reshape2", "brms", "cmdstanr", "bayesplot", "scales", "knitr"
)

ipak(packages)

# Compact structure only: one header + `str()`-style columns (`str()` on tbl_df already prints the tibble line)
brief_dataframe <- function(x, label = NULL) {
  if (!is.null(label) && nzchar(as.character(label)[1])) {
    cat(label, "\n\n", sep = "")
  }
  x_tb <- dplyr::as_tibble(x)
  ch <- utils::capture.output(utils::str(x_tb, give.attr = FALSE))
  hdr <- sprintf("tibble [%d × %d] (S3: tbl_df/tbl/data.frame)", nrow(x_tb), ncol(x_tb))
  drop_first <- length(ch) > 0L && grepl("obs\\.\\s+of\\s+\\d+\\s+variables", ch[1L])
  body <- if (drop_first) ch[-1L] else ch
  out <- if (drop_first) {
    c(hdr, body)
  } else if (length(body) && startsWith(body[[1L]], "tibble [")) {
    body
  } else {
    c(hdr, body)
  }
  cat(paste(out, collapse = "\n"), "\n\n", sep = "")
  invisible(x_tb)
}

# Raw binary outcomes (0/1): nudge below 0 / above 1 so jitter clears model intervals on [0,1]
binary_y_for_plot <- function(y, pad = 0.06) {
  dplyr::case_when(
    is.na(y) ~ NA_real_,
    as.numeric(y) < 0.5 ~ as.numeric(y) - pad,
    TRUE ~ as.numeric(y) + pad
  )
}

# Probability scale — include 1 so percent labels show 100%
prob_det_y_breaks <- c(0, 0.25, 0.5, 0.75, 1)

bayesplot::color_scheme_set("blue")
options(mc.cores = parallel::detectCores())

# Ensure plot output directory exists
dir.create(here::here("outputs/plots"), recursive = TRUE, showWarnings = FALSE)

# Production Sever Settings
cmdstanr::set_cmdstan_path("/powerplant/workspace/hrltxm/workbench-k8s/stan/vanilla_stan/cmdstan-2.36.0")
```

# Methods:

To compare the probability of detection (%) of GVB at different exposure levels (0, 2, 6, 16, 24, 48 hrs), with and without a Cq cut off (i.e., if a Cq value was greater than 39 then classify as no detection), and GVB density (where data was available: at 2hr and 24hrs), generalized linear mixed effect (GLMM) models were performed in R (R Core Team 2022) using the brms package (Bürkner, 2017).

For all models a Bernoulli distribution was selected with a logit link function to model the binary detection response (yes = 1, no = 0). The fixed effects of time and number of GBV were treated as monotonic predictors since they are ordered categories. An interaction term was included between time and number of GVB when this information was available. A random effect of sample was used for all models since measures were collected in triplicates. A weakly informative prior was used for the fixed effects to aid convergence (normal distribution with a mean of 0 and standard deviation of 1.5).

Model validation was performed using prior and posterior predictive checks, as well as effective samples sizes and r-hats (McElreath, 2020).

## Experiment 6

```{r exp6-wrangle, include=FALSE}
# read in data ----
df.6 <- read_excel(here::here("analysis/data/Exp6-time2.xlsx"), sheet = 1)
df.6 <- df.6 %>%
  clean_names() %>%
  select(-exp)

cols <- c("sample", "rep", "sample_id")
df.6[, cols] <- lapply(df.6[, cols], as.factor)

df.6 <- df.6 %>%
  mutate(time = ordered(time)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive))

df.6.2h <- read_excel(here::here("analysis/data/Exp6-Number-2h.xlsx"), sheet = 1) %>%
  mutate(time = "2")
df.6.24h <- read_excel(here::here("analysis/data/Exp6-Number-24h.xlsx"), sheet = 1) %>%
  mutate(time = "24")

df.6.d <- rbind(df.6.2h, df.6.24h) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.6.d[, cols] <- lapply(df.6.d[, cols], as.factor)

df.6.d <- df.6.d %>%
  mutate(time = ordered(time)) %>%
  mutate(number = ordered(number)) %>%
  mutate(positive = ifelse(is.na(positive), 0, positive)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive))

# GVB reads workbooks (Exp 6 / Exp 10) — map to shared column names; same schema as df.6 time-course extras
wrangle_exp6_reads_sheet <- function(
  rel_path,
  require_quantity_gbv = TRUE,
  require_total_reads = FALSE
) {
  path <- here::here(rel_path)
  d <- read_excel(path, sheet = 1) %>% clean_names()
  d <- dplyr::select(d, -dplyr::any_of("exp"))

  if (!"time" %in% names(d) && "exposure_time_hours" %in% names(d)) {
    d <- dplyr::rename(d, time = exposure_time_hours)
  }
  if (!"positive" %in% names(d) && "q_pcr_pos" %in% names(d)) {
    d <- dplyr::rename(d, positive = q_pcr_pos)
  }
  if (!"positive" %in% names(d) && "qpcr_pos" %in% names(d)) {
    d <- dplyr::rename(d, positive = qpcr_pos)
  }

  if (!"quantity_gbv" %in% names(d)) {
    q_alias <- intersect(
      c("quantity_gvb", "quanity_gvb", "quanity_gbv", "quantity_gbv"),
      names(d)
    )
    if (length(q_alias)) {
      d <- dplyr::rename(d, quantity_gbv = dplyr::all_of(q_alias[[1]]))
    }
  }
  if (!"quantity_gbv" %in% names(d)) {
    nms <- names(d)
    iq <- nms[grepl("gbv", nms, ignore.case = TRUE)]
    iq <- iq[grepl("quant|quani", iq, ignore.case = TRUE)]
    if (length(iq) == 1L) {
      d <- dplyr::rename(d, quantity_gbv = dplyr::all_of(iq))
    }
  }
  if (require_quantity_gbv && !"quantity_gbv" %in% names(d)) {
    stop(
      "Reads sheet: could not map any column to `quantity_gbv` (expected names like `quantity_gvb` or `quanity_gbv` after clean_names). Found: ",
      paste(names(d), collapse = ", ")
    )
  }

  if (!"total_reads" %in% names(d)) {
    tr_alias <- intersect(
      c(
        "total_reads", "reads_total", "total_read", "n_reads", "n_total_reads",
        "sequencing_reads", "total_sequencing_reads", "library_size"
      ),
      names(d)
    )
    if (length(tr_alias)) {
      d <- dplyr::rename(d, total_reads = dplyr::all_of(tr_alias[[1]]))
    }
  }
  if (!"total_reads" %in% names(d)) {
    nms <- names(d)
    cand <- nms[grepl("read", nms, ignore.case = TRUE)]
    cand <- cand[grepl("total|^n_", cand, ignore.case = TRUE) | grepl("^reads_", cand, ignore.case = TRUE)]
    if (length(cand) == 1L) {
      d <- dplyr::rename(d, total_reads = dplyr::all_of(cand))
    }
  }
  if (require_total_reads && !"total_reads" %in% names(d)) {
    stop(
      "Reads sheet: could not map any column to `total_reads` (e.g. `total_reads`, `reads_total`, `n_reads` after clean_names). Found: ",
      paste(names(d), collapse = ", ")
    )
  }
  if ("total_reads" %in% names(d)) {
    d$total_reads <- suppressWarnings(as.numeric(d$total_reads))
  }

  miss <- setdiff(c("time", "cq", "positive", "sample"), names(d))
  if (length(miss)) {
    stop(
      "After clean_names() and renames, `", rel_path, "` must include: ",
      paste(miss, collapse = ", "),
      ". Found: ",
      paste(names(d), collapse = ", ")
    )
  }

  d$cq <- suppressWarnings(as.numeric(d$cq))
  pv <- d$positive
  if (is.numeric(pv)) {
    d$positive <- ifelse(is.na(pv), NA_real_, as.integer(round(pv) != 0L))
  } else {
    ps <- toupper(stringr::str_trim(as.character(pv)))
    d$positive <- dplyr::case_when(
      is.na(ps) | ps == "" ~ NA_real_,
      ps %in% c("1", "Y", "YES", "TRUE", "T", "POS", "POSITIVE") ~ 1,
      TRUE ~ 0
    )
  }

  fac_cols <- intersect(c("sample", "rep", "sample_id"), names(d))
  for (cl in fac_cols) {
    d[[cl]] <- as.factor(d[[cl]])
  }

  tnum <- suppressWarnings(as.numeric(d$time))
  if (!all(is.na(tnum))) {
    tlv <- sort(unique(tnum[!is.na(tnum)]))
    d$time <- ordered(tnum, levels = tlv)
  } else {
    d$time <- ordered(as.character(d$time))
  }

  d %>%
    dplyr::mutate(
      positive_cut = dplyr::if_else(!is.na(.data$cq) & .data$cq > 39, 0L, as.integer(.data$positive))
    )
}

df.6.reads <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp6_reads_tom.xlsx",
  require_total_reads = TRUE
)
df.6.reads.no0 <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp6_reads_tom_no_time0.xlsx",
  require_total_reads = TRUE
)
```

```{r exp6-reads-fig34-prep, include=FALSE}
# Shared y-range (count scale) and discrete palette for Figure 3 & 4 — avoids blue/green used in Figures 1–2 (Wong-style hues).
exp6_reads_fig_y_max_filtered <- function(dr) {
  if (!nrow(dr)) {
    return(0)
  }
  dr %>%
    dplyr::mutate(
      gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
      quantity_gbv = as.numeric(.data$quantity_gbv)
    ) %>%
    dplyr::filter(
      !is.na(.data$gvb_sequence_number),
      !is.na(.data$quantity_gbv),
      .data$quantity_gbv != 0,
      suppressWarnings(as.numeric(as.character(.data$time))) %in% c(2, 24)
    ) %>%
    dplyr::summarise(mx = max(c(0L, .data$gvb_sequence_number), na.rm = TRUE), .groups = "drop") %>%
    dplyr::pull(1)
}
exp6_reads_fig_y_upper <- max(
  exp6_reads_fig_y_max_filtered(df.6.reads),
  exp6_reads_fig_y_max_filtered(df.6.reads.no0),
  na.rm = TRUE
) * 1.45
exp6_reads_fig_y_upper <- max(exp6_reads_fig_y_upper, 10)
exp6_reads_y_breaks <- sort(unique(c(0, 1, 10, 100, 1000, 10000, 100000)))
exp6_reads_y_breaks_use <- exp6_reads_y_breaks[exp6_reads_y_breaks <= exp6_reads_fig_y_upper]
exp6_reads_y_labels <- function(x) ifelse(x == 0, "0", scales::label_number(big.mark = ",")(x))
# Numeric x for 2 h / 24 h (1 and 2) so raw jitter can sit slightly right of estimates; axis still labelled in hours.
exp6_reads_fig_x_nudge_raw <- 0.1
exp6_reads_fig_xlim <- c(0.45, 2.72)
# Figure 3 only: small negative floor (log1p domain x > -1) so means/CIs near zero are not clipped.
exp6_reads_fig_y_lower_f3 <- -0.18
exp6_reads_fig_legend_npc <- c(
  x = (2 - exp6_reads_fig_xlim[1]) / diff(exp6_reads_fig_xlim),
  y = (log1p(10) - log1p(exp6_reads_fig_y_lower_f3)) /
    (log1p(exp6_reads_fig_y_upper) - log1p(exp6_reads_fig_y_lower_f3))
)
exp6_reads_qty_palette <- function(n) {
  wong <- c(
    "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
    "#D55E00", "#CC79A7"
  )
  wong[seq_len(min(n, length(wong)))]
}
```

<details><summary>Data frames</summary>

```{r exp6-summary-tibbles, echo=FALSE}
cat("Original time-course (Exp6-time2.xlsx)\n\n")
brief_dataframe(dplyr::count(df.6, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.6)

cat("\n\nOriginal density design (2 h / 24 h number sheets)\n\n")
brief_dataframe(dplyr::count(df.6.d, sample, number, name = "n"), "Counts by sample × number")
brief_dataframe(df.6.d)

cat("\n\nReads — full (GVB_exp6_reads_tom.xlsx)\n\n")
brief_dataframe(dplyr::count(df.6.reads, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.6.reads)

cat("\n\nReads — no time 0 (GVB_exp6_reads_tom_no_time0.xlsx)\n\n")
brief_dataframe(dplyr::count(df.6.reads.no0, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.6.reads.no0)
```

</details>

::: {.panel-tabset}

```{r exp6-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

priors <- prior(normal(0,1.5), class = b)

mod.1 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.1.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod.2 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.2.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive ~ mo(time) * mo(number) + (1|sample_id))

mod.3 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.3.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) * mo(number) + (1|sample_id))

mod.4 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.4.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp6-waic-compute, include=FALSE}
waic_tbl_one_exp <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_exp(mod.1, "mod.1 (time monotonic, no cut)"),
  waic_tbl_one_exp(mod.2, "mod.2 (time monotonic, cut)"),
  waic_tbl_one_exp(mod.3, "mod.3 (time monot. × number, no cut)"),
  waic_tbl_one_exp(mod.4, "mod.4 (time monot. × number, cut)")
) %>%
  dplyr::arrange(WAIC, Model) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

## Effect plots

```{r exp6-eff1, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_1 <- conditional_effects(mod.1)[[1]] %>% mutate(cutoff = "No")
c_eff_2 <- conditional_effects(mod.2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_1, c_eff_2)

raw_exp6_eff1 <- dplyr::bind_rows(
  dplyr::mutate(df.6, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.6, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
) %>%
  dplyr::mutate(y_plot = binary_y_for_plot(.data$y))

y_lim_exp6_eff1 <- c(-0.13, 1.13)
tlv_eff1 <- levels(c_eff$time)
n_t_eff1 <- length(tlv_eff1)
ichr <- as.character(tlv_eff1)
i24_eff1 <- which(ichr == "24")[1L]
if (is.na(i24_eff1)) i24_eff1 <- which(suppressWarnings(as.numeric(ichr)) == 24)[1L]
if (is.na(i24_eff1)) i24_eff1 <- max(1L, round(n_t_eff1 / 2))
x_npc_eff1 <- (i24_eff1 - 0.5) / n_t_eff1
y_npc_eff1 <- (0.3 - y_lim_exp6_eff1[1]) / (y_lim_exp6_eff1[2] - y_lim_exp6_eff1[1])

p_exp6_eff1 <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff1,
    aes(x = time, y = y_plot, group = cutoff, color = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.03, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  theme(
    legend.position = c(x_npc_eff1, y_npc_eff1),
    legend.justification = c(0.5, 0.5),
    legend.margin = ggplot2::margin(4, 4, 4, 4),
    legend.background = element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = y_lim_exp6_eff1,
    breaks = prob_det_y_breaks,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title="Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_1.png"),
  plot = p_exp6_eff1,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff1
```

Figure 1: Experiment 6, **monotonic time** (`mo(time)`). Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Jittered points are raw binary outcomes by exposure time, coloured by cutoff; error bars are 95% credible intervals and coloured points are posterior means.

```{r exp6-eff2, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5, out.width='100%', dpi = 600}
c_eff_3 <- conditional_effects(mod.3, effect = "time:number")[[1]] %>% mutate(cutoff = "No")
c_eff_4 <- conditional_effects(mod.4, effect = "time:number")[[1]] %>% mutate(cutoff = "Yes") %>%  rename(positive = positive_cut)

c_eff <- rbind(c_eff_3, c_eff_4) %>%
  mutate(time = paste0("Time: ", time, " hrs")) %>%
  mutate(number = paste0(number, " GVB"))

c_eff$cutoff <- factor(c_eff$cutoff, levels = c("Yes", "No"))

raw_exp6_eff2 <- df.6.d %>%
  dplyr::mutate(
    time = paste0("Time: ", .data$time, " hrs"),
    number = paste0(.data$number, " GVB")
  ) %>%
  tidyr::pivot_longer(
    cols = c(positive, positive_cut),
    names_to = "which_y",
    values_to = "y"
  ) %>%
  dplyr::mutate(
    cutoff = factor(
      dplyr::if_else(.data$which_y == "positive", "No", "Yes"),
      levels = c("Yes", "No")
    ),
    y_plot = binary_y_for_plot(.data$y)
  )

p_exp6_eff2 <- c_eff %>% ggplot(aes(x = cutoff, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff2,
    aes(x = cutoff, y = y_plot, color = cutoff),
    inherit.aes = FALSE,
    position = position_jitter(width = 0.15, height = 0.03),
    alpha = 0.35,
    size = 0.85,
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3, alpha = 0.5) +
  theme_classic() +
  theme(
    strip.text = element_text(size = 9, margin = ggplot2::margin(4, 4, 4, 4)),
    strip.background.x = element_rect(fill = "white", colour = "grey35", linewidth = 0.5),
    strip.background.y = element_rect(fill = "white", colour = "grey35", linewidth = 0.5),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.3),
    panel.grid.major.y = element_line(colour = "grey85", linewidth = 0.3)
  ) +
  scale_color_manual(values = c("darkgreen", "darkblue")) +
  labs(x = "Cq cut off (> 39)", y = "Probability of detection (%)") +
  guides(color = "none") +
  facet_grid(number ~ time) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(-0.13, 1.13),
    breaks = prob_det_y_breaks,
    expand = ggplot2::expansion(mult = c(0, 0))
  )

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_2.png"),
  plot = p_exp6_eff2,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff2
```

Figure 2: Experiment 6, **monotonic time**. Conditional effect plot of GBV exposure on probability of detection faceted by time and number of GVB with (yes = green) and without (no = blue) a Cq cut off of 39. Jittered points are raw outcomes, coloured by cutoff; error bars are 95% credible intervals and coloured points are posterior means.

## Tables

```{r exp6-estimates, echo=FALSE}
# Modeled estimates from conditional_effects() on the probability scale.
# Convert to percent for reporting.

ce6_time <- dplyr::bind_rows(
  conditional_effects(mod.1)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.2)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

ce6_time_number <- dplyr::bind_rows(
  conditional_effects(mod.3, effect = "time:number")[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.4, effect = "time:number")[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`, `Number (GVB)`)

knitr::kable(
  ce6_time,
  caption = "Modeled estimates from conditional effects (Experiment 6): time",
  format = "html"
)

knitr::kable(
  ce6_time_number,
  caption = "Modeled estimates from conditional effects (Experiment 6): time x number",
  format = "html"
)
```

## Model comparison

```{r exp6-waic-tab, echo=FALSE}
knitr::kable(
  exp6_waic_tbl,
  caption = "Experiment 6: WAIC (lower is better; `mo(time)` formulations).",
  format = "html"
)
```

## Models

```{r exp6-models-show, ref.label="exp6-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp6-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod.1, ndraws = 200, col = 2)
pp_check(mod.2, ndraws = 200, col = 2)
pp_check(mod.3, ndraws = 200, col = 2)
pp_check(mod.4, ndraws = 200, col = 2)
```

## Output

```{r exp6-summary, echo=FALSE}
summary(mod.1)
summary(mod.2)
summary(mod.3)
summary(mod.4)
```

<!-- Pairwise comparisons removed -->

```{r exp6-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.1, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  },
  {
    emm <- emmeans::emmeans(mod.2, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)

pairwise_number <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.3, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  },
  {
    emm <- emmeans::emmeans(mod.4, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  }
)

knitr::kable(
  pairwise_number,
  caption = "Pairwise comparisons between number levels within each time (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

### GVB sequence counts - reads

::: {.panel-tabset}

```{r exp6-reads-models-fit, include=FALSE}
df_exp6_reads_mod <- df.6.reads %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv),
    total_reads = as.numeric(.data$total_reads)
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    !is.na(.data$total_reads),
    .data$total_reads > 0,
    .data$gvb_sequence_number >= 0L,
    .data$quantity_gbv != 0,
    suppressWarnings(as.numeric(as.character(.data$time))) %in% c(2, 24)
  ) %>%
  dplyr::mutate(
    time = ordered(
      suppressWarnings(as.numeric(as.character(.data$time))),
      levels = c(2, 24)
    ),
    quantity_gbv = ordered(
      .data$quantity_gbv,
      levels = sort(unique(stats::na.omit(as.numeric(.data$quantity_gbv))))
    )
  )

qty_fac_lv_reads <- levels(df_exp6_reads_mod$quantity_gbv)
df_exp6_reads_fac <- df_exp6_reads_mod %>%
  dplyr::mutate(
    time = factor(as.character(.data$time), levels = c("2", "24")),
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = qty_fac_lv_reads)
  )

mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

mod_exp6_reads <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_mod,
  file = paste0(here::here("outputs/models/mod_exp6_reads_gvbseq_negbin_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

mod_exp6_reads_mo_int <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_mod,
  file = paste0(here::here("outputs/models/mod_exp6_reads_mo_interact_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

mod_exp6_reads_fac_add <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_fac,
  file = paste0(here::here("outputs/models/mod_exp6_reads_fac_add_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

mod_exp6_reads_fac_int <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_fac,
  file = paste0(here::here("outputs/models/mod_exp6_reads_fac_int_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp6-reads-waic-compute, include=FALSE}
waic_tbl_one_reads <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_reads_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_reads(mod_exp6_reads, "mo(time) + mo(quantity_gbv)"),
  waic_tbl_one_reads(mod_exp6_reads_mo_int, "mo(time) * mo(quantity_gbv)"),
  waic_tbl_one_reads(mod_exp6_reads_fac_add, "time + quantity_gbv (factor)"),
  waic_tbl_one_reads(mod_exp6_reads_fac_int, "time * quantity_gbv (factor)")
) %>%
  dplyr::arrange(WAIC) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

#### Effect plots

```{r exp6-reads-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4.5, out.width="100%", dpi=600}
raw <- df_exp6_reads_mod

ce_both <- conditional_effects(mod_exp6_reads_mo_int, effects = "time:quantity_gbv")[[1]]
qty_lv <- sort(unique(as.numeric(as.character(ce_both$quantity_gbv))))
ce_both <- ce_both %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv))
  ) %>%
  dplyr::mutate(
    tord = suppressWarnings(as.numeric(as.character(.data$time))),
    xtime = as.numeric(.data$time)
  ) %>%
  dplyr::arrange(.data$quantity_gbv, .data$tord)

raw_col <- raw %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both$quantity_gbv)),
    xtime = as.numeric(.data$time) + exp6_reads_fig_x_nudge_raw
  )

n_qty <- dplyr::n_distinct(ce_both$quantity_gbv)
pal <- exp6_reads_qty_palette(n_qty)
ce_pd <- ggplot2::position_dodge(width = 0.85)

exp6_reads_plot_theme <- ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
    panel.grid.minor.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
    legend.position = unname(exp6_reads_fig_legend_npc),
    legend.justification = c(0.5, 0.5),
    legend.title = ggplot2::element_text(size = 10),
    legend.text = ggplot2::element_text(size = 9),
    legend.margin = ggplot2::margin(2, 4, 2, 4),
    legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  )

p_exp6_reads_eff <- ggplot2::ggplot(
  ce_both,
  ggplot2::aes(
    x = .data$xtime,
    y = .data$estimate__,
    ymin = .data$lower__,
    ymax = .data$upper__,
    color = .data$quantity_gbv
  )
) +
  ggplot2::geom_point(
    data = raw_col,
    ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
    inherit.aes = FALSE,
    position = ggplot2::position_jitterdodge(jitter.width = 0.12, jitter.height = 0.175, dodge.width = 0.85),
    alpha = 0.4,
    size = 0.85,
    stroke = 0
  ) +
  ggplot2::geom_errorbar(position = ce_pd, width = 0.12, linewidth = 0.9) +
  ggplot2::geom_point(position = ce_pd, size = 2.5, alpha = 0.9) +
  ggplot2::scale_color_manual(values = pal) +
  ggplot2::scale_x_continuous(breaks = 1:2, labels = c("2", "24"), limits = exp6_reads_fig_xlim, expand = c(0, 0)) +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    limits = c(exp6_reads_fig_y_lower_f3, exp6_reads_fig_y_upper),
    breaks = exp6_reads_y_breaks_use,
    labels = exp6_reads_y_labels,
    minor_breaks = NULL,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  exp6_reads_plot_theme +
  ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
  ggplot2::guides(color = ggplot2::guide_legend(title = "GVB quantity", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_reads_effect_time_by_quantity.png"),
  plot = p_exp6_reads_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_reads_eff
```

Figure 3: Experiment 6, **GVB sequence counts** (reads workbook). Conditional effect plot of GVB exposure (2 h and 24 h) on mean sequence count by GVB quantity under **`mo(time) × mo(quantity)`** with **`offset(log(total_reads))`**. Vertical axis is **log(1 + count)** (lower limit slightly below zero so small estimates are visible). Jittered semi-transparent points are raw counts; error bars are 95% credible intervals and coloured points are posterior means.

#### Tables

Tabulated estimates use the **preferred monotonic interaction** model **`mo(time) * mo(quantity_gbv)`** (negative binomial mean count with **`offset(log(total_reads))`**; marginal **`conditional_effects()`**). See the **Model comparison** tab for **WAIC** across all candidate models.

```{r exp6-reads-estimates, echo=FALSE}
ce_exp6_t <- conditional_effects(mod_exp6_reads_mo_int, effects = "time")[[1]] %>%
  dplyr::transmute(
    `Time (hrs)` = time,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

ce_exp6_q <- conditional_effects(mod_exp6_reads_mo_int, effects = "quantity_gbv")[[1]] %>%
  dplyr::transmute(
    `Quantity GVB` = quantity_gbv,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

knitr::kable(
  ce_exp6_t,
  caption = "Experiment 6 reads — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
  format = "html"
)

knitr::kable(
  ce_exp6_q,
  caption = "Experiment 6 reads — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
  format = "html"
)
```

#### Models

```{r exp6-reads-models-show, ref.label="exp6-reads-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp6-reads-waic-tab, echo=FALSE}
knitr::kable(
  exp6_reads_waic_tbl,
  caption = "Experiment 6 reads: WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
```

#### Validation

```{r exp6-reads-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_exp6_reads, ndraws = 200)
pp_check(mod_exp6_reads_mo_int, ndraws = 200)
pp_check(mod_exp6_reads_fac_add, ndraws = 200)
pp_check(mod_exp6_reads_fac_int, ndraws = 200)
```

#### Output

```{r exp6-reads-summary, echo=FALSE}
summary(mod_exp6_reads)
summary(mod_exp6_reads_mo_int)
summary(mod_exp6_reads_fac_add)
summary(mod_exp6_reads_fac_int)
```

:::

### GVB sequence counts - no time 0

::: {.panel-tabset}

```{r exp6-reads-no0-models-fit, include=FALSE}
df_exp6_reads_no0_mod <- df.6.reads.no0 %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv),
    total_reads = as.numeric(.data$total_reads)
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    !is.na(.data$total_reads),
    .data$total_reads > 0,
    .data$gvb_sequence_number >= 0L,
    .data$quantity_gbv != 0,
    suppressWarnings(as.numeric(as.character(.data$time))) %in% c(2, 24)
  ) %>%
  dplyr::mutate(
    time = ordered(
      suppressWarnings(as.numeric(as.character(.data$time))),
      levels = c(2, 24)
    ),
    quantity_gbv = ordered(
      .data$quantity_gbv,
      levels = sort(unique(stats::na.omit(as.numeric(.data$quantity_gbv))))
    )
  )

qty_fac_lv_no0 <- levels(df_exp6_reads_no0_mod$quantity_gbv)
df_exp6_reads_no0_fac <- df_exp6_reads_no0_mod %>%
  dplyr::mutate(
    time = factor(as.character(.data$time), levels = c("2", "24")),
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = qty_fac_lv_no0)
  )

mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

mod_exp6_reads_no0 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_mod,
  file = paste0(here::here("outputs/models/mod_exp6_reads_no0_gvbseq_negbin_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

mod_exp6_reads_no0_mo_int <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_mod,
  file = paste0(here::here("outputs/models/mod_exp6_reads_no0_mo_interact_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

mod_exp6_reads_no0_fac_add <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_fac,
  file = paste0(here::here("outputs/models/mod_exp6_reads_no0_fac_add_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

mod_exp6_reads_no0_fac_int <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_fac,
  file = paste0(here::here("outputs/models/mod_exp6_reads_no0_fac_int_offset")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp6-reads-no0-waic-compute, include=FALSE}
waic_tbl_one_reads_no0 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_reads_no0_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_reads_no0(mod_exp6_reads_no0, "mo(time) + mo(quantity_gbv)"),
  waic_tbl_one_reads_no0(mod_exp6_reads_no0_mo_int, "mo(time) * mo(quantity_gbv)"),
  waic_tbl_one_reads_no0(mod_exp6_reads_no0_fac_add, "time + quantity_gbv (factor)"),
  waic_tbl_one_reads_no0(mod_exp6_reads_no0_fac_int, "time * quantity_gbv (factor)")
) %>%
  dplyr::arrange(WAIC) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

#### Effect plots

```{r exp6-reads-no0-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4.5, out.width="100%", dpi=600}
raw_no0 <- df_exp6_reads_no0_mod

ce_both_no0 <- conditional_effects(mod_exp6_reads_no0_mo_int, effects = "time:quantity_gbv")[[1]]
qty_lv_no0 <- sort(unique(as.numeric(as.character(ce_both_no0$quantity_gbv))))
ce_both_no0 <- ce_both_no0 %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv_no0))
  ) %>%
  dplyr::mutate(
    tord = suppressWarnings(as.numeric(as.character(.data$time))),
    xtime = as.numeric(.data$time)
  ) %>%
  dplyr::arrange(.data$quantity_gbv, .data$tord)

raw_col_no0 <- raw_no0 %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both_no0$quantity_gbv)),
    xtime = as.numeric(.data$time) + exp6_reads_fig_x_nudge_raw
  )

exp6_reads_fig_y_max_no0 <- max(
  0,
  suppressWarnings(max(as.numeric(ce_both_no0$upper__), na.rm = TRUE)),
  suppressWarnings(max(as.numeric(ce_both_no0$estimate__), na.rm = TRUE)),
  suppressWarnings(max(raw_col_no0$gvb_sequence_number, na.rm = TRUE)),
  na.rm = TRUE
)
# Generous headroom so means, 95% interval tops, and geom glyphs are not clipped (log1p axis, dodge=0.85).
exp6_reads_fig_y_upper_no0 <- max(
  exp6_reads_fig_y_upper,
  exp6_reads_fig_y_max_no0 * 1.72 + max(80, 0.12 * exp6_reads_fig_y_max_no0) + max(40, 0.04 * exp6_reads_fig_y_max_no0),
  10
)
exp6_reads_y_breaks_use_no0 <- exp6_reads_y_breaks[exp6_reads_y_breaks <= exp6_reads_fig_y_upper_no0]
exp6_reads_fig_legend_npc_no0 <- c(
  x = (2 - exp6_reads_fig_xlim[1]) / diff(exp6_reads_fig_xlim),
  y = log1p(10) / log1p(exp6_reads_fig_y_upper_no0)
)

n_qty_no0 <- dplyr::n_distinct(ce_both_no0$quantity_gbv)
pal_no0 <- exp6_reads_qty_palette(n_qty_no0)
ce_pd_no0 <- ggplot2::position_dodge(width = 0.85)

exp6_reads_plot_theme_no0 <- ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
    panel.grid.minor.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
    legend.position = unname(exp6_reads_fig_legend_npc_no0),
    legend.justification = c(0.5, 0.5),
    legend.title = ggplot2::element_text(size = 10),
    legend.text = ggplot2::element_text(size = 9),
    legend.margin = ggplot2::margin(2, 4, 2, 4),
    legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  )

p_exp6_reads_no0_eff <- ggplot2::ggplot(
  ce_both_no0,
  ggplot2::aes(
    x = .data$xtime,
    y = .data$estimate__,
    ymin = .data$lower__,
    ymax = .data$upper__,
    color = .data$quantity_gbv
  )
) +
  ggplot2::geom_point(
    data = raw_col_no0,
    ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
    inherit.aes = FALSE,
    position = ggplot2::position_jitterdodge(jitter.width = 0.12, jitter.height = 0.175, dodge.width = 0.85),
    alpha = 0.4,
    size = 0.85,
    stroke = 0
  ) +
  ggplot2::geom_errorbar(position = ce_pd_no0, width = 0.12, linewidth = 0.9) +
  ggplot2::geom_point(position = ce_pd_no0, size = 2.5, alpha = 0.9) +
  ggplot2::scale_color_manual(values = pal_no0) +
  ggplot2::scale_x_continuous(breaks = 1:2, labels = c("2", "24"), limits = exp6_reads_fig_xlim, expand = c(0, 0)) +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    limits = c(0, exp6_reads_fig_y_upper_no0),
    breaks = exp6_reads_y_breaks_use_no0,
    labels = exp6_reads_y_labels,
    minor_breaks = NULL,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  exp6_reads_plot_theme_no0 +
  ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
  ggplot2::guides(color = ggplot2::guide_legend(title = "GVB quantity", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_reads_no0_effect_time_by_quantity.png"),
  plot = p_exp6_reads_no0_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_reads_no0_eff
```

Figure 4: Experiment 6, **GVB sequence counts** (reads workbook without time 0). Same design as Figure 3: **`mo(time) × mo(quantity)`** with **`offset(log(total_reads))`**, **log(1 + count)** on the vertical axis. Jittered points are raw counts; error bars are 95% credible intervals and coloured points are posterior means. Source workbook omits 0 h exposure rows.

#### Tables

Tabulated estimates use the **preferred monotonic interaction** model **`mo(time) * mo(quantity_gbv)`** (negative binomial mean count with **`offset(log(total_reads))`**; marginal **`conditional_effects()`**). See the **Model comparison** tab for **WAIC** across all candidate models.

```{r exp6-reads-no0-estimates, echo=FALSE}
ce_exp6_no0_t <- conditional_effects(mod_exp6_reads_no0_mo_int, effects = "time")[[1]] %>%
  dplyr::transmute(
    `Time (hrs)` = time,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

ce_exp6_no0_q <- conditional_effects(mod_exp6_reads_no0_mo_int, effects = "quantity_gbv")[[1]] %>%
  dplyr::transmute(
    `Quantity GVB` = quantity_gbv,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

knitr::kable(
  ce_exp6_no0_t,
  caption = "Experiment 6 reads (no time 0) — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
  format = "html"
)

knitr::kable(
  ce_exp6_no0_q,
  caption = "Experiment 6 reads (no time 0) — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
  format = "html"
)
```

#### Models

```{r exp6-reads-no0-models-show, ref.label="exp6-reads-no0-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp6-reads-no0-waic-tab, echo=FALSE}
knitr::kable(
  exp6_reads_no0_waic_tbl,
  caption = "Experiment 6 reads (no time 0): WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
```

#### Validation

```{r exp6-reads-no0-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_exp6_reads_no0, ndraws = 200)
pp_check(mod_exp6_reads_no0_mo_int, ndraws = 200)
pp_check(mod_exp6_reads_no0_fac_add, ndraws = 200)
pp_check(mod_exp6_reads_no0_fac_int, ndraws = 200)
```

#### Output

```{r exp6-reads-no0-summary, echo=FALSE}
summary(mod_exp6_reads_no0)
summary(mod_exp6_reads_no0_mo_int)
summary(mod_exp6_reads_no0_fac_add)
summary(mod_exp6_reads_no0_fac_int)
```

:::

## Experiment 10

```{r exp10-wrangle, include=FALSE}
# Shared hour scale for detection (`df.10`) and reads (`df.10.reads`): Excel may use
# `16 hours`, `1 day`, `2 days`, … or short forms `16hrs`, `1d`, …; whitespace removed + lowercased first.
exp10_time_to_hours <- function(x) {
  xc <- trimws(as.character(x))
  xc <- gsub("\\s+", "", xc)
  xc <- tolower(xc)
  dplyr::case_when(
    xc == "" | is.na(xc) ~ NA_real_,
    xc == "0" ~ 0,
    xc %in% c("16hrs", "16hours", "16hour") ~ 16,
    xc %in% c("1d", "1day") ~ 24,
    xc %in% c("2d", "2day", "2days") ~ 48,
    xc %in% c("3d", "3day", "3days") ~ 72,
    xc %in% c("7d", "7day", "7days") ~ 168,
    grepl("^\\d+(\\.\\d+)?hours?$", xc) ~ suppressWarnings(
      as.numeric(sub("hours?$", "", xc, perl = TRUE))
    ),
    grepl("^\\d+(\\.\\d+)?days?$", xc) ~ suppressWarnings(
      as.numeric(sub("days?$", "", xc, perl = TRUE)) * 24
    ),
    grepl("^\\d+(\\.\\d+)?$", xc) ~ suppressWarnings(as.numeric(xc)),
    TRUE ~ NA_real_
  )
}

df.10 <- read_excel(here::here("analysis/data/Exp10- no cut off.xlsx"), sheet = 1) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.10[, cols] <- lapply(df.10[, cols], as.factor)

df.10 <- df.10 %>%
  dplyr::mutate(positive_cut = ifelse(cq > 39, 0, positive)) %>%
  dplyr::mutate(time_hours = exp10_time_to_hours(.data$time)) %>%
  dplyr::mutate(
    time = ordered(.data$time_hours, levels = sort(unique(stats::na.omit(.data$time_hours))))
  ) %>%
  dplyr::select(-dplyr::all_of("time_hours"))

df.10.reads <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp10_reads_tom.xlsx",
  require_quantity_gbv = FALSE,
  require_total_reads = TRUE
) %>%
  dplyr::mutate(time_hours = exp10_time_to_hours(.data$time)) %>%
  dplyr::mutate(
    time = ordered(.data$time_hours, levels = sort(unique(stats::na.omit(.data$time_hours))))
  )

df.10.reads.no0 <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp10_reads_tom_no_time_0.xlsx",
  require_quantity_gbv = FALSE,
  require_total_reads = TRUE
) %>%
  dplyr::mutate(time_hours = exp10_time_to_hours(.data$time)) %>%
  dplyr::mutate(
    time = ordered(.data$time_hours, levels = sort(unique(stats::na.omit(.data$time_hours))))
  )

# TRUE when sequence-count workbooks include a mappable quantity column (Experiment-6-style models).
exp10_reads_has_quantity <- "quantity_gbv" %in% names(df.10.reads)

# Shared row filter for Exp 10 negative-binomial reads (matches model fitting chunks).
# If `quantity_gbv` is absent, rows are not filtered on quantity (time-only models).
exp10_reads_filtered_for_nb <- function(dr) {
  has_q <- "quantity_gbv" %in% names(dr)
  has_th <- "time_hours" %in% names(dr)
  d <- dr %>%
    dplyr::mutate(
      gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
      total_reads = as.numeric(.data$total_reads),
      tnum = if (has_th) {
        suppressWarnings(as.numeric(.data$time_hours))
      } else {
        suppressWarnings(as.numeric(as.character(.data$time)))
      }
    )
  if (has_q) {
    d <- d %>% dplyr::mutate(quantity_gbv = as.numeric(.data$quantity_gbv))
  }
  d <- d %>%
    dplyr::filter(
      !is.na(.data$gvb_sequence_number),
      !is.na(.data$time),
      !is.na(.data$total_reads),
      .data$total_reads > 0,
      .data$gvb_sequence_number >= 0L
    )
  if (has_q) {
    d <- d %>% dplyr::filter(!is.na(.data$quantity_gbv), .data$quantity_gbv != 0)
  }
  d
}

exp10_reads_fig_y_max_filtered <- function(dr) {
  if (!nrow(dr)) {
    return(0)
  }
  exp10_reads_filtered_for_nb(dr) %>%
    dplyr::summarise(mx = max(c(0L, .data$gvb_sequence_number), na.rm = TRUE), .groups = "drop") %>%
    dplyr::pull(1)
}
```

```{r exp10-reads-fig34-prep, include=FALSE}
exp10_reads_fig_y_upper <- max(
  exp10_reads_fig_y_max_filtered(df.10.reads),
  exp10_reads_fig_y_max_filtered(df.10.reads.no0),
  na.rm = TRUE
) * 1.45
exp10_reads_fig_y_upper <- max(exp10_reads_fig_y_upper, 10)
exp10_reads_y_breaks <- sort(unique(c(0, 1, 10, 100, 1000, 10000, 100000)))
exp10_reads_y_breaks_use <- exp10_reads_y_breaks[exp10_reads_y_breaks <= exp10_reads_fig_y_upper]
exp10_reads_y_labels <- function(x) ifelse(x == 0, "0", scales::label_number(big.mark = ",")(x))
exp10_reads_fig_x_nudge_raw <- 0.1
exp10_reads_fig_y_lower_f3 <- -0.18
exp10_reads_xlim_for_n <- function(n) c(0.45, n + 0.27)
exp10_reads_qty_palette <- exp6_reads_qty_palette
```

<details><summary>Data frames</summary>

```{r exp10-summary-tibbles, echo=FALSE}
cat("Experiment 10 detection (Exp10- no cut off.xlsx)\n\n")
brief_dataframe(dplyr::count(df.10, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.10)

cat("\n\nReads — GVB_exp10_reads_tom.xlsx\n\n")
brief_dataframe(dplyr::count(df.10.reads, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.10.reads)

cat("\n\nReads — no time 0 (GVB_exp10_reads_tom_no_time_0.xlsx)\n\n")
brief_dataframe(dplyr::count(df.10.reads.no0, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.10.reads.no0)
```

</details>

::: {.panel-tabset}

```{r exp10-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

priors <- prior(normal(0,1.5), class = b)

mod.5 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.5.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod.6 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.6.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp10-waic-compute, include=FALSE}
waic_tbl_one_exp10 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp10_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_exp10(mod.5, "mod.5 (time monotonic, no cut)"),
  waic_tbl_one_exp10(mod.6, "mod.6 (time monotonic, cut)")
) %>%
  dplyr::arrange(WAIC, Model) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

## Effect plots

```{r exp10-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_5 <- conditional_effects(mod.5)[[1]] %>% mutate(cutoff = "No")
c_eff_6 <- conditional_effects(mod.6)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_5, c_eff_6)

raw_exp10_eff <- dplyr::bind_rows(
  dplyr::mutate(df.10, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.10, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
) %>%
  dplyr::mutate(y_plot = binary_y_for_plot(.data$y))

y_lim_exp10_eff1 <- c(-0.13, 1.13)
tlv_exp10_eff1 <- levels(c_eff$time)
n_t_exp10_eff1 <- length(tlv_exp10_eff1)
ichr_exp10 <- as.character(tlv_exp10_eff1)
i72_exp10_eff1 <- which(ichr_exp10 == "72")[1L]
if (is.na(i72_exp10_eff1)) {
  i72_exp10_eff1 <- which(suppressWarnings(as.numeric(ichr_exp10)) == 72)[1L]
}
if (is.na(i72_exp10_eff1)) i72_exp10_eff1 <- max(1L, round(n_t_exp10_eff1 / 2))
x_npc_exp10_eff1 <- (i72_exp10_eff1 - 0.5) / n_t_exp10_eff1
y_npc_exp10_eff1 <- (0.3 - y_lim_exp10_eff1[1]) / (y_lim_exp10_eff1[2] - y_lim_exp10_eff1[1])

p_exp10_eff <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp10_eff,
    aes(x = time, y = y_plot, group = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.03, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  theme(
    legend.position = c(x_npc_exp10_eff1, y_npc_exp10_eff1),
    legend.justification = c(0.5, 0.5),
    legend.margin = ggplot2::margin(4, 4, 4, 4),
    legend.background = element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = y_lim_exp10_eff1,
    breaks = prob_det_y_breaks,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title = "Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp10_effect_1.png"),
  plot = p_exp10_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_eff
```

Figure 5: Experiment 10, **monotonic time** (`mo(time)`). Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Black jittered points are raw binary outcomes by exposure time; error bars are 95% credible intervals and coloured points are posterior means.

## Tables

```{r exp10-estimates, echo=FALSE}
# Modeled estimates from conditional effects() on the probability scale.

ce10_time <- dplyr::bind_rows(
  conditional_effects(mod.5)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.6)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

knitr::kable(
  ce10_time,
  caption = "Modeled estimates from conditional effects (Experiment 10): time",
  format = "html"
)
```

## Model comparison

```{r exp10-waic-tab, echo=FALSE}
knitr::kable(
  exp10_waic_tbl,
  caption = "Experiment 10: WAIC (lower is better; `mo(time)` formulations).",
  format = "html"
)
```

## Models

```{r exp10-models-show, ref.label="exp10-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp10-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod.5, ndraws = 200, col = 2)
pp_check(mod.6, ndraws = 200, col = 2)
```

## Output

```{r exp10-summary, echo=FALSE}
summary(mod.5)
summary(mod.6)
```

<!-- Pairwise comparisons removed -->

```{r exp10-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.5, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  },
  {
    emm <- emmeans::emmeans(mod.6, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 10; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

### GVB sequence counts - reads

Negative binomial models on **`GVB_exp10_reads_tom.xlsx`** (`df.10.reads`): **`gvb_sequence_number`** with **`offset(log(total_reads))`**. If the sheet has **`quantity_gbv`** (or an alias **`wrangle_exp6_reads_sheet()`** recognizes), models match Experiment 6 (monotonic time × quantity + factor alternatives). Otherwise only **time** is modeled: **`mo(time)`** vs **factor `time`**. Exposure times use **`exp10_time_to_hours()`** (e.g. **1d → 24**). Rows require **`total_reads > 0`**; when present, **`quantity_gbv ≠ 0`**.

::: {.panel-tabset}

```{r exp10-reads-models-fit, include=FALSE}
if (exp10_reads_has_quantity) {
  d10r <- exp10_reads_filtered_for_nb(df.10.reads)
  t10_lv <- sort(unique(stats::na.omit(d10r$tnum)))
  if (length(t10_lv) < 2L) {
    stop(
      "Experiment 10 reads: need ≥ 2 time levels after filtering (quantity GVB, total reads). Found: ",
      paste(t10_lv, collapse = ", "),
      " (nrow = ", nrow(d10r), "). Check `time`/`time_hours` in the workbook and filters on total_reads / gvb_sequence_number."
    )
  }
  qty10_lv <- sort(unique(stats::na.omit(d10r$quantity_gbv)))
  df_exp10_reads_mod <- d10r %>%
    dplyr::mutate(
      time = ordered(.data$tnum, levels = t10_lv),
      quantity_gbv = ordered(.data$quantity_gbv, levels = qty10_lv)
    ) %>%
    dplyr::select(-tnum)

  qty_fac_lv_exp10_reads <- levels(df_exp10_reads_mod$quantity_gbv)
  df_exp10_reads_fac <- df_exp10_reads_mod %>%
    dplyr::mutate(
      time = factor(as.character(as.numeric(.data$time)), levels = as.character(t10_lv)),
      quantity_gbv = factor(as.character(as.numeric(.data$quantity_gbv)), levels = as.character(qty_fac_lv_exp10_reads))
    )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

  mod_exp10_reads <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_mod,
    file = paste0(here::here("outputs/models/mod_exp10_reads_gvbseq_negbin_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

  mod_exp10_reads_mo_int <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_mod,
    file = paste0(here::here("outputs/models/mod_exp10_reads_mo_interact_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

  mod_exp10_reads_fac_add <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_fac,
    file = paste0(here::here("outputs/models/mod_exp10_reads_fac_add_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

  mod_exp10_reads_fac_int <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_fac,
    file = paste0(here::here("outputs/models/mod_exp10_reads_fac_int_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )
} else {
  d10r <- exp10_reads_filtered_for_nb(df.10.reads)
  t10_lv <- sort(unique(stats::na.omit(d10r$tnum)))
  if (length(t10_lv) < 2L) {
    stop(
      "Experiment 10 reads: need ≥ 2 distinct time levels after filtering (total reads, counts). Found: ",
      paste(t10_lv, collapse = ", "),
      " (nrow = ", nrow(d10r), "). Check `time`/`time_hours` in the workbook and filters on total_reads / gvb_sequence_number."
    )
  }
  df_exp10_reads_mod <- d10r %>%
    dplyr::mutate(time = ordered(.data$tnum, levels = t10_lv)) %>%
    dplyr::select(-tnum)

  df_exp10_reads_fac <- df_exp10_reads_mod %>%
    dplyr::mutate(
      time = factor(as.character(as.numeric(.data$time)), levels = as.character(t10_lv))
    )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + offset(log(total_reads)))

  mod_exp10_reads <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_mod,
    file = paste0(here::here("outputs/models/mod_exp10_reads_mo_time_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_exp10_reads_mo_int <- mod_exp10_reads

  mod_formula <- bf(gvb_sequence_number ~ time + offset(log(total_reads)))

  mod_exp10_reads_fac_add <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_fac,
    file = paste0(here::here("outputs/models/mod_exp10_reads_fac_time_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_exp10_reads_fac_int <- mod_exp10_reads_fac_add
}
```

```{r exp10-reads-waic-compute, include=FALSE}
waic_tbl_one_exp10_reads <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

if (exp10_reads_has_quantity) {
  exp10_reads_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads(mod_exp10_reads, "mo(time) + mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads(mod_exp10_reads_mo_int, "mo(time) * mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads(mod_exp10_reads_fac_add, "time + quantity_gbv (factor)"),
    waic_tbl_one_exp10_reads(mod_exp10_reads_fac_int, "time * quantity_gbv (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
} else {
  exp10_reads_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads(mod_exp10_reads, "mo(time)"),
    waic_tbl_one_exp10_reads(mod_exp10_reads_fac_add, "time (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
}
```

#### Effect plots

```{r exp10-reads-eff, warning=FALSE, echo=FALSE, fig.width=8, fig.height=4.5, out.width="100%", dpi=600}
raw <- df_exp10_reads_mod
n_t_exp10 <- length(t10_lv)
xlim_exp10 <- exp10_reads_xlim_for_n(n_t_exp10)
x_lab_exp10 <- as.character(t10_lv)
leg_npc_exp10_f3 <- c(
  x = (n_t_exp10 - xlim_exp10[1]) / diff(xlim_exp10),
  y = (log1p(10) - log1p(exp10_reads_fig_y_lower_f3)) /
    (log1p(exp10_reads_fig_y_upper) - log1p(exp10_reads_fig_y_lower_f3))
)

if (exp10_reads_has_quantity) {
  ce_both <- conditional_effects(mod_exp10_reads_mo_int, effects = "time:quantity_gbv")[[1]]
  qty_lv <- sort(unique(as.numeric(as.character(ce_both$quantity_gbv))))
  ce_both <- ce_both %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv))
    ) %>%
    dplyr::mutate(
      tord = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = as.numeric(.data$time)
    ) %>%
    dplyr::arrange(.data$quantity_gbv, .data$tord)

  raw_col <- raw %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both$quantity_gbv)),
      xtime = as.numeric(.data$time) + exp10_reads_fig_x_nudge_raw
    )

  n_qty <- dplyr::n_distinct(ce_both$quantity_gbv)
  pal <- exp10_reads_qty_palette(n_qty)
  ce_pd <- ggplot2::position_dodge(width = 0.85)

  exp10_reads_plot_theme <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
      legend.position = unname(leg_npc_exp10_f3),
      legend.justification = c(0.5, 0.5),
      legend.title = ggplot2::element_text(size = 10),
      legend.text = ggplot2::element_text(size = 9),
      legend.margin = ggplot2::margin(2, 4, 2, 4),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
    )

  p_exp10_reads_eff <- ggplot2::ggplot(
    ce_both,
    ggplot2::aes(
      x = .data$xtime,
      y = .data$estimate__,
      ymin = .data$lower__,
      ymax = .data$upper__,
      color = .data$quantity_gbv
    )
  ) +
    ggplot2::geom_point(
      data = raw_col,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
      inherit.aes = FALSE,
      position = ggplot2::position_jitterdodge(jitter.width = 0.12, jitter.height = 0.175, dodge.width = 0.85),
      alpha = 0.4,
      size = 0.85,
      stroke = 0
    ) +
    ggplot2::geom_errorbar(position = ce_pd, width = 0.12, linewidth = 0.9) +
    ggplot2::geom_point(position = ce_pd, size = 2.5, alpha = 0.9) +
    ggplot2::scale_color_manual(values = pal) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10),
      labels = x_lab_exp10,
      limits = xlim_exp10,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(exp10_reads_fig_y_lower_f3, exp10_reads_fig_y_upper),
      breaks = exp10_reads_y_breaks_use,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
    ggplot2::guides(color = ggplot2::guide_legend(title = "GVB quantity", reverse = TRUE))

  fn_exp10_reads_plot <- here::here("outputs/plots", "exp10_reads_effect_time_by_quantity.png")
} else {
  ce_time <- conditional_effects(mod_exp10_reads, effects = "time")[[1]] %>%
    dplyr::mutate(xtime = as.numeric(.data$time))

  raw_col <- raw %>%
    dplyr::mutate(xtime = as.numeric(.data$time) + exp10_reads_fig_x_nudge_raw)

  exp10_reads_plot_theme <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
    )

  col_mu <- "#0072B2"
  p_exp10_reads_eff <- ggplot2::ggplot(
    ce_time,
    ggplot2::aes(x = .data$xtime, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
  ) +
    ggplot2::geom_point(
      data = raw_col,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number),
      inherit.aes = FALSE,
      position = ggplot2::position_jitter(width = 0.1, height = 0.175),
      alpha = 0.4,
      size = 0.85,
      color = "grey40",
      stroke = 0
    ) +
    ggplot2::geom_errorbar(width = 0.12, linewidth = 0.9, color = col_mu) +
    ggplot2::geom_point(size = 2.5, alpha = 0.9, color = col_mu) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10),
      labels = x_lab_exp10,
      limits = xlim_exp10,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(exp10_reads_fig_y_lower_f3, exp10_reads_fig_y_upper),
      breaks = exp10_reads_y_breaks_use,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

  fn_exp10_reads_plot <- here::here("outputs/plots", "exp10_reads_effect_time.png")
}

ggplot2::ggsave(
  filename = fn_exp10_reads_plot,
  plot = p_exp10_reads_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_reads_eff
```

```{r exp10-reads-fig6-caption, echo=FALSE, results="asis"}
if (exp10_reads_has_quantity) {
  cat("Figure 6: Experiment 10, **GVB sequence counts** (reads workbook). Conditional effect of GVB exposure on mean sequence count by GVB quantity under **`mo(time) × mo(quantity)`** with **`offset(log(total_reads))`**. Vertical axis is **log(1 + count)** (lower limit slightly below zero). Jittered points are raw counts; error bars are 95% credible intervals and coloured points are posterior means.\n\n")
} else {
  cat("Figure 6: Experiment 10, **GVB sequence counts** (reads workbook). Conditional effect of GVB exposure on mean sequence count from **`mo(time)`** with **`offset(log(total_reads))`** (no **quantity GVB** column in this sheet). Vertical axis is **log(1 + count)** (lower limit slightly below zero). Grey jittered points are raw counts; blue error bars and points are 95% credible intervals and posterior means.\n\n")
}
```

#### Tables

```{r exp10-reads-estimates, echo=FALSE}
if (exp10_reads_has_quantity) {
  ce_exp10_t <- conditional_effects(mod_exp10_reads_mo_int, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  ce_exp10_q <- conditional_effects(mod_exp10_reads_mo_int, effects = "quantity_gbv")[[1]] %>%
    dplyr::transmute(
      `Quantity GVB` = quantity_gbv,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_t,
    caption = "Experiment 10 reads — primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
    format = "html"
  )

  knitr::kable(
    ce_exp10_q,
    caption = "Experiment 10 reads — primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
    format = "html"
  )
} else {
  ce_exp10_t <- conditional_effects(mod_exp10_reads, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_t,
    caption = "Experiment 10 reads — model `mo(time)` with offset: conditional effect of time (mean count).",
    format = "html"
  )
}
```

#### Models

```{r exp10-reads-models-show, ref.label="exp10-reads-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp10-reads-waic-tab, echo=FALSE}
knitr::kable(
  exp10_reads_waic_tbl,
  caption = "Experiment 10 reads: WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
```

#### Validation

```{r exp10-reads-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
if (exp10_reads_has_quantity) {
  pp_check(mod_exp10_reads, ndraws = 200)
  pp_check(mod_exp10_reads_mo_int, ndraws = 200)
  pp_check(mod_exp10_reads_fac_add, ndraws = 200)
  pp_check(mod_exp10_reads_fac_int, ndraws = 200)
} else {
  pp_check(mod_exp10_reads, ndraws = 200)
  pp_check(mod_exp10_reads_fac_add, ndraws = 200)
}
```

#### Output

```{r exp10-reads-summary, echo=FALSE}
if (exp10_reads_has_quantity) {
  summary(mod_exp10_reads)
  summary(mod_exp10_reads_mo_int)
  summary(mod_exp10_reads_fac_add)
  summary(mod_exp10_reads_fac_int)
} else {
  summary(mod_exp10_reads)
  summary(mod_exp10_reads_fac_add)
}
```

:::

### GVB sequence counts - no time 0

::: {.panel-tabset}

```{r exp10-reads-no0-models-fit, include=FALSE}
if (exp10_reads_has_quantity) {
  d10rn0 <- exp10_reads_filtered_for_nb(df.10.reads.no0)
  t10n0_lv <- sort(unique(stats::na.omit(d10rn0$tnum)))
  if (length(t10n0_lv) < 2L) {
    stop(
      "Experiment 10 reads (no time 0): need ≥ 2 time levels after filtering. Found: ",
      paste(t10n0_lv, collapse = ", "),
      " (nrow = ", nrow(d10rn0), ")."
    )
  }
  qty10n0_lv <- sort(unique(stats::na.omit(d10rn0$quantity_gbv)))
  df_exp10_reads_no0_mod <- d10rn0 %>%
    dplyr::mutate(
      time = ordered(.data$tnum, levels = t10n0_lv),
      quantity_gbv = ordered(.data$quantity_gbv, levels = qty10n0_lv)
    ) %>%
    dplyr::select(-tnum)

  qty_fac_lv_exp10_no0 <- levels(df_exp10_reads_no0_mod$quantity_gbv)
  df_exp10_reads_no0_fac <- df_exp10_reads_no0_mod %>%
    dplyr::mutate(
      time = factor(as.character(as.numeric(.data$time)), levels = as.character(t10n0_lv)),
      quantity_gbv = factor(as.character(as.numeric(.data$quantity_gbv)), levels = as.character(qty_fac_lv_exp10_no0))
    )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

  mod_exp10_reads_no0 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_mod,
    file = paste0(here::here("outputs/models/mod_exp10_reads_no0_gvbseq_negbin_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

  mod_exp10_reads_no0_mo_int <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_mod,
    file = paste0(here::here("outputs/models/mod_exp10_reads_no0_mo_interact_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

  mod_exp10_reads_no0_fac_add <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_fac,
    file = paste0(here::here("outputs/models/mod_exp10_reads_no0_fac_add_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

  mod_exp10_reads_no0_fac_int <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_fac,
    file = paste0(here::here("outputs/models/mod_exp10_reads_no0_fac_int_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )
} else {
  d10rn0 <- exp10_reads_filtered_for_nb(df.10.reads.no0)
  t10n0_lv <- sort(unique(stats::na.omit(d10rn0$tnum)))
  if (length(t10n0_lv) < 2L) {
    stop(
      "Experiment 10 reads (no time 0): need ≥ 2 time levels after filtering. Found: ",
      paste(t10n0_lv, collapse = ", "),
      " (nrow = ", nrow(d10rn0), ")."
    )
  }
  df_exp10_reads_no0_mod <- d10rn0 %>%
    dplyr::mutate(time = ordered(.data$tnum, levels = t10n0_lv)) %>%
    dplyr::select(-tnum)

  df_exp10_reads_no0_fac <- df_exp10_reads_no0_mod %>%
    dplyr::mutate(
      time = factor(as.character(as.numeric(.data$time)), levels = as.character(t10n0_lv))
    )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + offset(log(total_reads)))

  mod_exp10_reads_no0 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_mod,
    file = paste0(here::here("outputs/models/mod_exp10_reads_no0_mo_time_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_exp10_reads_no0_mo_int <- mod_exp10_reads_no0

  mod_formula <- bf(gvb_sequence_number ~ time + offset(log(total_reads)))

  mod_exp10_reads_no0_fac_add <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_fac,
    file = paste0(here::here("outputs/models/mod_exp10_reads_no0_fac_time_offset")),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_exp10_reads_no0_fac_int <- mod_exp10_reads_no0_fac_add
}
```

```{r exp10-reads-no0-waic-compute, include=FALSE}
waic_tbl_one_exp10_reads_no0 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

if (exp10_reads_has_quantity) {
  exp10_reads_no0_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads_no0(mod_exp10_reads_no0, "mo(time) + mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads_no0(mod_exp10_reads_no0_mo_int, "mo(time) * mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads_no0(mod_exp10_reads_no0_fac_add, "time + quantity_gbv (factor)"),
    waic_tbl_one_exp10_reads_no0(mod_exp10_reads_no0_fac_int, "time * quantity_gbv (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
} else {
  exp10_reads_no0_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads_no0(mod_exp10_reads_no0, "mo(time)"),
    waic_tbl_one_exp10_reads_no0(mod_exp10_reads_no0_fac_add, "time (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
}
```

#### Effect plots

```{r exp10-reads-no0-eff, warning=FALSE, echo=FALSE, fig.width=8, fig.height=4.5, out.width="100%", dpi=600}
raw_no0 <- df_exp10_reads_no0_mod
n_t_exp10_n0 <- length(t10n0_lv)
xlim_exp10_n0 <- exp10_reads_xlim_for_n(n_t_exp10_n0)
x_lab_exp10_n0 <- as.character(t10n0_lv)

if (exp10_reads_has_quantity) {
  ce_both_no0 <- conditional_effects(mod_exp10_reads_no0_mo_int, effects = "time:quantity_gbv")[[1]]
  qty_lv_no0 <- sort(unique(as.numeric(as.character(ce_both_no0$quantity_gbv))))
  ce_both_no0 <- ce_both_no0 %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv_no0))
    ) %>%
    dplyr::mutate(
      tord = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = as.numeric(.data$time)
    ) %>%
    dplyr::arrange(.data$quantity_gbv, .data$tord)

  raw_col_no0 <- raw_no0 %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both_no0$quantity_gbv)),
      xtime = as.numeric(.data$time) + exp10_reads_fig_x_nudge_raw
    )

  exp10_reads_fig_y_max_no0 <- max(
    0,
    suppressWarnings(max(as.numeric(ce_both_no0$upper__), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(ce_both_no0$estimate__), na.rm = TRUE)),
    suppressWarnings(max(raw_col_no0$gvb_sequence_number, na.rm = TRUE)),
    na.rm = TRUE
  )
  exp10_reads_fig_y_upper_no0 <- max(
    exp10_reads_fig_y_upper,
    exp10_reads_fig_y_max_no0 * 1.72 + max(80, 0.12 * exp10_reads_fig_y_max_no0) + max(40, 0.04 * exp10_reads_fig_y_max_no0),
    10
  )
  exp10_reads_y_breaks_use_no0 <- exp10_reads_y_breaks[exp10_reads_y_breaks <= exp10_reads_fig_y_upper_no0]
  exp10_reads_fig_legend_npc_no0 <- c(
    x = (n_t_exp10_n0 - xlim_exp10_n0[1]) / diff(xlim_exp10_n0),
    y = log1p(10) / log1p(exp10_reads_fig_y_upper_no0)
  )

  n_qty_no0 <- dplyr::n_distinct(ce_both_no0$quantity_gbv)
  pal_no0 <- exp10_reads_qty_palette(n_qty_no0)
  ce_pd_no0 <- ggplot2::position_dodge(width = 0.85)

  exp10_reads_plot_theme_no0 <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
      legend.position = unname(exp10_reads_fig_legend_npc_no0),
      legend.justification = c(0.5, 0.5),
      legend.title = ggplot2::element_text(size = 10),
      legend.text = ggplot2::element_text(size = 9),
      legend.margin = ggplot2::margin(2, 4, 2, 4),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
    )

  p_exp10_reads_no0_eff <- ggplot2::ggplot(
    ce_both_no0,
    ggplot2::aes(
      x = .data$xtime,
      y = .data$estimate__,
      ymin = .data$lower__,
      ymax = .data$upper__,
      color = .data$quantity_gbv
    )
  ) +
    ggplot2::geom_point(
      data = raw_col_no0,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
      inherit.aes = FALSE,
      position = ggplot2::position_jitterdodge(jitter.width = 0.12, jitter.height = 0.175, dodge.width = 0.85),
      alpha = 0.4,
      size = 0.85,
      stroke = 0
    ) +
    ggplot2::geom_errorbar(position = ce_pd_no0, width = 0.12, linewidth = 0.9) +
    ggplot2::geom_point(position = ce_pd_no0, size = 2.5, alpha = 0.9) +
    ggplot2::scale_color_manual(values = pal_no0) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10_n0),
      labels = x_lab_exp10_n0,
      limits = xlim_exp10_n0,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(0, exp10_reads_fig_y_upper_no0),
      breaks = exp10_reads_y_breaks_use_no0,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme_no0 +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
    ggplot2::guides(color = ggplot2::guide_legend(title = "GVB quantity", reverse = TRUE))

  fn_exp10_no0_plot <- here::here("outputs/plots", "exp10_reads_no0_effect_time_by_quantity.png")
} else {
  ce_both_no0 <- conditional_effects(mod_exp10_reads_no0, effects = "time")[[1]] %>%
    dplyr::mutate(xtime = as.numeric(.data$time))

  raw_col_no0 <- raw_no0 %>%
    dplyr::mutate(xtime = as.numeric(.data$time) + exp10_reads_fig_x_nudge_raw)

  exp10_reads_fig_y_max_no0 <- max(
    0,
    suppressWarnings(max(as.numeric(ce_both_no0$upper__), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(ce_both_no0$estimate__), na.rm = TRUE)),
    suppressWarnings(max(raw_col_no0$gvb_sequence_number, na.rm = TRUE)),
    na.rm = TRUE
  )
  exp10_reads_fig_y_upper_no0 <- max(
    exp10_reads_fig_y_upper,
    exp10_reads_fig_y_max_no0 * 1.72 + max(80, 0.12 * exp10_reads_fig_y_max_no0) + max(40, 0.04 * exp10_reads_fig_y_max_no0),
    10
  )
  exp10_reads_y_breaks_use_no0 <- exp10_reads_y_breaks[exp10_reads_y_breaks <= exp10_reads_fig_y_upper_no0]

  col_mu_n0 <- "#0072B2"
  exp10_reads_plot_theme_no0 <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
    )

  p_exp10_reads_no0_eff <- ggplot2::ggplot(
    ce_both_no0,
    ggplot2::aes(x = .data$xtime, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
  ) +
    ggplot2::geom_point(
      data = raw_col_no0,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number),
      inherit.aes = FALSE,
      position = ggplot2::position_jitter(width = 0.1, height = 0.175),
      alpha = 0.4,
      size = 0.85,
      color = "grey40",
      stroke = 0
    ) +
    ggplot2::geom_errorbar(width = 0.12, linewidth = 0.9, color = col_mu_n0) +
    ggplot2::geom_point(size = 2.5, alpha = 0.9, color = col_mu_n0) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10_n0),
      labels = x_lab_exp10_n0,
      limits = xlim_exp10_n0,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(0, exp10_reads_fig_y_upper_no0),
      breaks = exp10_reads_y_breaks_use_no0,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme_no0 +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

  fn_exp10_no0_plot <- here::here("outputs/plots", "exp10_reads_no0_effect_time.png")
}

ggplot2::ggsave(
  filename = fn_exp10_no0_plot,
  plot = p_exp10_reads_no0_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_reads_no0_eff
```

```{r exp10-reads-fig7-caption, echo=FALSE, results="asis"}
if (exp10_reads_has_quantity) {
  cat("Figure 7: Experiment 10, **GVB sequence counts** (reads workbook without time 0). Same design as Figure 6. Source workbook **`GVB_exp10_reads_tom_no_time0.xlsx`** omits 0 h exposure rows where applicable.\n\n")
} else {
  cat("Figure 7: Experiment 10, **GVB sequence counts** (no time 0 workbook). **`mo(time)`** and **`offset(log(total_reads))`** (no quantity column). Same plotting style as Figure 6 time-only panel.\n\n")
}
```

#### Tables

```{r exp10-reads-no0-estimates, echo=FALSE}
if (exp10_reads_has_quantity) {
  ce_exp10_no0_t <- conditional_effects(mod_exp10_reads_no0_mo_int, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  ce_exp10_no0_q <- conditional_effects(mod_exp10_reads_no0_mo_int, effects = "quantity_gbv")[[1]] %>%
    dplyr::transmute(
      `Quantity GVB` = quantity_gbv,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_no0_t,
    caption = "Experiment 10 reads (no time 0) — primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
    format = "html"
  )

  knitr::kable(
    ce_exp10_no0_q,
    caption = "Experiment 10 reads (no time 0) — primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
    format = "html"
  )
} else {
  ce_exp10_no0_t <- conditional_effects(mod_exp10_reads_no0, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_no0_t,
    caption = "Experiment 10 reads (no time 0) — model `mo(time)` with offset: conditional effect of time (mean count).",
    format = "html"
  )
}
```

#### Models

```{r exp10-reads-no0-models-show, ref.label="exp10-reads-no0-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp10-reads-no0-waic-tab, echo=FALSE}
knitr::kable(
  exp10_reads_no0_waic_tbl,
  caption = "Experiment 10 reads (no time 0): WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
```

#### Validation

```{r exp10-reads-no0-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
if (exp10_reads_has_quantity) {
  pp_check(mod_exp10_reads_no0, ndraws = 200)
  pp_check(mod_exp10_reads_no0_mo_int, ndraws = 200)
  pp_check(mod_exp10_reads_no0_fac_add, ndraws = 200)
  pp_check(mod_exp10_reads_no0_fac_int, ndraws = 200)
} else {
  pp_check(mod_exp10_reads_no0, ndraws = 200)
  pp_check(mod_exp10_reads_no0_fac_add, ndraws = 200)
}
```

#### Output

```{r exp10-reads-no0-summary, echo=FALSE}
if (exp10_reads_has_quantity) {
  summary(mod_exp10_reads_no0)
  summary(mod_exp10_reads_no0_mo_int)
  summary(mod_exp10_reads_no0_fac_add)
  summary(mod_exp10_reads_no0_fac_int)
} else {
  summary(mod_exp10_reads_no0)
  summary(mod_exp10_reads_no0_fac_add)
}
```

:::

## QPCR experiment

<details><summary>Data frames</summary>
```{r}
#read in data ----
df <- read_excel(here::here("analysis/data/qPCR_results251017_forTom.xlsx"), sheet = 1) # Import .xlsx file

df <- df %>% clean_names() %>% select(-location_in_orchard, -note)

cols <- c("sample", "method", "method_name")
df[, cols] <- lapply(df[, cols], as.factor) # Coerce variables to factors

cols <- c("col", "row", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml", "cq1", "cq2", "cq3")
df[, cols] <- lapply(df[, cols], as.numeric) # Coerce variables to factors

df <- reshape2::melt(
  df,
  id = c("sample", "col", "row", "method", "method_name", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml")
) # Make it long

df <- df |>
  rename(replicate = variable, cq = value) |>
  filter(!is.na(row)) |>
  mutate(week_ord = ordered(week)) |> # ordered factor
  mutate(method = as.factor(make_clean_names(method, allow_dupes = TRUE))) |>
  mutate(method_name = as.factor(make_clean_names(method_name, allow_dupes = TRUE))) |>
  mutate(cq = if_else(is.nan(cq), 0, cq)) |>
  mutate(pa = ifelse(cq > 0, 1, cq)) |> # Create binary variable
  filter(method %in% c("b", "c", "d")) |>
  mutate(sub_method = paste0(method, sep = "_", method_name)) |>
  droplevels()

brief_dataframe(df)
```
</details>

::: {.panel-tabset}

```{r qpcr-models-fit-submethod, include=FALSE}
priors <- prior(normal(0, 1.5), class = b)

mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, id = "sub_method") + (1|replicate))

mod_3 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.9.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: linear (separate slopes per sub-method)
mod_formula <- bf(pa ~ sub_method * increase_in_pvh_number + (1|replicate))

mod_4 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.10.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: one shared smooth; sub-method shifts intercept only
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number) + (1|replicate))

mod_5 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.11.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: separate smooths per sub-method (mgcv-style by)
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, by = sub_method) + (1|replicate))

mod_6 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.12.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# + monotonic week within sub_method (parallel to mod_3–mod_6)
mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, id = "sub_method") + (1|replicate)
)

mod_7 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.14.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method * increase_in_pvh_number + mo(week_ord, id = "sub_method") + (1|replicate)
)

mod_8 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.15.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number) + (1|replicate)
)

mod_9 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.16.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, by = sub_method) + (1|replicate)
)

mod_10 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.17.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × monotonic week + separate PVH smooths per sub-method
mod_formula <- bf(
  pa ~ sub_method * mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, by = sub_method) + (1|replicate)
)

mod_11 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.18.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)
```

```{r qpcr-eff-helpers, include=FALSE}
qpcr_sub_method_label <- function(x) {
  lbl <- vapply(as.character(x), function(s) {
    p <- strsplit(s, "_", fixed = TRUE)[[1]]
    if (length(p) == 0L) {
      return(s)
    }
    letter_lab <- dplyr::case_match(
      p[[1]],
      "b" ~ "B",
      "c" ~ "C",
      "d" ~ "D",
      .default = stringr::str_to_title(p[[1]])
    )
    if (length(p) == 1L) {
      return(paste0("Method ", letter_lab))
    }
    rest <- stringr::str_replace_all(paste(p[-1L], collapse = "_"), "_", " ")
    rest <- stringr::str_to_title(rest)
    paste0("Method ", letter_lab, " — ", rest)
  }, character(1))
  factor(lbl, levels = sort(unique(lbl)))
}

# Shared qPCR figure: PVH × sub_method conditional effects (ghost + focal facets)
qpcr_plot_pvh_by_submethod <- function(fit) {
  c_eff <- conditional_effects(fit, effects = "increase_in_pvh_number:sub_method")[[1]] %>%
    dplyr::arrange(sub_method, increase_in_pvh_number) %>%
    dplyr::mutate(sub_method_l = qpcr_sub_method_label(sub_method))
  panel_levels <- levels(c_eff$sub_method_l)
  ghost <- purrr::map_dfr(panel_levels, function(f) {
    c_eff %>%
      dplyr::filter(as.character(.data$sub_method_l) != as.character(f)) %>%
      dplyr::mutate(facet_panel = factor(f, levels = panel_levels))
  })
  focal <- c_eff %>%
    dplyr::mutate(facet_panel = factor(.data$sub_method_l, levels = panel_levels))
  obs_raw <- fit$data %>%
    dplyr::mutate(
      sub_method_l = qpcr_sub_method_label(.data$sub_method),
      facet_panel = factor(.data$sub_method_l, levels = panel_levels),
      pa_plot = binary_y_for_plot(.data$pa)
    )
  ggplot2::ggplot() +
    ggplot2::geom_jitter(
      data = obs_raw,
      ggplot2::aes(x = increase_in_pvh_number, y = pa_plot),
      inherit.aes = FALSE,
      height = 0.035,
      width = 0.15,
      alpha = 0.35,
      size = 0.85,
      color = "black",
      stroke = 0
    ) +
    ggplot2::geom_ribbon(
      data = ghost,
      ggplot2::aes(
        x = increase_in_pvh_number, ymin = lower__, ymax = upper__,
        group = sub_method_l
      ),
      inherit.aes = FALSE,
      fill = "gray45",
      alpha = 0.04
    ) +
    ggplot2::geom_line(
      data = ghost,
      ggplot2::aes(x = increase_in_pvh_number, y = estimate__, group = sub_method_l),
      inherit.aes = FALSE,
      color = "gray40",
      linewidth = 0.5,
      alpha = 0.1
    ) +
    ggplot2::geom_ribbon(
      data = focal,
      ggplot2::aes(
        x = increase_in_pvh_number, ymin = lower__, ymax = upper__,
        fill = sub_method_l, group = sub_method_l
      ),
      inherit.aes = FALSE,
      alpha = 0.2,
      color = NA
    ) +
    ggplot2::geom_line(
      data = focal,
      ggplot2::aes(
        x = increase_in_pvh_number, y = estimate__,
        color = sub_method_l, group = sub_method_l
      ),
      inherit.aes = FALSE,
      linewidth = 1
    ) +
    ggplot2::facet_wrap(ggplot2::vars(facet_panel), ncol = 3, scales = "fixed") +
    ggplot2::theme_classic() +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 9),
      legend.position = "none"
    ) +
    ggplot2::scale_y_continuous(
      labels = scales::percent,
      limits = c(-0.13, 1.13),
      breaks = prob_det_y_breaks,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    ggplot2::scale_color_brewer(palette = "Dark2", type = "qual") +
    ggplot2::scale_fill_brewer(palette = "Dark2", type = "qual") +
    ggplot2::labs(x = "PVH number", y = "Probability of detection (%)")
}

# Week × sub_method: posterior mean response at each (week_ord, sub_method), holding
# increase_in_pvh_number at its within–sub-method median and replicate at a reference level.
# Uses posterior_epred() so it behaves for all PVH structures (linear, shared spline, `by`)
# — conditional_effects("week_ord:sub_method") is unreliable for e.g. mod_10 / mod_11.
qpcr_plot_week_by_submethod <- function(fit) {
  d <- fit$data

  wk <- d$week_ord
  if (is.factor(wk)) {
    wk_lvls <- levels(wk)
    wk_ordered <- is.ordered(wk)
  } else {
    wk_lvls <- unique(as.character(wk[!is.na(wk)]))
    wk_ordered <- TRUE
  }

  sm <- d$sub_method
  if (!is.factor(sm)) {
    sm <- factor(sm, levels = unique(as.character(sm[!is.na(sm)])))
  } else {
    sm <- droplevels(sm)
  }
  sm_lvls <- levels(sm)

  rp <- d$replicate
  if (!is.factor(rp)) {
    rp <- factor(rp, levels = unique(as.character(rp[!is.na(rp)])))
  } else {
    rp <- droplevels(rp)
  }
  rep_lvls <- levels(rp)
  rep_ref <- rep_lvls[[1L]]

  if (length(wk_lvls) < 1L || length(sm_lvls) < 1L) {
    stop("fit$data is missing week_ord or sub_method factor levels.")
  }
  med_pvh <- tapply(d$increase_in_pvh_number, sm, stats::median, na.rm = TRUE)
  med_pvh <- med_pvh[sm_lvls]
  if (any(is.na(med_pvh))) {
    stop("Could not compute median PVH for every sub_method level.")
  }
  rows <- tidyr::expand_grid(
    week_ord = factor(wk_lvls, levels = wk_lvls, ordered = wk_ordered),
    sub_method = factor(sm_lvls, levels = sm_lvls)
  )
  rows$increase_in_pvh_number <- as.numeric(med_pvh[as.character(rows$sub_method)])
  rows$replicate <- factor(rep_ref, levels = rep_lvls)

  nd <- tryCatch(
    min(1000L, as.integer(brms::nsamples(fit))),
    error = function(...) 1000L
  )
  if (!is.finite(nd) || nd < 1L) nd <- 1000L
  ep <- brms::posterior_epred(
    object = fit,
    newdata = rows,
    re_formula = NA,
    ndraws = nd
  )
  rq <- apply(ep, 2L, stats::quantile, probs = c(0.025, 0.5, 0.975), na.rm = TRUE)
  c_eff <- rows %>%
    dplyr::mutate(
      lower__ = as.numeric(rq[1, ]),
      estimate__ = as.numeric(rq[2, ]),
      upper__ = as.numeric(rq[3, ])
    ) %>%
    dplyr::filter(is.finite(estimate__)) %>%
    dplyr::mutate(sub_method_l = qpcr_sub_method_label(sub_method))
  if (nrow(c_eff) < 1L) {
    stop("posterior_epred returned no usable week × sub-method estimates.")
  }

  panel_levels <- levels(c_eff$sub_method_l)
  plot_dat <- c_eff %>%
    dplyr::mutate(facet_panel = factor(.data$sub_method_l, levels = panel_levels))
  obs_raw <- d %>%
    dplyr::mutate(
      week_ord = factor(as.character(.data$week_ord), levels = wk_lvls, ordered = wk_ordered),
      sub_method_l = qpcr_sub_method_label(.data$sub_method),
      facet_panel = factor(.data$sub_method_l, levels = panel_levels),
      pa_plot = binary_y_for_plot(.data$pa)
    )

  n_col <- length(unique(c_eff$sub_method_l))
  p <- ggplot2::ggplot(plot_dat) +
    ggplot2::geom_jitter(
      data = obs_raw,
      ggplot2::aes(x = week_ord, y = pa_plot),
      inherit.aes = FALSE,
      height = 0.035,
      width = 0.18,
      alpha = 0.35,
      size = 0.85,
      color = "black",
      stroke = 0
    ) +
    ggplot2::geom_linerange(
      ggplot2::aes(
        x = week_ord,
        ymin = lower__, ymax = upper__,
        color = sub_method_l,
        group = week_ord
      ),
      linewidth = 1
    ) +
    ggplot2::geom_point(
      ggplot2::aes(
        x = week_ord, y = estimate__,
        color = sub_method_l,
        group = week_ord
      ),
      size = 3,
      alpha = 0.55
    ) +
    ggplot2::facet_wrap(ggplot2::vars(facet_panel), ncol = 3, scales = "fixed") +
    ggplot2::theme_classic() +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 9),
      legend.position = "none",
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.3),
      panel.grid.major.y = ggplot2::element_line(colour = "grey85", linewidth = 0.3)
    ) +
    ggplot2::scale_y_continuous(
      labels = scales::percent,
      limits = c(-0.13, 1.13),
      breaks = prob_det_y_breaks,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    ggplot2::labs(x = "Week", y = "Probability of detection (%)")
  if (n_col <= 8L) {
    p <- p + ggplot2::scale_color_brewer(palette = "Dark2", type = "qual")
  } else {
    p <- p + ggplot2::scale_color_hue(l = 55, c = 80)
  }
  p
}
```

```{r qpcr-waic-compute, include=FALSE}
waic_tbl_one <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

waic_tbl_pre <- dplyr::bind_rows(
  waic_tbl_one(mod_3, "mod_3"),
  waic_tbl_one(mod_4, "mod_4"),
  waic_tbl_one(mod_5, "mod_5"),
  waic_tbl_one(mod_6, "mod_6"),
  waic_tbl_one(mod_7, "mod_7"),
  waic_tbl_one(mod_8, "mod_8"),
  waic_tbl_one(mod_9, "mod_9"),
  waic_tbl_one(mod_10, "mod_10"),
  waic_tbl_one(mod_11, "mod_11")
) %>%
  dplyr::mutate(model_id = as.integer(sub("^mod_", "", Model))) %>%
  dplyr::arrange(WAIC, model_id)

waic_tbl <- waic_tbl_pre %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)

qpcr_model_formula_lines <- c(
  "3" = "bf(pa ~ sub_method + s(increase_in_pvh_number, id = \"sub_method\") + (1|replicate))",
  "4" = "bf(pa ~ sub_method * increase_in_pvh_number + (1|replicate))",
  "5" = "bf(pa ~ sub_method + s(increase_in_pvh_number) + (1|replicate))",
  "6" = "bf(pa ~ sub_method + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "7" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, id = \"sub_method\") + (1|replicate))",
  "8" = "bf(pa ~ sub_method * increase_in_pvh_number + mo(week_ord, id = \"sub_method\") + (1|replicate))",
  "9" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number) + (1|replicate))",
  "10" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "11" = "bf(pa ~ sub_method * mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, by = sub_method) + (1|replicate))"
)
```

## Effect plots

Side-by-side tabs for **`mod_8`** and **`mod_10`**. Each tab shows **week × sub-method** (`posterior_epred` on a full week × sub-method grid: median PVH per sub-method, reference replicate, `re_formula = NA`) and **PVH number × sub-method** (`conditional_effects`; other predictors at **brms** defaults). Figures are written to `outputs/plots/` as `qpcr_effect_mod8_*` and `qpcr_effect_mod10_*`.

::: {.panel-tabset}

## Effect plots (mod 8)

```{r qpcr-effect-mod8, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_8
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["8"]], "\n```\n\n", sep = "")
cat("**Week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod8_week.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n**PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod8_pvh.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
```

## Effect plots (mod 10)

```{r qpcr-effect-mod10, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_10
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["10"]], "\n```\n\n", sep = "")
cat("**Week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod10_week.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n**PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod10_pvh.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
```

:::

## Models

Sub-method formulations (`mod_3`–`mod_11`; includes monotonic week in `mod_7`–`mod_11`):

```{r qpcr-models-show, ref.label="qpcr-models-fit-submethod", echo=TRUE, eval=FALSE}

```

## Model Comparison

```{r qpcr-waic, echo=FALSE}
knitr::kable(
  waic_tbl,
  caption = "Model comparison (qPCR Experiment): WAIC (lower is better).",
  format = "html"
)
```

## Validation

```{r qpcr-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_3, ndraws = 200, col = 2)
pp_check(mod_4, ndraws = 200, col = 2)
pp_check(mod_5, ndraws = 200, col = 2)
pp_check(mod_6, ndraws = 200, col = 2)
pp_check(mod_7, ndraws = 200, col = 2)
pp_check(mod_8, ndraws = 200, col = 2)
pp_check(mod_9, ndraws = 200, col = 2)
pp_check(mod_10, ndraws = 200, col = 2)
pp_check(mod_11, ndraws = 200, col = 2)
```

## Output

```{r qpcr-summary, echo=FALSE}
summary(mod_3)
summary(mod_4)
summary(mod_5)
summary(mod_6)
summary(mod_7)
summary(mod_8)
summary(mod_9)
summary(mod_10)
summary(mod_11)
```


:::

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Session Information</summary>
```{r}
utils::sessionInfo()
```
</details>

