---
title: "GBV Detection Analysis"
subtitle: "*Document structure and formatting were assisted with Cursor.*"
author: "Dr Tom Moore - Statistical Scientist"
date: "`r Sys.Date()`"
format:
  html:
    include-in-header:
      text: |
        <style>
        .quarto-title .subtitle,
        p.subtitle,
        .subtitle.lead {
          font-size: 0.88rem;
          font-weight: normal;
        }
        </style>
    df-print: paged
    code-fold: false
    toc: true
    toc-float: true
    toc-depth: 5
    theme: united
    highlight-style: tango
    fig-width: 9
    fig-height: 6
    fig-dpi: 600
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup-packages, message=FALSE, warning=FALSE}
# Install packages and functions (based on existing Rmd setups)
ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  invisible(sapply(pkg, require, character.only = TRUE))
}

packages <- c(
  "tidyverse", "readxl", "janitor", "here", "reshape2", "brms", "cmdstanr", "bayesplot", "scales", "knitr"
)

ipak(packages)

bayesplot::color_scheme_set("blue")
options(mc.cores = parallel::detectCores())

# Ensure plot output directory exists
dir.create(here::here("outputs/plots"), recursive = TRUE, showWarnings = FALSE)

# Production Sever Settings
cmdstanr::set_cmdstan_path("/powerplant/workspace/hrltxm/workbench-k8s/stan/vanilla_stan/cmdstan-2.36.0")
```

# Methods:

To compare the probability of detection (%) of GVB at different exposure levels (0, 2, 6, 16, 24, 48 hrs), with and without a Cq cut off (i.e., if a Cq value was greater than 39 then classify as no detection), and GVB density (where data was available: at 2hr and 24hrs), generalized linear mixed effect (GLMM) models were performed in R (R Core Team 2022) using the brms package (Bürkner, 2017).

For all models a Bernoulli distribution was selected with a logit link function to model the binary detection response (yes = 1, no = 0). The fixed effects of time and number of GBV were treated as monotonic predictors since they are ordered categories. An interaction term was included between time and number of GVB when this information was available. A random effect of sample was used for all models since measures were collected in triplicates. A weakly informative prior was used for the fixed effects to aid convergence (normal distribution with a mean of 0 and standard deviation of 1.5).

Model validation was performed using prior and posterior predictive checks, as well as effective samples sizes and r-hats (McElreath, 2020).

## Experiment 6

```{r exp6-wrangle, include=FALSE}
# read in data ----
df.6 <- read_excel(here::here("analysis/data/Exp6-time2.xlsx"), sheet = 1)
df.6 <- df.6 %>%
  clean_names() %>%
  select(-exp)

cols <- c("sample", "rep", "sample_id")
df.6[, cols] <- lapply(df.6[, cols], as.factor)

df.6 <- df.6 %>%
  mutate(time = ordered(time)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive))

df.6.2h <- read_excel(here::here("analysis/data/Exp6-Number-2h.xlsx"), sheet = 1) %>%
  mutate(time = "2")
df.6.24h <- read_excel(here::here("analysis/data/Exp6-Number-24h.xlsx"), sheet = 1) %>%
  mutate(time = "24")

df.6.d <- rbind(df.6.2h, df.6.24h) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.6.d[, cols] <- lapply(df.6.d[, cols], as.factor)

df.6.d <- df.6.d %>%
  mutate(time = ordered(time)) %>%
  mutate(number = ordered(number)) %>%
  mutate(positive = ifelse(is.na(positive), 0, positive)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive))
```

<details><summary>Data frames</summary>

```{r exp6-summary-tibbles, echo=FALSE}
print(df.6 %>% group_by(sample, time) %>% tally())
print(df.6.d %>% group_by(sample, number) %>% tally())
```

</details>

::: {.panel-tabset}

```{r exp6-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

priors <- prior(normal(0,1.5), class = b)

mod.1 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.1.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod.2 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.2.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive ~ mo(time) * mo(number) + (1|sample_id))

mod.3 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.3.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) * mo(number) + (1|sample_id))

mod.4 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.4.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

## Effect plots

```{r exp6-eff1, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_1 <- conditional_effects(mod.1)[[1]] %>% mutate(cutoff = "No")
c_eff_2 <- conditional_effects(mod.2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_1, c_eff_2)

p_exp6_eff1 <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title="Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_1.png"),
  plot = p_exp6_eff1,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff1
```

Figure 1: Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Error bars represent 95% credible intervals with points representing the mean estimate.

```{r exp6-eff2, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5, out.width='100%', dpi = 600}
c_eff_3 <- conditional_effects(mod.3, effect = "time:number")[[1]] %>% mutate(cutoff = "No")
c_eff_4 <- conditional_effects(mod.4, effect = "time:number")[[1]] %>% mutate(cutoff = "Yes") %>%  rename(positive = positive_cut)

c_eff <- rbind(c_eff_3, c_eff_4) %>%
  mutate(time = paste0("Time: ", time, " hrs")) %>%
  mutate(number = paste0(number, " GVB"))

c_eff$cutoff <- factor(c_eff$cutoff, levels = c("Yes", "No"))

p_exp6_eff2 <- c_eff %>% ggplot(aes(x = cutoff, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("darkgreen", "darkblue")) +
  labs(x = "Cq cut off (> 39)", y = "Probability of detection (%)") +
  guides(color=guide_legend(title="Cq cut off (> 39)")) +
  facet_grid(number ~ time)

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_2.png"),
  plot = p_exp6_eff2,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff2
```

Figure 2: Conditional effect plot of GBV exposure on probability of detection faceted by time and number of GVB with (yes = green) and without (no = blue) a Cq cut off of 39. Error bars represent 95% credible intervals with points representing the mean estimate.

## Models

```{r exp6-models-show, ref.label="exp6-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp6-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod.1, ndraws = 200, col = 2)
pp_check(mod.2, ndraws = 200, col = 2)
pp_check(mod.3, ndraws = 200, col = 2)
pp_check(mod.4, ndraws = 200, col = 2)
```

## Output

```{r exp6-summary, echo=FALSE}
summary(mod.1)
summary(mod.2)
summary(mod.3)
summary(mod.4)
```

## Tables

```{r exp6-estimates, echo=FALSE}
# Modeled estimates from conditional_effects() on the probability scale.
# Convert to percent for reporting.

ce6_time <- dplyr::bind_rows(
  conditional_effects(mod.1)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.2)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

ce6_time_number <- dplyr::bind_rows(
  conditional_effects(mod.3, effect = "time:number")[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.4, effect = "time:number")[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`, `Number (GVB)`)

knitr::kable(
  ce6_time,
  caption = "Modeled estimates from conditional effects (Experiment 6): time",
  format = "html"
)

knitr::kable(
  ce6_time_number,
  caption = "Modeled estimates from conditional effects (Experiment 6): time x number",
  format = "html"
)
```

<!-- Pairwise comparisons removed -->

```{r exp6-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.1, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  },
  {
    emm <- emmeans::emmeans(mod.2, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)

pairwise_number <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.3, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  },
  {
    emm <- emmeans::emmeans(mod.4, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  }
)

knitr::kable(
  pairwise_number,
  caption = "Pairwise comparisons between number levels within each time (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

## Experiment 10

```{r exp10-wrangle, include=FALSE}
df.10 <- read_excel(here::here("analysis/data/Exp10- no cut off.xlsx"), sheet = 1) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.10[, cols] <- lapply(df.10[, cols], as.factor)

df.10 <- df.10 %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive)) %>%
  mutate(time = case_when(
    time == "16hrs" ~ 16,
    time == "1d" ~ 24,
    time == "2d" ~ 48,
    time == "3d" ~ 72,
    time == "7d" ~ 168,
    time == "0" ~ 0
  )) %>%
  mutate(time = ordered(time))
```

<details><summary>Data frames</summary>

```{r exp10-summary-tibbles, echo=FALSE}
print(df.10 %>% group_by(sample, time) %>% tally())
```

</details>

::: {.panel-tabset}

```{r exp10-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

priors <- prior(normal(0,1.5), class = b)

mod.5 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.5.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod.6 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.6.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

## Effect plots

```{r exp10-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_5 <- conditional_effects(mod.5)[[1]] %>% mutate(cutoff = "No")
c_eff_6 <- conditional_effects(mod.6)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_5, c_eff_6)

p_exp10_eff <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color=guide_legend(title="Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp10_effect_1.png"),
  plot = p_exp10_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_eff
```

Figure 3: Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Error bars represent 95% credible intervals with points representing the mean estimate.

## Models

```{r exp10-models-show, ref.label="exp10-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp10-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod.5, ndraws = 200, col = 2)
pp_check(mod.6, ndraws = 200, col = 2)
```

## Output

```{r exp10-summary, echo=FALSE}
summary(mod.5)
summary(mod.6)
```

## Tables

```{r exp10-estimates, echo=FALSE}
# Modeled estimates from conditional effects() on the probability scale.

ce10_time <- dplyr::bind_rows(
  conditional_effects(mod.5)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.6)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

knitr::kable(
  ce10_time,
  caption = "Modeled estimates from conditional effects (Experiment 10): time",
  format = "html"
)
```

<!-- Pairwise comparisons removed -->

```{r exp10-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.5, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  },
  {
    emm <- emmeans::emmeans(mod.6, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 10; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

## QPCR results 2025

<details><summary>Wrangling & Summary</summary>
```{r}
#read in data ----
df <- read_excel(here::here("analysis/data/qPCR_results251017_forTom.xlsx"), sheet = 1) # Import .xlsx file

df <- df %>% clean_names() %>% select(-location_in_orchard, -note)

cols <- c("sample", "method", "method_name")
df[, cols] <- lapply(df[, cols], as.factor) # Coerce variables to factors

cols <- c("col", "row", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml", "cq1", "cq2", "cq3")
df[, cols] <- lapply(df[, cols], as.numeric) # Coerce variables to factors

df <- reshape2::melt(
  df,
  id = c("sample", "col", "row", "method", "method_name", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml")
) # Make it long

df <- df |>
  rename(replicate = variable, cq = value) |>
  filter(!is.na(row)) |>
  mutate(week_ord = ordered(week)) |> # ordered factor
  mutate(method = as.factor(make_clean_names(method, allow_dupes = TRUE))) |>
  mutate(method_name = as.factor(make_clean_names(method_name, allow_dupes = TRUE))) |>
  mutate(cq = if_else(is.nan(cq), 0, cq)) |>
  mutate(pa = ifelse(cq > 0, 1, cq)) |> # Create binary variable
  filter(method %in% c("b", "c", "d")) |>
  mutate(sub_method = paste0(method, sep = "_", method_name)) |>
  droplevels()

str(df)
```
</details>

# Methods:

To compare the probability of detection (%) of GVB at different exposure levels (0, 2, 6, 16, 24, 48 hrs), with and without a Cq cut off (i.e., if a Cq value was greater than 39 then classify as no detection), and GVB density (where data was available: at 2hr and 24hrs), generalized linear mixed effect (GLMM) models were performed in R (R Core Team 2022) using the brms package (Bürkner, 2017).

For all models a Bernoulli distribution was selected with a logit link function to model the binary detection response (yes = 1, no = 0). The fixed effects of time and number of GBV were treated as monotonic predictors since they are ordered categories. An interaction term was included between time and number of GVB when this information was available. A random effect of sample was used for all models since measures were collected in triplicates. A weakly informative prior was used for the fixed effects to aid convergence (normal distribution with a mean of 0 and standard deviation of 1.5).

Model validation was performed using prior and posterior predictive checks, as well as effective samples sizes and r-hats (McElreath, 2020).

## qPCR Experiment

::: {.panel-tabset}

```{r qpcr-models-fit, include=FALSE}
mod_formula <- bf(pa ~ method + mo(week_ord, id = "method") + gp(row, col) + s(increase_in_pvh_number) + s(volume_in_container_ml) + (1|replicate))

priors <- prior(normal(0, 1.5), class = b)

mod_1 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.7.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(pa ~ method + s(increase_in_pvh_number, id = "method") + (1|replicate))

mod_2 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.8.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, id = "sub_method") + (1|replicate))

mod_3 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.9.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)
```

## Effect plots

### Model 1 (all effects)

```{r qpcr-eff1, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
ce_qpcr_1 <- conditional_effects(mod_1, effects = "week_ord:method")
p_qpcr_1 <- plot(ce_qpcr_1)[[1]]

ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_1.png"),
  plot = p_qpcr_1,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_qpcr_1
```

Figure 1: Conditional effect plot of GBV exposure on the probability of detection. Error bars represent 95% credible intervals with points representing the mean estimate.

### Model 2 (increase in pvh and method only)

```{r qpcr-eff2, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
ce_qpcr_2 <- conditional_effects(mod_2, effects = "increase_in_pvh_number:method")
p_qpcr_2 <- plot(ce_qpcr_2)[[1]]

ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_2.png"),
  plot = p_qpcr_2,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_qpcr_2
```

Figure 2: Conditional effect plot of GBV exposure on the probability of detection. Error bars represent 95% credible intervals with points representing the mean estimate.

### Model 3 (increase in pvh and sub method only)

```{r qpcr-eff3, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
ce_qpcr_3 <- conditional_effects(mod_3, effects = "increase_in_pvh_number:sub_method")
p_qpcr_3 <- plot(ce_qpcr_3)[[1]]

ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_3.png"),
  plot = p_qpcr_3,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_qpcr_3
```

## Models

```{r qpcr-models-show, ref.label="qpcr-models-fit", echo=TRUE, eval=FALSE}

```

## Model Comparison

```{r qpcr-waic, echo=FALSE}
waic_tbl_one <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

waic_tbl <- dplyr::bind_rows(
  waic_tbl_one(mod_1, "mod_1"),
  waic_tbl_one(mod_2, "mod_2"),
  waic_tbl_one(mod_3, "mod_3")
) %>%
  dplyr::mutate(
    delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)
  ) %>%
  dplyr::arrange(WAIC) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  )

knitr::kable(
  waic_tbl,
  caption = "Model comparison (qPCR Experiment): WAIC (lower is better)",
  format = "html"
)
```

## Validation

```{r qpcr-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_1, ndraws = 200, col = 2)
pp_check(mod_2, ndraws = 200, col = 2)
pp_check(mod_3, ndraws = 200, col = 2)
```

## Output

```{r qpcr-summary, echo=FALSE}
summary(mod_1)
summary(mod_2)
summary(mod_3)
```


:::

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Session Information</summary>
```{r}
utils::sessionInfo()
```
</details>

