---
title: "GBV Detection Analysis"
subtitle: "*Document structure and formatting were assisted with Cursor.*"
author: "Dr Tom Moore - Statistical Scientist"
date: "`r Sys.Date()`"
format:
  html:
    include-in-header:
      text: |
        <style>
        .quarto-title .subtitle,
        p.subtitle,
        .subtitle.lead {
          font-size: 0.88rem;
          font-weight: normal;
        }
        </style>
    df-print: paged
    code-fold: false
    toc: true
    toc-float: true
    toc-depth: 5
    theme: united
    highlight-style: tango
    fig-width: 9
    fig-height: 6
    fig-dpi: 600
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup-packages, message=FALSE, warning=FALSE}
# Install packages and functions (based on existing Rmd setups)
ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  invisible(sapply(pkg, require, character.only = TRUE))
}

packages <- c(
  "tidyverse", "readxl", "janitor", "here", "reshape2", "brms", "cmdstanr", "bayesplot", "scales", "knitr"
)

ipak(packages)

bayesplot::color_scheme_set("blue")
options(mc.cores = parallel::detectCores())

# Ensure plot output directory exists
dir.create(here::here("outputs/plots"), recursive = TRUE, showWarnings = FALSE)

# Production Sever Settings
cmdstanr::set_cmdstan_path("/powerplant/workspace/hrltxm/workbench-k8s/stan/vanilla_stan/cmdstan-2.36.0")
```

# Methods:

To compare the probability of detection (%) of GVB at different exposure levels (0, 2, 6, 16, 24, 48 hrs), with and without a Cq cut off (i.e., if a Cq value was greater than 39 then classify as no detection), and GVB density (where data was available: at 2hr and 24hrs), generalized linear mixed effect (GLMM) models were performed in R (R Core Team 2022) using the brms package (Bürkner, 2017).

For all models a Bernoulli distribution was selected with a logit link function to model the binary detection response (yes = 1, no = 0). The fixed effects of time and number of GBV were treated as monotonic predictors since they are ordered categories. An interaction term was included between time and number of GVB when this information was available. A random effect of sample was used for all models since measures were collected in triplicates. A weakly informative prior was used for the fixed effects to aid convergence (normal distribution with a mean of 0 and standard deviation of 1.5).

Model validation was performed using prior and posterior predictive checks, as well as effective samples sizes and r-hats (McElreath, 2020).

## Experiment 6

```{r exp6-wrangle, include=FALSE}
# read in data ----
df.6 <- read_excel(here::here("analysis/data/Exp6-time2.xlsx"), sheet = 1)
df.6 <- df.6 %>%
  clean_names() %>%
  select(-exp)

cols <- c("sample", "rep", "sample_id")
df.6[, cols] <- lapply(df.6[, cols], as.factor)

df.6 <- df.6 %>%
  mutate(time = ordered(time)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive)) %>%
  mutate(time_cat = factor(as.character(time), levels = levels(time), ordered = FALSE))

df.6.2h <- read_excel(here::here("analysis/data/Exp6-Number-2h.xlsx"), sheet = 1) %>%
  mutate(time = "2")
df.6.24h <- read_excel(here::here("analysis/data/Exp6-Number-24h.xlsx"), sheet = 1) %>%
  mutate(time = "24")

df.6.d <- rbind(df.6.2h, df.6.24h) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.6.d[, cols] <- lapply(df.6.d[, cols], as.factor)

df.6.d <- df.6.d %>%
  mutate(time = ordered(time)) %>%
  mutate(number = ordered(number)) %>%
  mutate(positive = ifelse(is.na(positive), 0, positive)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive)) %>%
  mutate(time_cat = factor(as.character(time), levels = levels(time), ordered = FALSE))

# GVB reads workbooks (Exp 6 / Exp 10) — map to shared column names; same schema as df.6 time-course extras
wrangle_exp6_reads_sheet <- function(rel_path, require_quantity_gbv = TRUE) {
  path <- here::here(rel_path)
  d <- read_excel(path, sheet = 1) %>% clean_names()
  d <- dplyr::select(d, -dplyr::any_of("exp"))

  if (!"time" %in% names(d) && "exposure_time_hours" %in% names(d)) {
    d <- dplyr::rename(d, time = exposure_time_hours)
  }
  if (!"positive" %in% names(d) && "q_pcr_pos" %in% names(d)) {
    d <- dplyr::rename(d, positive = q_pcr_pos)
  }
  if (!"positive" %in% names(d) && "qpcr_pos" %in% names(d)) {
    d <- dplyr::rename(d, positive = qpcr_pos)
  }

  if (!"quantity_gbv" %in% names(d)) {
    q_alias <- intersect(
      c("quantity_gvb", "quanity_gvb", "quanity_gbv", "quantity_gbv"),
      names(d)
    )
    if (length(q_alias)) {
      d <- dplyr::rename(d, quantity_gbv = dplyr::all_of(q_alias[[1]]))
    }
  }
  if (!"quantity_gbv" %in% names(d)) {
    nms <- names(d)
    iq <- nms[grepl("gbv", nms, ignore.case = TRUE)]
    iq <- iq[grepl("quant|quani", iq, ignore.case = TRUE)]
    if (length(iq) == 1L) {
      d <- dplyr::rename(d, quantity_gbv = dplyr::all_of(iq))
    }
  }
  if (require_quantity_gbv && !"quantity_gbv" %in% names(d)) {
    stop(
      "Reads sheet: could not map any column to `quantity_gbv` (expected names like `quantity_gvb` or `quanity_gvb` after clean_names). Found: ",
      paste(names(d), collapse = ", ")
    )
  }

  miss <- setdiff(c("time", "cq", "positive", "sample"), names(d))
  if (length(miss)) {
    stop(
      "After clean_names() and renames, `", rel_path, "` must include: ",
      paste(miss, collapse = ", "),
      ". Found: ",
      paste(names(d), collapse = ", ")
    )
  }

  d$cq <- suppressWarnings(as.numeric(d$cq))
  pv <- d$positive
  if (is.numeric(pv)) {
    d$positive <- ifelse(is.na(pv), NA_real_, as.integer(round(pv) != 0L))
  } else {
    ps <- toupper(stringr::str_trim(as.character(pv)))
    d$positive <- dplyr::case_when(
      is.na(ps) | ps == "" ~ NA_real_,
      ps %in% c("1", "Y", "YES", "TRUE", "T", "POS", "POSITIVE") ~ 1,
      TRUE ~ 0
    )
  }

  fac_cols <- intersect(c("sample", "rep", "sample_id"), names(d))
  for (cl in fac_cols) {
    d[[cl]] <- as.factor(d[[cl]])
  }

  tnum <- suppressWarnings(as.numeric(d$time))
  if (!all(is.na(tnum))) {
    tlv <- sort(unique(tnum[!is.na(tnum)]))
    d$time <- ordered(tnum, levels = tlv)
  } else {
    d$time <- ordered(as.character(d$time))
  }

  d %>%
    dplyr::mutate(
      positive_cut = dplyr::if_else(!is.na(.data$cq) & .data$cq > 39, 0L, as.integer(.data$positive)),
      time_cat = factor(as.character(.data$time), levels = levels(.data$time), ordered = FALSE)
    )
}

df.6.reads <- wrangle_exp6_reads_sheet("analysis/data/GVB_exp6_reads_tom.xlsx")
df.6.reads.no0 <- wrangle_exp6_reads_sheet("analysis/data/GVB_exp6_reads_tom_no_time0.xlsx")
```

<details><summary>Data frames</summary>

```{r exp6-summary-tibbles, echo=FALSE}
cat("Original time-course (Exp6-time2.xlsx)\n\n")
print(df.6 %>% group_by(sample, time) %>% tally())
str(df.6)

cat("\n\nOriginal density design (2 h / 24 h number sheets)\n\n")
print(df.6.d %>% group_by(sample, number) %>% tally())
str(df.6.d)

cat("\n\nReads — full (GVB_exp6_reads_tom.xlsx)\n\n")
print(df.6.reads %>% group_by(sample, time) %>% tally())
str(df.6.reads)

cat("\n\nReads — no time 0 (GVB_exp6_reads_tom_no_time0.xlsx)\n\n")
print(df.6.reads.no0 %>% group_by(sample, time) %>% tally())
str(df.6.reads.no0)
```

</details>

::: {.panel-tabset}

```{r exp6-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

priors <- prior(normal(0,1.5), class = b)

mod.1 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.1.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod.2 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.2.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive ~ mo(time) * mo(number) + (1|sample_id))

mod.3 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.3.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) * mo(number) + (1|sample_id))

mod.4 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.4.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

# Factor time (same mean structure; time as categorical, not monotonic)
mod_formula <- bf(positive ~ time_cat + (1|sample_id))

mod.1b <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.1b.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ time_cat + (1|sample_id))

mod.2b <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6,
  file = paste0(here::here("outputs/models/mod.2b.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive ~ time_cat * mo(number) + (1|sample_id))

mod.3b <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.3b.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ time_cat * mo(number) + (1|sample_id))

mod.4b <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.6.d,
  file = paste0(here::here("outputs/models/mod.4b.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp6-waic-compute, include=FALSE}
waic_tbl_one_exp <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_exp(mod.1, "mod.1 (time monotonic, no cut)"),
  waic_tbl_one_exp(mod.1b, "mod.1b (time factor, no cut)"),
  waic_tbl_one_exp(mod.2, "mod.2 (time monotonic, cut)"),
  waic_tbl_one_exp(mod.2b, "mod.2b (time factor, cut)"),
  waic_tbl_one_exp(mod.3, "mod.3 (time monot. × number, no cut)"),
  waic_tbl_one_exp(mod.3b, "mod.3b (time factor × number, no cut)"),
  waic_tbl_one_exp(mod.4, "mod.4 (time monot. × number, cut)"),
  waic_tbl_one_exp(mod.4b, "mod.4b (time factor × number, cut)")
) %>%
  dplyr::arrange(WAIC, Model) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

## Effect plots (monotonic time)

```{r exp6-eff1, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_1 <- conditional_effects(mod.1)[[1]] %>% mutate(cutoff = "No")
c_eff_2 <- conditional_effects(mod.2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_1, c_eff_2)

raw_exp6_eff1 <- dplyr::bind_rows(
  dplyr::mutate(df.6, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.6, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
)

p_exp6_eff1 <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff1,
    aes(x = time, y = y, group = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.04, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title="Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_1.png"),
  plot = p_exp6_eff1,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff1
```

Figure 1: Experiment 6, **monotonic time** (`mo(time)`). Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Black jittered points are raw binary outcomes by exposure time; error bars are 95% credible intervals and coloured points are posterior means.

```{r exp6-eff2, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5, out.width='100%', dpi = 600}
c_eff_3 <- conditional_effects(mod.3, effect = "time:number")[[1]] %>% mutate(cutoff = "No")
c_eff_4 <- conditional_effects(mod.4, effect = "time:number")[[1]] %>% mutate(cutoff = "Yes") %>%  rename(positive = positive_cut)

c_eff <- rbind(c_eff_3, c_eff_4) %>%
  mutate(time = paste0("Time: ", time, " hrs")) %>%
  mutate(number = paste0(number, " GVB"))

c_eff$cutoff <- factor(c_eff$cutoff, levels = c("Yes", "No"))

raw_exp6_eff2 <- df.6.d %>%
  dplyr::mutate(
    time = paste0("Time: ", .data$time, " hrs"),
    number = paste0(.data$number, " GVB")
  ) %>%
  tidyr::pivot_longer(
    cols = c(positive, positive_cut),
    names_to = "which_y",
    values_to = "y"
  ) %>%
  dplyr::mutate(
    cutoff = factor(
      dplyr::if_else(.data$which_y == "positive", "No", "Yes"),
      levels = c("Yes", "No")
    )
  )

p_exp6_eff2 <- c_eff %>% ggplot(aes(x = cutoff, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff2,
    aes(x = cutoff, y = y),
    inherit.aes = FALSE,
    position = position_jitter(width = 0.15, height = 0.04),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(values = c("darkgreen", "darkblue")) +
  labs(x = "Cq cut off (> 39)", y = "Probability of detection (%)") +
  guides(color=guide_legend(title="Cq cut off (> 39)")) +
  facet_grid(number ~ time)

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_2.png"),
  plot = p_exp6_eff2,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff2
```

Figure 2: Experiment 6, **monotonic time**. Conditional effect plot of GBV exposure on probability of detection faceted by time and number of GVB with (yes = green) and without (no = blue) a Cq cut off of 39. Black jittered points are raw outcomes; error bars are 95% credible intervals and coloured points are posterior means.

## Effect plots (factor time)

```{r exp6-eff1-fct, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
c_eff_1b <- conditional_effects(mod.1b)[[1]] %>% mutate(cutoff = "No")
c_eff_2b <- conditional_effects(mod.2b)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_1b, c_eff_2b)

raw_exp6_eff1b <- dplyr::bind_rows(
  dplyr::mutate(df.6, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.6, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
)

p_exp6_eff1b <- c_eff %>% ggplot(aes(x = time_cat, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff1b,
    aes(x = time_cat, y = y, group = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.04, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title = "Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_1_factor_time.png"),
  plot = p_exp6_eff1b,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff1b
```

Figure 1b: Experiment 6, **factor time** (`time_cat`). Same layout as Figure 1; black jittered points are raw outcomes.

```{r exp6-eff2-fct, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5, out.width='100%', dpi = 600}
c_eff_3b <- conditional_effects(mod.3b, effect = "time_cat:number")[[1]] %>% mutate(cutoff = "No")
c_eff_4b <- conditional_effects(mod.4b, effect = "time_cat:number")[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)

c_eff <- rbind(c_eff_3b, c_eff_4b) %>%
  mutate(time = paste0("Time: ", time_cat, " hrs")) %>%
  mutate(number = paste0(number, " GVB"))

c_eff$cutoff <- factor(c_eff$cutoff, levels = c("Yes", "No"))

raw_exp6_eff2b <- df.6.d %>%
  dplyr::mutate(
    time = paste0("Time: ", .data$time_cat, " hrs"),
    number = paste0(.data$number, " GVB")
  ) %>%
  tidyr::pivot_longer(
    cols = c(positive, positive_cut),
    names_to = "which_y",
    values_to = "y"
  ) %>%
  dplyr::mutate(
    cutoff = factor(
      dplyr::if_else(.data$which_y == "positive", "No", "Yes"),
      levels = c("Yes", "No")
    )
  )

p_exp6_eff2b <- c_eff %>% ggplot(aes(x = cutoff, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff2b,
    aes(x = cutoff, y = y),
    inherit.aes = FALSE,
    position = position_jitter(width = 0.15, height = 0.04),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(values = c("darkgreen", "darkblue")) +
  labs(x = "Cq cut off (> 39)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title = "Cq cut off (> 39)")) +
  facet_grid(number ~ time)

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_effect_2_factor_time.png"),
  plot = p_exp6_eff2b,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff2b
```

Figure 2b: Experiment 6, **factor time** (`time_cat × number`). Same layout as Figure 2.

## Model comparison

```{r exp6-waic-tab, echo=FALSE}
knitr::kable(
  exp6_waic_tbl,
  caption = "Experiment 6: WAIC (lower is better). Compare monotonic (`mo(time)`) vs factor (`time_cat`) formulations within each response and design.",
  format = "html"
)
```

## Models

```{r exp6-models-show, ref.label="exp6-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp6-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod.1, ndraws = 200, col = 2)
pp_check(mod.2, ndraws = 200, col = 2)
pp_check(mod.3, ndraws = 200, col = 2)
pp_check(mod.4, ndraws = 200, col = 2)
pp_check(mod.1b, ndraws = 200, col = 2)
pp_check(mod.2b, ndraws = 200, col = 2)
pp_check(mod.3b, ndraws = 200, col = 2)
pp_check(mod.4b, ndraws = 200, col = 2)
```

## Output

```{r exp6-summary, echo=FALSE}
summary(mod.1)
summary(mod.2)
summary(mod.3)
summary(mod.4)
summary(mod.1b)
summary(mod.2b)
summary(mod.3b)
summary(mod.4b)
```

## Tables

```{r exp6-estimates, echo=FALSE}
# Modeled estimates from conditional_effects() on the probability scale.
# Convert to percent for reporting.

ce6_time <- dplyr::bind_rows(
  conditional_effects(mod.1)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.2)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

ce6_time_number <- dplyr::bind_rows(
  conditional_effects(mod.3, effect = "time:number")[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.4, effect = "time:number")[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`, `Number (GVB)`)

knitr::kable(
  ce6_time,
  caption = "Modeled estimates from conditional effects (Experiment 6): time",
  format = "html"
)

knitr::kable(
  ce6_time_number,
  caption = "Modeled estimates from conditional effects (Experiment 6): time x number",
  format = "html"
)
```

<!-- Pairwise comparisons removed -->

```{r exp6-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.1, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  },
  {
    emm <- emmeans::emmeans(mod.2, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)

pairwise_number <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.3, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  },
  {
    emm <- emmeans::emmeans(mod.4, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  }
)

knitr::kable(
  pairwise_number,
  caption = "Pairwise comparisons between number levels within each time (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

## Experiment 6 — reads (GVB sequence counts)

Negative binomial regression on **`GVB_exp6_reads_tom.xlsx`** (`df.6.reads`): **`gvb_sequence_number ~ mo(time) + mo(quantity_gbv)`** (quantity GVB as an ordered factor for the monotonic term; default **brms** priors).

::: {.panel-tabset}

```{r exp6-reads-models-fit, include=FALSE}
df_exp6_reads_mod <- df.6.reads %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv)
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    .data$gvb_sequence_number >= 0L
  ) %>%
  dplyr::mutate(
    quantity_gbv = ordered(
      .data$quantity_gbv,
      levels = sort(unique(stats::na.omit(as.numeric(.data$quantity_gbv))))
    )
  )

mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv))

mod_exp6_reads <- brm(
  mod_formula,
  family = negbinomial(),
  data = df_exp6_reads_mod,
  file = paste0(here::here("outputs/models/mod_exp6_reads_gvbseq_negbin.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr",
  silent = 0,
  refresh = 1000,
  chains = 4,
  cores = 4,
  warmup = 2000,
  iter = 4000
)
```

## Effect plots

```{r exp6-reads-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width="100%", dpi=600}
#| fig-keep: all
raw <- df_exp6_reads_mod
exp6_reads_y_breaks <- c(0, 1, 10, 100, 1000, 10000, 100000)

exp6_reads_plot_theme <- ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
    panel.grid.minor.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
  )

ce_time <- conditional_effects(mod_exp6_reads, effects = "time")[[1]] %>%
  dplyr::arrange(.data$time)

p_exp6_reads_eff_time <- ggplot2::ggplot(
  ce_time,
  ggplot2::aes(x = .data$time, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
) +
  ggplot2::geom_jitter(
    data = raw,
    ggplot2::aes(x = .data$time, y = .data$gvb_sequence_number),
    inherit.aes = FALSE,
    position = ggplot2::position_jitter(width = 0.18, height = 0.35),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  ggplot2::geom_errorbar(width = 0.3, size = 1, color = "darkblue") +
  ggplot2::geom_point(size = 3, alpha = 0.5, color = "darkblue") +
  exp6_reads_plot_theme +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    breaks = exp6_reads_y_breaks,
    labels = scales::label_number(big.mark = ","),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

ce_qty <- conditional_effects(mod_exp6_reads, effects = "quantity_gbv")[[1]] %>%
  dplyr::arrange(suppressWarnings(as.numeric(as.character(.data$quantity_gbv))))

p_exp6_reads_eff_qty <- ggplot2::ggplot(
  ce_qty,
  ggplot2::aes(x = .data$quantity_gbv, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
) +
  ggplot2::geom_jitter(
    data = raw,
    ggplot2::aes(x = .data$quantity_gbv, y = .data$gvb_sequence_number),
    inherit.aes = FALSE,
    position = ggplot2::position_jitter(width = 0.12, height = 0.35),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  ggplot2::geom_errorbar(width = 0.3, size = 1, color = "darkblue") +
  ggplot2::geom_point(size = 3, alpha = 0.5, color = "darkblue") +
  exp6_reads_plot_theme +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    breaks = exp6_reads_y_breaks,
    labels = scales::label_number(big.mark = ","),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  ggplot2::labs(x = "Quantity GVB", y = "GVB sequence number")

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_reads_effect_time.png"),
  plot = p_exp6_reads_eff_time,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp6_reads_effect_quantity.png"),
  plot = p_exp6_reads_eff_qty,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_reads_eff_time
p_exp6_reads_eff_qty
```

Black jittered points are observed **GVB sequence numbers**; blue error bars and points are posterior mean and 95% credible intervals from **`conditional_effects()`** (**`mo(time)`** and **`mo(quantity_gbv)`**; other predictors at **brms** defaults). The vertical axis uses a **log(1 + count)** scale (natural log) so zeros remain visible; reference ticks **0, 1, 10, 100, 1,000, 10,000, 100,000**.

## Models

```{r exp6-reads-models-show, ref.label="exp6-reads-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp6-reads-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_exp6_reads)
```

## Output

```{r exp6-reads-summary, echo=FALSE}
summary(mod_exp6_reads)
```

## Tables

```{r exp6-reads-estimates, echo=FALSE}
ce_exp6_t <- conditional_effects(mod_exp6_reads, effects = "time")[[1]] %>%
  dplyr::transmute(
    `Time (hrs)` = time,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

ce_exp6_q <- conditional_effects(mod_exp6_reads, effects = "quantity_gbv")[[1]] %>%
  dplyr::transmute(
    `Quantity GVB` = quantity_gbv,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

knitr::kable(
  ce_exp6_t,
  caption = "Experiment 6 reads: conditional effects of time on the negative binomial mean (GVB sequence number).",
  format = "html"
)

knitr::kable(
  ce_exp6_q,
  caption = "Experiment 6 reads: conditional effects of quantity GVB on the negative binomial mean (GVB sequence number).",
  format = "html"
)
```

:::

## Experiment 10

```{r exp10-wrangle, include=FALSE}
df.10 <- read_excel(here::here("analysis/data/Exp10- no cut off.xlsx"), sheet = 1) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.10[, cols] <- lapply(df.10[, cols], as.factor)

df.10 <- df.10 %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive)) %>%
  mutate(time = case_when(
    time == "16hrs" ~ 16,
    time == "1d" ~ 24,
    time == "2d" ~ 48,
    time == "3d" ~ 72,
    time == "7d" ~ 168,
    time == "0" ~ 0
  )) %>%
  mutate(time = ordered(time)) %>%
  mutate(time_cat = factor(as.character(time), levels = levels(time), ordered = FALSE))

df.10.reads <- wrangle_exp6_reads_sheet("analysis/data/GVB_exp10_reads_tom.xlsx", require_quantity_gbv = FALSE)
```

<details><summary>Data frames</summary>

```{r exp10-summary-tibbles, echo=FALSE}
print(df.10 %>% group_by(sample, time) %>% tally())
str(df.10)

cat("\n\nReads — GVB_exp10_reads_tom.xlsx\n\n")
print(df.10.reads %>% group_by(sample, time) %>% tally())
str(df.10.reads)
```

</details>

::: {.panel-tabset}

```{r exp10-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

priors <- prior(normal(0,1.5), class = b)

mod.5 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.5.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod.6 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.6.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive ~ time_cat + (1|sample_id))

mod.5b <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.5b.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ time_cat + (1|sample_id))

mod.6b <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  prior = priors,
  data = df.10,
  file = paste0(here::here("outputs/models/mod.6b.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp10-waic-compute, include=FALSE}
waic_tbl_one_exp10 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp10_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_exp10(mod.5, "mod.5 (time monotonic, no cut)"),
  waic_tbl_one_exp10(mod.5b, "mod.5b (time factor, no cut)"),
  waic_tbl_one_exp10(mod.6, "mod.6 (time monotonic, cut)"),
  waic_tbl_one_exp10(mod.6b, "mod.6b (time factor, cut)")
) %>%
  dplyr::arrange(WAIC, Model) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

## Effect plots (monotonic time)

```{r exp10-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_5 <- conditional_effects(mod.5)[[1]] %>% mutate(cutoff = "No")
c_eff_6 <- conditional_effects(mod.6)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_5, c_eff_6)

raw_exp10_eff <- dplyr::bind_rows(
  dplyr::mutate(df.10, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.10, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
)

p_exp10_eff <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp10_eff,
    aes(x = time, y = y, group = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.04, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color=guide_legend(title="Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp10_effect_1.png"),
  plot = p_exp10_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_eff
```

Figure 3: Experiment 10, **monotonic time** (`mo(time)`). Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Black jittered points are raw binary outcomes by exposure time; error bars are 95% credible intervals and coloured points are posterior means.

## Effect plots (factor time)

```{r exp10-eff-fct, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
c_eff_5b <- conditional_effects(mod.5b)[[1]] %>% mutate(cutoff = "No")
c_eff_6b <- conditional_effects(mod.6b)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_5b, c_eff_6b)

raw_exp10_effb <- dplyr::bind_rows(
  dplyr::mutate(df.10, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.10, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
)

p_exp10_effb <- c_eff %>% ggplot(aes(x = time_cat, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp10_effb,
    aes(x = time_cat, y = y, group = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.04, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(0, 1),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title = "Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp10_effect_1_factor_time.png"),
  plot = p_exp10_effb,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_effb
```

Figure 3b: Experiment 10, **factor time** (`time_cat`). Same layout as Figure 3.

## Model comparison

```{r exp10-waic-tab, echo=FALSE}
knitr::kable(
  exp10_waic_tbl,
  caption = "Experiment 10: WAIC (lower is better). Compare monotonic (`mo(time)`) vs factor (`time_cat`) formulations.",
  format = "html"
)
```

## Models

```{r exp10-models-show, ref.label="exp10-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp10-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod.5, ndraws = 200, col = 2)
pp_check(mod.6, ndraws = 200, col = 2)
pp_check(mod.5b, ndraws = 200, col = 2)
pp_check(mod.6b, ndraws = 200, col = 2)
```

## Output

```{r exp10-summary, echo=FALSE}
summary(mod.5)
summary(mod.6)
summary(mod.5b)
summary(mod.6b)
```

## Tables

```{r exp10-estimates, echo=FALSE}
# Modeled estimates from conditional effects() on the probability scale.

ce10_time <- dplyr::bind_rows(
  conditional_effects(mod.5)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod.6)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

knitr::kable(
  ce10_time,
  caption = "Modeled estimates from conditional effects (Experiment 10): time",
  format = "html"
)
```

<!-- Pairwise comparisons removed -->

```{r exp10-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod.5, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  },
  {
    emm <- emmeans::emmeans(mod.6, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 10; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

## Experiment 10 — reads (GVB sequence counts)

Negative binomial regression on **`GVB_exp10_reads_tom.xlsx`** (`df.10.reads`): **`gvb_sequence_number ~ mo(time)`** (no quantity-GVB column in this workbook; default **brms** priors).

::: {.panel-tabset}

```{r exp10-reads-models-fit, include=FALSE}
df_exp10_reads_mod <- df.10.reads %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number))))
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$time),
    .data$gvb_sequence_number >= 0L
  )

if (length(levels(droplevels(df_exp10_reads_mod$time))) < 2L) {
  stop(
    "Experiment 10 reads: mo(time) needs at least two distinct time levels after filtering. Found: ",
    paste(levels(droplevels(df_exp10_reads_mod$time)), collapse = ", ")
  )
}

mod_formula <- bf(gvb_sequence_number ~ mo(time))

# file: brms adds .rds (avoid .Rmd here; threading off — reduces cmdstanr metadata failures on some drives)
mod_exp10_reads <- brm(
  mod_formula,
  family = negbinomial(),
  data = df_exp10_reads_mod,
  file = here::here("outputs/models/mod_exp10_reads_time_nb"),
  control = list(adapt_delta = 0.99, max_treedepth = 12L),
  backend = "cmdstanr",
  silent = 0,
  refresh = 500,
  chains = 4,
  cores = 4,
  warmup = 2000,
  iter = 4000
)
```

## Effect plots

```{r exp10-reads-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width="100%", dpi=600}
#| fig-keep: all
raw <- df_exp10_reads_mod
exp10_reads_y_breaks <- c(0, 1, 10, 100, 1000, 10000, 100000)

exp10_reads_plot_theme <- ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
    panel.grid.minor.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
  )

ce_time <- conditional_effects(mod_exp10_reads, effects = "time")[[1]] %>%
  dplyr::arrange(.data$time)

p_exp10_reads_eff_time <- ggplot2::ggplot(
  ce_time,
  ggplot2::aes(x = .data$time, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
) +
  ggplot2::geom_jitter(
    data = raw,
    ggplot2::aes(x = .data$time, y = .data$gvb_sequence_number),
    inherit.aes = FALSE,
    position = ggplot2::position_jitter(width = 0.18, height = 0.35),
    alpha = 0.35,
    size = 0.85,
    color = "black",
    stroke = 0
  ) +
  ggplot2::geom_errorbar(width = 0.3, size = 1, color = "darkblue") +
  ggplot2::geom_point(size = 3, alpha = 0.5, color = "darkblue") +
  exp10_reads_plot_theme +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    breaks = exp10_reads_y_breaks,
    labels = scales::label_number(big.mark = ","),
    expand = ggplot2::expansion(mult = c(0.02, 0.02))
  ) +
  ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

ggplot2::ggsave(
  filename = here::here("outputs/plots", "exp10_reads_effect_time.png"),
  plot = p_exp10_reads_eff_time,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_reads_eff_time
```

Black jittered points are observed **GVB sequence numbers**; blue error bars and points are posterior mean and 95% credible intervals from **`conditional_effects()`** for **`mo(time)`** only. The vertical axis uses a **log(1 + count)** scale with ticks at **0, 1, 10, 100, 1,000, 10,000, 100,000**.

## Models

```{r exp10-reads-models-show, ref.label="exp10-reads-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp10-reads-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_exp10_reads)
```

## Output

```{r exp10-reads-summary, echo=FALSE}
summary(mod_exp10_reads)
```

## Tables

```{r exp10-reads-estimates, echo=FALSE}
ce_exp10_t <- conditional_effects(mod_exp10_reads, effects = "time")[[1]] %>%
  dplyr::transmute(
    `Time (hrs)` = time,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

knitr::kable(
  ce_exp10_t,
  caption = "Experiment 10 reads: conditional effects of time on the negative binomial mean (GVB sequence number).",
  format = "html"
)
```

:::

## QPCR experiment

<details><summary>Data frames</summary>
```{r}
#read in data ----
df <- read_excel(here::here("analysis/data/qPCR_results251017_forTom.xlsx"), sheet = 1) # Import .xlsx file

df <- df %>% clean_names() %>% select(-location_in_orchard, -note)

cols <- c("sample", "method", "method_name")
df[, cols] <- lapply(df[, cols], as.factor) # Coerce variables to factors

cols <- c("col", "row", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml", "cq1", "cq2", "cq3")
df[, cols] <- lapply(df[, cols], as.numeric) # Coerce variables to factors

df <- reshape2::melt(
  df,
  id = c("sample", "col", "row", "method", "method_name", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml")
) # Make it long

df <- df |>
  rename(replicate = variable, cq = value) |>
  filter(!is.na(row)) |>
  mutate(week_ord = ordered(week)) |> # ordered factor
  mutate(method = as.factor(make_clean_names(method, allow_dupes = TRUE))) |>
  mutate(method_name = as.factor(make_clean_names(method_name, allow_dupes = TRUE))) |>
  mutate(cq = if_else(is.nan(cq), 0, cq)) |>
  mutate(pa = ifelse(cq > 0, 1, cq)) |> # Create binary variable
  filter(method %in% c("b", "c", "d")) |>
  mutate(sub_method = paste0(method, sep = "_", method_name)) |>
  droplevels()

str(df)
```
</details>

::: {.panel-tabset}

```{r qpcr-models-fit-submethod, include=FALSE}
priors <- prior(normal(0, 1.5), class = b)

mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, id = "sub_method") + (1|replicate))

mod_3 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.9.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: linear (separate slopes per sub-method)
mod_formula <- bf(pa ~ sub_method * increase_in_pvh_number + (1|replicate))

mod_4 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.10.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: one shared smooth; sub-method shifts intercept only
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number) + (1|replicate))

mod_5 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.11.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: separate smooths per sub-method (mgcv-style by)
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, by = sub_method) + (1|replicate))

mod_6 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.12.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# + monotonic week within sub_method (parallel to mod_3–mod_6)
mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, id = "sub_method") + (1|replicate)
)

mod_7 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.14.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method * increase_in_pvh_number + mo(week_ord, id = "sub_method") + (1|replicate)
)

mod_8 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.15.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number) + (1|replicate)
)

mod_9 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.16.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, by = sub_method) + (1|replicate)
)

mod_10 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.17.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × monotonic week + separate PVH smooths per sub-method
mod_formula <- bf(
  pa ~ sub_method * mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, by = sub_method) + (1|replicate)
)

mod_11 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  prior = priors,
  data = df,
  file = paste0(here::here("outputs/models/mod.18.Rmd")),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)
```

```{r qpcr-eff-helpers, include=FALSE}
qpcr_sub_method_label <- function(x) {
  lbl <- vapply(as.character(x), function(s) {
    p <- strsplit(s, "_", fixed = TRUE)[[1]]
    if (length(p) == 0L) {
      return(s)
    }
    letter_lab <- dplyr::case_match(
      p[[1]],
      "b" ~ "B",
      "c" ~ "C",
      "d" ~ "D",
      .default = stringr::str_to_title(p[[1]])
    )
    if (length(p) == 1L) {
      return(paste0("Method ", letter_lab))
    }
    rest <- stringr::str_replace_all(paste(p[-1L], collapse = "_"), "_", " ")
    rest <- stringr::str_to_title(rest)
    paste0("Method ", letter_lab, " — ", rest)
  }, character(1))
  factor(lbl, levels = sort(unique(lbl)))
}

# Shared qPCR figure: PVH × sub_method conditional effects (ghost + focal facets)
qpcr_plot_pvh_by_submethod <- function(fit) {
  c_eff <- conditional_effects(fit, effects = "increase_in_pvh_number:sub_method")[[1]] %>%
    dplyr::arrange(sub_method, increase_in_pvh_number) %>%
    dplyr::mutate(sub_method_l = qpcr_sub_method_label(sub_method))
  panel_levels <- levels(c_eff$sub_method_l)
  ghost <- purrr::map_dfr(panel_levels, function(f) {
    c_eff %>%
      dplyr::filter(as.character(.data$sub_method_l) != as.character(f)) %>%
      dplyr::mutate(facet_panel = factor(f, levels = panel_levels))
  })
  focal <- c_eff %>%
    dplyr::mutate(facet_panel = factor(.data$sub_method_l, levels = panel_levels))
  obs_raw <- fit$data %>%
    dplyr::mutate(
      sub_method_l = qpcr_sub_method_label(.data$sub_method),
      facet_panel = factor(.data$sub_method_l, levels = panel_levels)
    )
  ggplot2::ggplot() +
    ggplot2::geom_jitter(
      data = obs_raw,
      ggplot2::aes(x = increase_in_pvh_number, y = pa),
      inherit.aes = FALSE,
      height = 0.04,
      width = 0.15,
      alpha = 0.35,
      size = 0.85,
      color = "black",
      stroke = 0
    ) +
    ggplot2::geom_ribbon(
      data = ghost,
      ggplot2::aes(
        x = increase_in_pvh_number, ymin = lower__, ymax = upper__,
        group = sub_method_l
      ),
      inherit.aes = FALSE,
      fill = "gray45",
      alpha = 0.04
    ) +
    ggplot2::geom_line(
      data = ghost,
      ggplot2::aes(x = increase_in_pvh_number, y = estimate__, group = sub_method_l),
      inherit.aes = FALSE,
      color = "gray40",
      linewidth = 0.5,
      alpha = 0.1
    ) +
    ggplot2::geom_ribbon(
      data = focal,
      ggplot2::aes(
        x = increase_in_pvh_number, ymin = lower__, ymax = upper__,
        fill = sub_method_l, group = sub_method_l
      ),
      inherit.aes = FALSE,
      alpha = 0.2,
      color = NA
    ) +
    ggplot2::geom_line(
      data = focal,
      ggplot2::aes(
        x = increase_in_pvh_number, y = estimate__,
        color = sub_method_l, group = sub_method_l
      ),
      inherit.aes = FALSE,
      linewidth = 1
    ) +
    ggplot2::facet_wrap(ggplot2::vars(facet_panel), ncol = 3, scales = "fixed") +
    ggplot2::theme_classic() +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 9),
      legend.position = "none"
    ) +
    ggplot2::scale_y_continuous(
      labels = scales::percent,
      limits = c(0, 1),
      expand = ggplot2::expansion(mult = c(0.02, 0.02))
    ) +
    ggplot2::scale_color_brewer(palette = "Dark2", type = "qual") +
    ggplot2::scale_fill_brewer(palette = "Dark2", type = "qual") +
    ggplot2::labs(x = "PVH number", y = "Probability of detection (%)")
}

# Week × sub_method: posterior mean response at each (week_ord, sub_method), holding
# increase_in_pvh_number at its within–sub-method median and replicate at a reference level.
# Uses posterior_epred() so it behaves for all PVH structures (linear, shared spline, `by`)
# — conditional_effects("week_ord:sub_method") is unreliable for e.g. mod_10 / mod_11.
qpcr_plot_week_by_submethod <- function(fit) {
  d <- fit$data

  wk <- d$week_ord
  if (is.factor(wk)) {
    wk_lvls <- levels(wk)
    wk_ordered <- is.ordered(wk)
  } else {
    wk_lvls <- unique(as.character(wk[!is.na(wk)]))
    wk_ordered <- TRUE
  }

  sm <- d$sub_method
  if (!is.factor(sm)) {
    sm <- factor(sm, levels = unique(as.character(sm[!is.na(sm)])))
  } else {
    sm <- droplevels(sm)
  }
  sm_lvls <- levels(sm)

  rp <- d$replicate
  if (!is.factor(rp)) {
    rp <- factor(rp, levels = unique(as.character(rp[!is.na(rp)])))
  } else {
    rp <- droplevels(rp)
  }
  rep_lvls <- levels(rp)
  rep_ref <- rep_lvls[[1L]]

  if (length(wk_lvls) < 1L || length(sm_lvls) < 1L) {
    stop("fit$data is missing week_ord or sub_method factor levels.")
  }
  med_pvh <- tapply(d$increase_in_pvh_number, sm, stats::median, na.rm = TRUE)
  med_pvh <- med_pvh[sm_lvls]
  if (any(is.na(med_pvh))) {
    stop("Could not compute median PVH for every sub_method level.")
  }
  rows <- tidyr::expand_grid(
    week_ord = factor(wk_lvls, levels = wk_lvls, ordered = wk_ordered),
    sub_method = factor(sm_lvls, levels = sm_lvls)
  )
  rows$increase_in_pvh_number <- as.numeric(med_pvh[as.character(rows$sub_method)])
  rows$replicate <- factor(rep_ref, levels = rep_lvls)

  nd <- tryCatch(
    min(1000L, as.integer(brms::nsamples(fit))),
    error = function(...) 1000L
  )
  if (!is.finite(nd) || nd < 1L) nd <- 1000L
  ep <- brms::posterior_epred(
    object = fit,
    newdata = rows,
    re_formula = NA,
    ndraws = nd
  )
  rq <- apply(ep, 2L, stats::quantile, probs = c(0.025, 0.5, 0.975), na.rm = TRUE)
  c_eff <- rows %>%
    dplyr::mutate(
      lower__ = as.numeric(rq[1, ]),
      estimate__ = as.numeric(rq[2, ]),
      upper__ = as.numeric(rq[3, ])
    ) %>%
    dplyr::filter(is.finite(estimate__)) %>%
    dplyr::mutate(sub_method_l = qpcr_sub_method_label(sub_method))
  if (nrow(c_eff) < 1L) {
    stop("posterior_epred returned no usable week × sub-method estimates.")
  }

  panel_levels <- levels(c_eff$sub_method_l)
  plot_dat <- c_eff %>%
    dplyr::mutate(facet_panel = factor(.data$sub_method_l, levels = panel_levels))
  obs_raw <- d %>%
    dplyr::mutate(
      week_ord = factor(as.character(.data$week_ord), levels = wk_lvls, ordered = wk_ordered),
      sub_method_l = qpcr_sub_method_label(.data$sub_method),
      facet_panel = factor(.data$sub_method_l, levels = panel_levels)
    )

  n_col <- length(unique(c_eff$sub_method_l))
  p <- ggplot2::ggplot(plot_dat) +
    ggplot2::geom_jitter(
      data = obs_raw,
      ggplot2::aes(x = week_ord, y = pa),
      inherit.aes = FALSE,
      height = 0.04,
      width = 0.18,
      alpha = 0.35,
      size = 0.85,
      color = "black",
      stroke = 0
    ) +
    ggplot2::geom_linerange(
      ggplot2::aes(
        x = week_ord,
        ymin = lower__, ymax = upper__,
        color = sub_method_l,
        group = week_ord
      ),
      linewidth = 1
    ) +
    ggplot2::geom_point(
      ggplot2::aes(
        x = week_ord, y = estimate__,
        color = sub_method_l,
        group = week_ord
      ),
      size = 3,
      alpha = 0.55
    ) +
    ggplot2::facet_wrap(ggplot2::vars(facet_panel), ncol = 3, scales = "fixed") +
    ggplot2::theme_classic() +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 9),
      legend.position = "none",
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.3),
      panel.grid.major.y = ggplot2::element_line(colour = "grey85", linewidth = 0.3)
    ) +
    ggplot2::scale_y_continuous(
      labels = scales::percent,
      limits = c(0, 1),
      expand = ggplot2::expansion(mult = c(0.02, 0.02))
    ) +
    ggplot2::labs(x = "Week", y = "Probability of detection (%)")
  if (n_col <= 8L) {
    p <- p + ggplot2::scale_color_brewer(palette = "Dark2", type = "qual")
  } else {
    p <- p + ggplot2::scale_color_hue(l = 55, c = 80)
  }
  p
}
```

```{r qpcr-waic-compute, include=FALSE}
waic_tbl_one <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

waic_tbl_pre <- dplyr::bind_rows(
  waic_tbl_one(mod_3, "mod_3"),
  waic_tbl_one(mod_4, "mod_4"),
  waic_tbl_one(mod_5, "mod_5"),
  waic_tbl_one(mod_6, "mod_6"),
  waic_tbl_one(mod_7, "mod_7"),
  waic_tbl_one(mod_8, "mod_8"),
  waic_tbl_one(mod_9, "mod_9"),
  waic_tbl_one(mod_10, "mod_10"),
  waic_tbl_one(mod_11, "mod_11")
) %>%
  dplyr::mutate(model_id = as.integer(sub("^mod_", "", Model))) %>%
  dplyr::arrange(WAIC, model_id)

waic_tbl <- waic_tbl_pre %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)

qpcr_model_formula_lines <- c(
  "3" = "bf(pa ~ sub_method + s(increase_in_pvh_number, id = \"sub_method\") + (1|replicate))",
  "4" = "bf(pa ~ sub_method * increase_in_pvh_number + (1|replicate))",
  "5" = "bf(pa ~ sub_method + s(increase_in_pvh_number) + (1|replicate))",
  "6" = "bf(pa ~ sub_method + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "7" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, id = \"sub_method\") + (1|replicate))",
  "8" = "bf(pa ~ sub_method * increase_in_pvh_number + mo(week_ord, id = \"sub_method\") + (1|replicate))",
  "9" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number) + (1|replicate))",
  "10" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "11" = "bf(pa ~ sub_method * mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, by = sub_method) + (1|replicate))"
)
```

## Effect plots

Side-by-side tabs for **`mod_8`** and **`mod_10`**. Each tab shows **week × sub-method** (`posterior_epred` on a full week × sub-method grid: median PVH per sub-method, reference replicate, `re_formula = NA`) and **PVH number × sub-method** (`conditional_effects`; other predictors at **brms** defaults). Figures are written to `outputs/plots/` as `qpcr_effect_mod8_*` and `qpcr_effect_mod10_*`.

::: {.panel-tabset}

## Effect plots (mod 8)

```{r qpcr-effect-mod8, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_8
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["8"]], "\n```\n\n", sep = "")
cat("**Week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod8_week.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n**PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod8_pvh.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
```

## Effect plots (mod 10)

```{r qpcr-effect-mod10, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_10
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["10"]], "\n```\n\n", sep = "")
cat("**Week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod10_week.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n**PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "qpcr_effect_mod10_pvh.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
```

:::

## Models

Sub-method formulations (`mod_3`–`mod_11`; includes monotonic week in `mod_7`–`mod_11`):

```{r qpcr-models-show, ref.label="qpcr-models-fit-submethod", echo=TRUE, eval=FALSE}

```

## Model Comparison

```{r qpcr-waic, echo=FALSE}
knitr::kable(
  waic_tbl,
  caption = "Model comparison (qPCR Experiment): WAIC (lower is better).",
  format = "html"
)
```

## Validation

```{r qpcr-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_3, ndraws = 200, col = 2)
pp_check(mod_4, ndraws = 200, col = 2)
pp_check(mod_5, ndraws = 200, col = 2)
pp_check(mod_6, ndraws = 200, col = 2)
pp_check(mod_7, ndraws = 200, col = 2)
pp_check(mod_8, ndraws = 200, col = 2)
pp_check(mod_9, ndraws = 200, col = 2)
pp_check(mod_10, ndraws = 200, col = 2)
pp_check(mod_11, ndraws = 200, col = 2)
```

## Output

```{r qpcr-summary, echo=FALSE}
summary(mod_3)
summary(mod_4)
summary(mod_5)
summary(mod_6)
summary(mod_7)
summary(mod_8)
summary(mod_9)
summary(mod_10)
summary(mod_11)
```


:::

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Session Information</summary>
```{r}
utils::sessionInfo()
```
</details>

