---
title: "GBV Detection Analysis"
subtitle: "*Document structure and formatting were assisted with Cursor.*"
author: "Dr Tom Moore - Statistical Scientist"
date: "`r Sys.Date()`"
format:
  html:
    include-in-header:
      text: |
        <style>
        .quarto-title .subtitle,
        p.subtitle,
        .subtitle.lead {
          font-size: 0.88rem;
          font-weight: normal;
        }
        /* Model summary blocks (brms `summary()`): smaller type, horizontal scroll, no soft wrap */
        :root {
          --model-summary-font-rem: 0.72;
        }
        .model-output-panel {
          overflow-x: auto;
          overflow-y: hidden;
          max-width: 100%;
          margin: 0.35rem 0 0.9rem;
          padding: 0.45rem 0.55rem;
          border: 1px solid rgba(0, 0, 0, 0.12);
          border-radius: 4px;
          background: rgba(0, 0, 0, 0.02);
          -webkit-overflow-scrolling: touch;
        }
        .model-output-panel pre,
        .model-output-panel .ansi,
        .model-output-panel code {
          font-size: calc(var(--model-summary-font-rem) * 1rem) !important;
          line-height: 1.28;
          white-space: pre !important;
          word-break: normal !important;
          overflow-wrap: normal !important;
          margin: 0;
        }
        #model-output-global-toolbar {
          position: fixed;
          right: max(12px, env(safe-area-inset-right));
          bottom: max(12px, env(safe-area-inset-bottom));
          z-index: 1080;
          display: none;
          align-items: center;
          gap: 0.45rem;
          padding: 0.35rem 0.65rem;
          font-size: 0.78rem;
          line-height: 1.2;
          background: var(--bs-body-bg, #fff);
          color: var(--bs-body-color, #333);
          border: 1px solid rgba(0, 0, 0, 0.15);
          border-radius: 6px;
          box-shadow: 0 2px 8px rgba(0, 0, 0, 0.08);
          max-width: calc(100vw - 24px);
        }
        #model-output-global-toolbar.visible {
          display: flex;
        }
        #model-output-global-toolbar label {
          margin: 0;
          white-space: nowrap;
        }
        #model-output-global-toolbar input[type="range"] {
          width: min(140px, 35vw);
        }
        </style>
        <script>
        (function () {
          var root = document.documentElement;
          var debounceTimer = null;
          function debounced(fn, ms) {
            return function () {
              clearTimeout(debounceTimer);
              debounceTimer = setTimeout(fn, ms);
            };
          }
          function ensureToolbar() {
            var panels = document.querySelectorAll(".model-output-panel");
            var tb = document.getElementById("model-output-global-toolbar");
            var slider = document.getElementById("summary-font-size-slider");
            if (!tb || !slider) return;
            if (panels.length) {
              tb.classList.add("visible");
              root.style.setProperty("--model-summary-font-rem", slider.value);
            } else {
              tb.classList.remove("visible");
            }
          }
          function initModelSummaryToolbar() {
            var tb = document.getElementById("model-output-global-toolbar");
            var slider;
            if (!tb) {
              tb = document.createElement("div");
              tb.id = "model-output-global-toolbar";
              tb.innerHTML =
                '<label for="summary-font-size-slider">Output font size</label>' +
                '<input type="range" id="summary-font-size-slider" min="0.55" max="1" step="0.03" value="0.72" ' +
                'aria-label="Adjust output font size" title="Font size for printed model output (brms summary)"/>';
              document.body.appendChild(tb);
              slider = tb.querySelector("#summary-font-size-slider");
              slider.addEventListener("input", function () {
                root.style.setProperty("--model-summary-font-rem", slider.value);
              });
            }
            ensureToolbar();
            var mo = new MutationObserver(debounced(ensureToolbar, 80));
            mo.observe(document.body, { childList: true, subtree: true });
            document.addEventListener("shown.bs.tab", ensureToolbar);
          }
          document.addEventListener("DOMContentLoaded", initModelSummaryToolbar);
        })();
        </script>
    df-print: paged
    code-fold: false
    toc: true
    toc-float: true
    toc-depth: 5
    theme: united
    highlight-style: tango
    fig-width: 9
    fig-height: 6
    fig-dpi: 600
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup-packages, message=FALSE, warning=FALSE}
# Install packages and functions (based on existing Rmd setups)
ipak <- function(pkg) {
  new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
  if (length(new.pkg)) {
    install.packages(new.pkg, dependencies = TRUE)
  }
  invisible(sapply(pkg, require, character.only = TRUE))
}

packages <- c(
  "tidyverse", "readxl", "janitor", "here", "reshape2", "brms", "cmdstanr", "bayesplot", "scales", "knitr", "patchwork"
)

ipak(packages)

# Compact structure only: one header + `str()`-style columns (`str()` on tbl_df already prints the tibble line)
brief_dataframe <- function(x, label = NULL) {
  if (!is.null(label) && nzchar(as.character(label)[1])) {
    cat(label, "\n\n", sep = "")
  }
  x_tb <- dplyr::as_tibble(x)
  ch <- utils::capture.output(utils::str(x_tb, give.attr = FALSE))
  hdr <- sprintf("tibble [%d × %d] (S3: tbl_df/tbl/data.frame)", nrow(x_tb), ncol(x_tb))
  drop_first <- length(ch) > 0L && grepl("obs\\.\\s+of\\s+\\d+\\s+variables", ch[1L])
  body <- if (drop_first) ch[-1L] else ch
  out <- if (drop_first) {
    c(hdr, body)
  } else if (length(body) && startsWith(body[[1L]], "tibble [")) {
    body
  } else {
    c(hdr, body)
  }
  cat(paste(out, collapse = "\n"), "\n\n", sep = "")
  invisible(x_tb)
}

# Raw binary outcomes (0/1): nudge below 0 / above 1 so jitter clears model intervals on [0,1]
binary_y_for_plot <- function(y, pad = 0.06) {
  dplyr::case_when(
    is.na(y) ~ NA_real_,
    as.numeric(y) < 0.5 ~ as.numeric(y) - pad,
    TRUE ~ as.numeric(y) + pad
  )
}

# Probability scale — include 1 so percent labels show 100%
prob_det_y_breaks <- c(0, 0.25, 0.5, 0.75, 1)

bayesplot::color_scheme_set("blue")
options(mc.cores = parallel::detectCores())

# Ensure plot output directory exists
dir.create(here::here("outputs/plots"), recursive = TRUE, showWarnings = FALSE)

# Production Sever Settings
cmdstanr::set_cmdstan_path("/powerplant/workspace/hrltxm/workbench-k8s/stan/vanilla_stan/cmdstan-2.36.0")
```

# Methods:

To compare the probability of detection (%) of GVB at different exposure levels (0, 2, 6, 16, 24, 48 hrs), with and without a Cq cut off (i.e., if a Cq value was greater than 39 then classify as no detection), and GVB density (where data was available: at 2hr and 24hrs), generalized linear mixed effect (GLMM) models were performed in R (R Core Team 2022) using the brms package (Bürkner, 2017).

For all models a Bernoulli distribution was selected with a logit link function to model the binary detection response (yes = 1, no = 0). The fixed effects of time and number of GBV were treated as monotonic predictors since they are ordered categories. Where both were available (density design at 2 h and 24 h), they were modeled additively as `mo(time) + mo(number)` (no interaction). A random effect of sample was used for all models since measures were collected in triplicates.

Model validation was performed using prior and posterior predictive checks, as well as effective samples sizes and r-hats (McElreath, 2020).

## Experiment 6

```{r exp6-wrangle, include=FALSE}
# read in data ----
df.6 <- read_excel(here::here("analysis/data/Exp6-time2.xlsx"), sheet = 1)
df.6 <- df.6 %>%
  clean_names() %>%
  select(-exp)

cols <- c("sample", "rep", "sample_id")
df.6[, cols] <- lapply(df.6[, cols], as.factor)

df.6 <- df.6 %>%
  mutate(time = ordered(time)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive))

df.6.2h <- read_excel(here::here("analysis/data/Exp6-Number-2h.xlsx"), sheet = 1) %>%
  mutate(time = "2")
df.6.24h <- read_excel(here::here("analysis/data/Exp6-Number-24h.xlsx"), sheet = 1) %>%
  mutate(time = "24")

df.6.d <- rbind(df.6.2h, df.6.24h) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.6.d[, cols] <- lapply(df.6.d[, cols], as.factor)

df.6.d <- df.6.d %>%
  mutate(time = ordered(time)) %>%
  mutate(number = ordered(number)) %>%
  mutate(positive = ifelse(is.na(positive), 0, positive)) %>%
  mutate(positive_cut = ifelse(cq > 39, 0, positive))

cond_ce6_time_number <- dplyr::distinct(df.6.d, .data$number) %>% dplyr::arrange(.data$number)

# GVB reads workbooks (Exp 6 / Exp 10) — map to shared column names; same schema as df.6 time-course extras
wrangle_exp6_reads_sheet <- function(
  rel_path,
  require_quantity_gbv = TRUE,
  require_total_reads = FALSE
) {
  path <- here::here(rel_path)
  d <- read_excel(path, sheet = 1) %>% clean_names()
  d <- dplyr::select(d, -dplyr::any_of("exp"))

  # If the sheet has "Exposure time (hours)" and also another `time` column (e.g. replicate index),
  # always take the exposure column as the experimental factor.
  if ("exposure_time_hours" %in% names(d)) {
    d <- dplyr::mutate(d, time = .data$exposure_time_hours) %>%
      dplyr::select(-dplyr::all_of("exposure_time_hours"))
  } else if (!"time" %in% names(d)) {
    et_alias <- intersect(
      c("exposure_time_h", "exposure_time", "exposure_h", "time_exposure"),
      names(d)
    )
    if (length(et_alias)) {
      d <- dplyr::rename(d, time = dplyr::all_of(et_alias[[1]]))
    }
  }
  if (!"positive" %in% names(d) && "q_pcr_pos" %in% names(d)) {
    d <- dplyr::rename(d, positive = q_pcr_pos)
  }
  if (!"positive" %in% names(d) && "qpcr_pos" %in% names(d)) {
    d <- dplyr::rename(d, positive = qpcr_pos)
  }

  if (!"quantity_gbv" %in% names(d)) {
    q_alias <- intersect(
      c("quantity_gvb", "quanity_gvb", "quanity_gbv", "quantity_gbv"),
      names(d)
    )
    if (length(q_alias)) {
      d <- dplyr::rename(d, quantity_gbv = dplyr::all_of(q_alias[[1]]))
    }
  }
  if (!"quantity_gbv" %in% names(d)) {
    nms <- names(d)
    iq <- nms[grepl("gbv", nms, ignore.case = TRUE)]
    iq <- iq[grepl("quant|quani", iq, ignore.case = TRUE)]
    if (length(iq) == 1L) {
      d <- dplyr::rename(d, quantity_gbv = dplyr::all_of(iq))
    }
  }
  if (require_quantity_gbv && !"quantity_gbv" %in% names(d)) {
    stop(
      "Reads sheet: could not map any column to `quantity_gbv` (expected names like `quantity_gvb` or `quanity_gbv` after clean_names). Found: ",
      paste(names(d), collapse = ", ")
    )
  }

  if (!"total_reads" %in% names(d)) {
    tr_alias <- intersect(
      c(
        "total_reads", "reads_total", "total_read", "n_reads", "n_total_reads",
        "sequencing_reads", "total_sequencing_reads", "library_size"
      ),
      names(d)
    )
    if (length(tr_alias)) {
      d <- dplyr::rename(d, total_reads = dplyr::all_of(tr_alias[[1]]))
    }
  }
  if (!"total_reads" %in% names(d)) {
    nms <- names(d)
    cand <- nms[grepl("read", nms, ignore.case = TRUE)]
    cand <- cand[grepl("total|^n_", cand, ignore.case = TRUE) | grepl("^reads_", cand, ignore.case = TRUE)]
    if (length(cand) == 1L) {
      d <- dplyr::rename(d, total_reads = dplyr::all_of(cand))
    }
  }
  if (require_total_reads && !"total_reads" %in% names(d)) {
    stop(
      "Reads sheet: could not map any column to `total_reads` (e.g. `total_reads`, `reads_total`, `n_reads` after clean_names). Found: ",
      paste(names(d), collapse = ", ")
    )
  }
  if ("total_reads" %in% names(d)) {
    d$total_reads <- suppressWarnings(as.numeric(d$total_reads))
  }

  if (!"gvb_sequence_number" %in% names(d)) {
    gvb_alias <- intersect(
      c("gvb_sequence_number", "gbv_sequence_number", "gvb_seq_number", "gbv_seq_n"),
      names(d)
    )
    if (length(gvb_alias)) {
      d <- dplyr::rename(d, gvb_sequence_number = dplyr::all_of(gvb_alias[[1]]))
    }
  }
  if (!"gvb_sequence_number" %in% names(d)) {
    nms <- names(d)
    cand <- nms[
      grepl("sequence", nms, ignore.case = TRUE) &
        grepl("gvb|gbv", nms, ignore.case = TRUE) &
        !grepl("quantity|quant|quani", nms, ignore.case = TRUE)
    ]
    if (length(cand) == 1L) {
      d <- dplyr::rename(d, gvb_sequence_number = dplyr::all_of(cand))
    }
  }

  miss <- setdiff(c("time", "cq", "positive", "sample"), names(d))
  if (length(miss)) {
    stop(
      "After clean_names() and renames, `", rel_path, "` must include: ",
      paste(miss, collapse = ", "),
      ". Found: ",
      paste(names(d), collapse = ", ")
    )
  }

  d$cq <- suppressWarnings(as.numeric(d$cq))
  pv <- d$positive
  if (is.numeric(pv)) {
    d$positive <- ifelse(is.na(pv), NA_real_, as.integer(round(pv) != 0L))
  } else {
    ps <- toupper(stringr::str_trim(as.character(pv)))
    d$positive <- dplyr::case_when(
      is.na(ps) | ps == "" ~ NA_real_,
      ps %in% c("1", "Y", "YES", "TRUE", "T", "POS", "POSITIVE") ~ 1,
      TRUE ~ 0
    )
  }

  fac_cols <- intersect(c("sample", "rep", "sample_id"), names(d))
  for (cl in fac_cols) {
    d[[cl]] <- as.factor(d[[cl]])
  }

  # Preserve raw exposure text for `exp10_time_to_hours()` (ordered `time` can obscure cell text).
  exposure_label <- stringr::str_trim(as.character(d$time))

  tnum <- suppressWarnings(as.numeric(d$time))
  if (!all(is.na(tnum))) {
    tlv <- sort(unique(tnum[!is.na(tnum)]))
    d$time <- ordered(tnum, levels = tlv)
  } else {
    d$time <- ordered(as.character(d$time))
  }

  d$exposure_label <- exposure_label

  d %>%
    dplyr::mutate(
      positive_cut = dplyr::if_else(!is.na(.data$cq) & .data$cq > 39, 0L, as.integer(.data$positive))
    )
}

df.6.reads <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp6_reads_tom.xlsx",
  require_total_reads = TRUE
)
df.6.reads.no0 <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp6_reads_tom_no_time0.xlsx",
  require_total_reads = TRUE
)

# Design correction: 0 h exposure with quantity GVB coded as 0 (e.g. T0-* rows) are analyzed as quantity 1.
exp6_reads_recode_qty0_at_t0 <- function(dr) {
  dr %>%
    dplyr::mutate(
      .tnum = suppressWarnings(as.numeric(as.character(.data$time))),
      quantity_gbv = dplyr::if_else(
        .data$.tnum == 0 & dplyr::coalesce(as.numeric(.data$quantity_gbv), NA_real_) == 0,
        1,
        as.numeric(.data$quantity_gbv)
      )
    ) %>%
    dplyr::select(-.tnum)
}
df.6.reads <- exp6_reads_recode_qty0_at_t0(df.6.reads)
df.6.reads.no0 <- exp6_reads_recode_qty0_at_t0(df.6.reads.no0)
```

```{r exp6-reads-fig34-prep, include=FALSE}
# Shared y-range (count scale) and discrete palette for Figure 3 & 4 — avoids blue/green used in Figures 1–2 (Wong-style hues).
exp6_reads_fig_y_max_filtered <- function(dr) {
  if (!nrow(dr)) {
    return(0)
  }
  dr %>%
    dplyr::mutate(
      gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
      quantity_gbv = as.numeric(.data$quantity_gbv),
      total_reads = as.numeric(.data$total_reads)
    ) %>%
    dplyr::filter(
      !is.na(.data$gvb_sequence_number),
      !is.na(.data$quantity_gbv),
      !is.na(.data$total_reads),
      .data$total_reads >= 10000,
      .data$quantity_gbv != 0,
      suppressWarnings(as.numeric(as.character(.data$time))) != 0,
      suppressWarnings(as.numeric(as.character(.data$time))) %in% c(2, 24)
    ) %>%
    dplyr::summarise(mx = max(c(0L, .data$gvb_sequence_number), na.rm = TRUE), .groups = "drop") %>%
    dplyr::pull(1)
}
exp6_reads_fig_y_upper <- max(
  exp6_reads_fig_y_max_filtered(df.6.reads),
  exp6_reads_fig_y_max_filtered(df.6.reads.no0),
  na.rm = TRUE
) * 1.45
exp6_reads_fig_y_upper <- max(exp6_reads_fig_y_upper, 10)
exp6_reads_y_breaks <- sort(unique(c(0, 1, 10, 100, 1000, 10000, 100000)))
exp6_reads_y_breaks_use <- exp6_reads_y_breaks[exp6_reads_y_breaks <= exp6_reads_fig_y_upper]
exp6_reads_y_labels <- function(x) ifelse(x == 0, "0", scales::label_number(big.mark = ",")(x))
# Numeric x for 2 h / 24 h (1 and 2) so raw jitter can sit slightly right of estimates; axis still labelled in hours.
exp6_reads_fig_x_nudge_raw <- 0.1
exp6_reads_fig_xlim <- c(0.45, 2.72)
exp6_reads_xlim_for_n <- function(n) c(0.45, n + 0.27)
exp6_reads_qty_dodge_w <- 0.85 * 0.75
# Figure 3 only: small negative floor (log1p domain x > -1) so means/CIs near zero are not clipped.
exp6_reads_fig_y_lower_f3 <- -0.18
exp6_reads_fig_legend_npc <- c(
  x = (2 - exp6_reads_fig_xlim[1]) / diff(exp6_reads_fig_xlim),
  y = (log1p(10) - log1p(exp6_reads_fig_y_lower_f3)) /
    (log1p(exp6_reads_fig_y_upper) - log1p(exp6_reads_fig_y_lower_f3))
)
exp6_reads_qty_palette <- function(n) {
  wong <- c(
    "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
    "#D55E00", "#CC79A7"
  )
  wong[seq_len(min(n, length(wong)))]
}
```

<details><summary>Data frames</summary>

```{r exp6-summary-tibbles, echo=FALSE}
cat("Original time-course (Exp6-time2.xlsx)\n\n")
brief_dataframe(dplyr::count(df.6, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.6)

cat("\n\nOriginal density design (2 h / 24 h number sheets)\n\n")
brief_dataframe(dplyr::count(df.6.d, sample, number, name = "n"), "Counts by sample × number")
brief_dataframe(df.6.d)

cat("\n\nReads — full (GVB_exp6_reads_tom.xlsx)\n\n")
brief_dataframe(dplyr::count(df.6.reads, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.6.reads)

cat("\n\nReads — no time 0 (GVB_exp6_reads_tom_no_time0.xlsx)\n\n")
brief_dataframe(dplyr::count(df.6.reads.no0, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.6.reads.no0)
```

</details>

::: {.panel-tabset}

```{r exp6-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

mod_1 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  data = df.6,
  file = here::here("outputs/models", "mod_1"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod_2 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  data = df.6,
  file = here::here("outputs/models", "mod_2"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive ~ mo(time) + mo(number) + (1|sample_id))

mod_3 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  data = df.6.d,
  file = here::here("outputs/models", "mod_3"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + mo(number) + (1|sample_id))

mod_4 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  data = df.6.d,
  file = here::here("outputs/models", "mod_4"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp6-waic-compute, include=FALSE}
waic_tbl_one_exp <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_exp(mod_1, "mod_1 (time monotonic, no cut)"),
  waic_tbl_one_exp(mod_2, "mod_2 (time monotonic, cut)"),
  waic_tbl_one_exp(mod_3, "mod_3 (time + number additive, no cut)"),
  waic_tbl_one_exp(mod_4, "mod_4 (time + number additive, cut)")
) %>%
  dplyr::arrange(WAIC, Model) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

## Effect plots

```{r exp6-eff1, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_1 <- conditional_effects(mod_1)[[1]] %>% mutate(cutoff = "No")
c_eff_2 <- conditional_effects(mod_2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_1, c_eff_2)

raw_exp6_eff1 <- dplyr::bind_rows(
  dplyr::mutate(df.6, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.6, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
) %>%
  dplyr::mutate(y_plot = binary_y_for_plot(.data$y))

y_lim_exp6_eff1 <- c(-0.13, 1.13)
tlv_eff1 <- levels(c_eff$time)
n_t_eff1 <- length(tlv_eff1)
ichr <- as.character(tlv_eff1)
i24_eff1 <- which(ichr == "24")[1L]
if (is.na(i24_eff1)) i24_eff1 <- which(suppressWarnings(as.numeric(ichr)) == 24)[1L]
if (is.na(i24_eff1)) i24_eff1 <- max(1L, round(n_t_eff1 / 2))
x_npc_eff1 <- (i24_eff1 - 0.5) / n_t_eff1
y_npc_eff1 <- (0.3 - y_lim_exp6_eff1[1]) / (y_lim_exp6_eff1[2] - y_lim_exp6_eff1[1])

p_exp6_eff1 <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff1,
    aes(x = time, y = y_plot, group = cutoff, color = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.03, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  theme(
    legend.position = c(x_npc_eff1, y_npc_eff1),
    legend.justification = c(0.5, 0.5),
    legend.margin = ggplot2::margin(4, 4, 4, 4),
    legend.background = element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = y_lim_exp6_eff1,
    breaks = prob_det_y_breaks,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title="Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-01-exp6-detection-mo-time-cutoff.png"),
  plot = p_exp6_eff1,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff1
```

Figure 1: Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Jittered points are raw binary outcomes by exposure time, coloured by cutoff; error bars are 95% credible intervals and coloured points are posterior means.

```{r exp6-eff2, warning=FALSE, echo=FALSE, fig.width=7, fig.height=5, out.width='100%', dpi = 600}
# conditional_effects(c("time","number")) fixes number at its reference level; use
# effects = "time" with conditions = cond_ce6_time_number for all time × number cells.
c_eff_3 <- conditional_effects(mod_3, effects = "time", conditions = cond_ce6_time_number)[[1]] %>% mutate(cutoff = "No")
c_eff_4 <- conditional_effects(mod_4, effects = "time", conditions = cond_ce6_time_number)[[1]] %>% mutate(cutoff = "Yes") %>% dplyr::rename(positive = positive_cut)

c_eff <- rbind(c_eff_3, c_eff_4) %>%
  mutate(time = paste0("Time: ", time, " hrs")) %>%
  mutate(number = paste0(number, " GVB"))

c_eff$cutoff <- factor(c_eff$cutoff, levels = c("No", "Yes"))

raw_exp6_eff2 <- df.6.d %>%
  dplyr::mutate(
    time = paste0("Time: ", .data$time, " hrs"),
    number = paste0(.data$number, " GVB")
  ) %>%
  tidyr::pivot_longer(
    cols = c(positive, positive_cut),
    names_to = "which_y",
    values_to = "y"
  ) %>%
  dplyr::mutate(
    cutoff = factor(
      dplyr::if_else(.data$which_y == "positive", "No", "Yes"),
      levels = c("No", "Yes")
    ),
    y_plot = binary_y_for_plot(.data$y)
  )

p_exp6_eff2 <- c_eff %>% ggplot(aes(x = cutoff, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp6_eff2,
    aes(x = cutoff, y = y_plot, color = cutoff),
    inherit.aes = FALSE,
    position = position_jitter(width = 0.15, height = 0.03),
    alpha = 0.35,
    size = 0.85,
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = 0.5), size = 3, alpha = 0.5) +
  theme_classic() +
  theme(
    strip.text = element_text(size = 9, margin = ggplot2::margin(4, 4, 4, 4)),
    strip.background.x = element_rect(fill = "white", colour = "grey35", linewidth = 0.5),
    strip.background.y = element_rect(fill = "white", colour = "grey35", linewidth = 0.5),
    panel.background = element_rect(fill = "white", colour = NA),
    panel.border = element_rect(colour = "grey35", fill = NA, linewidth = 0.6),
    panel.grid.major.x = element_line(colour = "grey85", linewidth = 0.3),
    panel.grid.major.y = element_line(colour = "grey85", linewidth = 0.3)
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "Cq cut off (> 39)", y = "Probability of detection (%)") +
  guides(color = "none") +
  facet_grid(number ~ time) +
  scale_y_continuous(
    labels = scales::percent,
    limits = c(-0.13, 1.13),
    breaks = prob_det_y_breaks,
    expand = ggplot2::expansion(mult = c(0, 0))
  )

ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-02-exp6-detection-mo-time-by-time-number.png"),
  plot = p_exp6_eff2,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp6_eff2
```

Figure 2: Conditional effect plot of GBV exposure on probability of detection, faceted by time and number of GVB, with (yes = green) and without (no = blue) a Cq cut off of 39. Jittered points are raw outcomes, coloured by cutoff; error bars are 95% credible intervals and coloured points are posterior means.

## Tables

```{r exp6-estimates, echo=FALSE}
# Modeled estimates from conditional_effects() on the probability scale.
# Convert to percent for reporting.

ce6_time <- dplyr::bind_rows(
  conditional_effects(mod_1)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod_2)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

ce6_time_number <- dplyr::bind_rows(
  conditional_effects(mod_3, effects = "time", conditions = cond_ce6_time_number)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod_4, effects = "time", conditions = cond_ce6_time_number)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Number (GVB)` = number,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`, `Number (GVB)`)

knitr::kable(
  ce6_time,
  caption = "Modeled estimates from conditional effects (Experiment 6): time",
  format = "html"
)

knitr::kable(
  ce6_time_number,
  caption = "Modeled estimates from conditional effects (Experiment 6): time and number (additive)",
  format = "html"
)
```

## Model comparison

```{r exp6-waic-tab, echo=FALSE}
knitr::kable(
  exp6_waic_tbl,
  caption = "Experiment 6: WAIC (lower is better; `mo(time)` formulations).",
  format = "html"
)
```

## Models

```{r exp6-models-show, ref.label="exp6-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp6-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_1, ndraws = 200, col = 2)
pp_check(mod_2, ndraws = 200, col = 2)
pp_check(mod_3, ndraws = 200, col = 2)
pp_check(mod_4, ndraws = 200, col = 2)
```

## Output

::: {.panel-tabset}

### mod_1

```{r exp6-summary-mod1, echo=FALSE}
#| class-output: model-output-panel
summary(mod_1)
```

### mod_2

```{r exp6-summary-mod2, echo=FALSE}
#| class-output: model-output-panel
summary(mod_2)
```

### mod_3

```{r exp6-summary-mod3, echo=FALSE}
#| class-output: model-output-panel
summary(mod_3)
```

### mod_4

```{r exp6-summary-mod4, echo=FALSE}
#| class-output: model-output-panel
summary(mod_4)
```

:::

<!-- Pairwise comparisons removed -->

```{r exp6-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod_1, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  },
  {
    emm <- emmeans::emmeans(mod_2, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d <- d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    d
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)

pairwise_number <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod_3, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  },
  {
    emm <- emmeans::emmeans(mod_4, ~ number | time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", by = "time", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    if ("time" %in% names(d)) {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = time, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    } else {
      d <- d %>%
        mutate(
          `Difference (pp)` = round(.data[[est_col]] * 100, 2),
          `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
          `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
        ) %>%
        transmute(Cutoff, `Time (hrs)` = NA, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
    }
    d
  }
)

knitr::kable(
  pairwise_number,
  caption = "Pairwise comparisons between number levels within each time (Experiment 6; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

### GVB sequence counts - reads

::: {.panel-tabset}

```{r exp6-reads-models-fit, include=FALSE}
# `df_exp6_reads_mod` (mod_5–mod_8): 2 h & 24 h only — density design for Figure 3a / 4a — not Figure 4b.
# Negative-binomial reads models: require total_reads >= 10,000 (excludes low-library rows).
df_exp6_reads_mod <- df.6.reads %>%
  dplyr::mutate(
    quantity_gbv = ifelse(quantity_gbv == 0, 1, quantity_gbv),
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv),
    total_reads = as.numeric(.data$total_reads)
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    !is.na(.data$total_reads),
    .data$total_reads >= 10000,
    .data$gvb_sequence_number >= 0L,
    suppressWarnings(as.numeric(as.character(.data$time))) %in% c(2, 24)
  ) %>%
  dplyr::mutate(
    time = ordered(
      suppressWarnings(as.numeric(as.character(.data$time))),
      levels = c(2, 24)
    ),
    quantity_gbv = ordered(
      .data$quantity_gbv,
      levels = sort(unique(stats::na.omit(as.numeric(.data$quantity_gbv))))
    )
  )

qty_fac_lv_reads <- levels(df_exp6_reads_mod$quantity_gbv)
df_exp6_reads_fac <- df_exp6_reads_mod %>%
  dplyr::mutate(
    time = factor(as.character(.data$time), levels = c("2", "24")),
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = qty_fac_lv_reads)
  )

mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

mod_5 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_mod,
  file = here::here("outputs/models", "mod_5"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

mod_6 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_mod,
  file = here::here("outputs/models", "mod_6"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

mod_7 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_fac,
  file = here::here("outputs/models", "mod_7"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

mod_8 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_fac,
  file = here::here("outputs/models", "mod_8"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

# Quantity GVB = 1 marginal (Figure 3b); 0 h q=1 comes from `exp6_reads_recode_qty0_at_t0()` on df.6.reads.
df_exp6_reads_q1_base <- df.6.reads %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv),
    total_reads = as.numeric(.data$total_reads),
    tnum = suppressWarnings(as.numeric(as.character(.data$time)))
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    !is.na(.data$total_reads),
    !is.na(.data$tnum),
    # At 0 h keep rows with any positive library; elsewhere keep 10k floor (so 0 h is not dropped only for low depth).
    ((.data$tnum == 0) & (.data$total_reads > 0)) | ((.data$tnum != 0) & (.data$total_reads >= 10000)),
    .data$gvb_sequence_number >= 0L,
    .data$quantity_gbv == 1
  )

exp6_reads_q1_time_levels_fit <- sort(unique(stats::na.omit(df_exp6_reads_q1_base$tnum)))
exp6_reads_design_hours <- sort(unique(suppressWarnings(as.numeric(as.character(df.6.reads$time)))))
# X-axis for 3b: include 0 h whenever the workbook has that exposure (even if no q=1 row at 0).
exp6_reads_q1_time_levels_plot <- exp6_reads_q1_time_levels_fit
if (length(exp6_reads_design_hours) && 0 %in% exp6_reads_design_hours) {
  exp6_reads_q1_time_levels_plot <- sort(unique(c(0, exp6_reads_q1_time_levels_fit)))
}
exp6_reads_q1_ok <- nrow(df_exp6_reads_q1_base) > 0L && length(exp6_reads_q1_time_levels_fit) >= 2L

if (exp6_reads_q1_ok) {
  df_exp6_reads_q1_mod <- df_exp6_reads_q1_base %>%
    dplyr::select(-tnum) %>%
    dplyr::mutate(
      time = ordered(
        suppressWarnings(as.numeric(as.character(.data$time))),
        levels = exp6_reads_q1_time_levels_fit
      )
    )

  df_exp6_reads_q1_fac <- df_exp6_reads_q1_mod %>%
    dplyr::mutate(
      time = factor(
        as.character(suppressWarnings(as.numeric(as.character(.data$time)))),
        levels = as.character(exp6_reads_q1_time_levels_fit)
      )
    )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + offset(log(total_reads)))

  mod_9 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp6_reads_q1_mod |> filter(total_reads >= 10000),
    file = here::here("outputs/models", "mod_9"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time + offset(log(total_reads)))

  mod_10 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp6_reads_q1_fac |> filter(total_reads >= 10000),
    file = here::here("outputs/models", "mod_10"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )
} else {
  mod_9 <- NULL
  mod_10 <- NULL
}
```

```{r exp6-reads-waic-compute, include=FALSE}
waic_tbl_one_reads <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_reads_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_reads(mod_5, "mo(time) + mo(quantity_gbv)"),
  waic_tbl_one_reads(mod_6, "mo(time) * mo(quantity_gbv)"),
  waic_tbl_one_reads(mod_7, "time + quantity_gbv (factor)"),
  waic_tbl_one_reads(mod_8, "time * quantity_gbv (factor)")
) %>%
  dplyr::arrange(WAIC) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)

exp6_reads_q1_waic_tbl <- if (exp6_reads_q1_ok) {
  dplyr::bind_rows(
    waic_tbl_one_reads(mod_9, "mo(time)"),
    waic_tbl_one_reads(mod_10, "time (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
} else {
  NULL
}

exp6_reads_q1_plot_mod <- if (
  exp6_reads_q1_ok &&
    !is.null(exp6_reads_q1_waic_tbl) &&
    nrow(exp6_reads_q1_waic_tbl) > 0L
) {
  if (identical(as.character(exp6_reads_q1_waic_tbl$Model[[1]]), "time (factor)")) {
    mod_10
  } else {
    mod_9
  }
} else {
  NULL
}
exp6_reads_q1_plot_label <- if (
  exp6_reads_q1_ok &&
    !is.null(exp6_reads_q1_waic_tbl) &&
    nrow(exp6_reads_q1_waic_tbl) > 0L
) {
  as.character(exp6_reads_q1_waic_tbl$Model[[1]])
} else {
  ""
}
```

#### Effect plots

```{r exp6-reads-eff, warning=FALSE, echo=FALSE, fig.width=14, fig.height=4.5, out.width="100%", dpi=600}
raw <- df_exp6_reads_mod

ce_both <- conditional_effects(mod_6, effects = "time:quantity_gbv")[[1]]
qty_lv <- sort(unique(as.numeric(as.character(ce_both$quantity_gbv))))
ce_both <- ce_both %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv))
  ) %>%
  dplyr::mutate(
    tord = suppressWarnings(as.numeric(as.character(.data$time))),
    xtime = as.numeric(.data$time)
  ) %>%
  dplyr::arrange(.data$quantity_gbv, .data$tord)

raw_col <- raw %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both$quantity_gbv)),
    xtime = as.numeric(.data$time) + exp6_reads_fig_x_nudge_raw
  )

n_qty <- dplyr::n_distinct(ce_both$quantity_gbv)
pal <- exp6_reads_qty_palette(n_qty)
ce_pd <- ggplot2::position_dodge(width = exp6_reads_qty_dodge_w)

exp6_reads_plot_theme <- ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
    panel.grid.minor.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
    legend.position = unname(exp6_reads_fig_legend_npc),
    legend.justification = c(0.5, 0.5),
    legend.title = ggplot2::element_text(size = 10),
    legend.text = ggplot2::element_text(size = 9),
    legend.margin = ggplot2::margin(2, 4, 2, 4),
    legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  )

p_exp6_reads_eff <- ggplot2::ggplot(
  ce_both,
  ggplot2::aes(
    x = .data$xtime,
    y = .data$estimate__,
    ymin = .data$lower__,
    ymax = .data$upper__,
    color = .data$quantity_gbv
  )
) +
  ggplot2::geom_point(
    data = raw_col,
    ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
    inherit.aes = FALSE,
    position = ggplot2::position_jitterdodge(
      jitter.width = 0.12,
      jitter.height = 0.175,
      dodge.width = exp6_reads_qty_dodge_w
    ),
    alpha = 0.4,
    size = 0.85,
    stroke = 0
  ) +
  ggplot2::geom_errorbar(position = ce_pd, width = 0.12, linewidth = 0.9) +
  ggplot2::geom_point(position = ce_pd, size = 2.5, alpha = 0.9) +
  ggplot2::scale_color_manual(values = pal) +
  ggplot2::scale_x_continuous(breaks = 1:2, labels = c("2", "24"), limits = exp6_reads_fig_xlim, expand = c(0, 0)) +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    limits = c(exp6_reads_fig_y_lower_f3, exp6_reads_fig_y_upper),
    breaks = exp6_reads_y_breaks_use,
    labels = exp6_reads_y_labels,
    minor_breaks = NULL,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  exp6_reads_plot_theme +
  ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      title = "GVB quantity",
      reverse = TRUE
    )
  )

exp6_reads_combo_tag_theme <- ggplot2::theme(
  plot.tag = ggplot2::element_text(size = 11, face = "bold", hjust = 0, vjust = 1)
)
exp6_reads_combo_legend_theme <- ggplot2::theme(
  legend.position = c(0.97, 0.2),
  legend.justification = c(1, 0.5)
)

if (exp6_reads_q1_ok) {
  n_q1 <- length(exp6_reads_q1_time_levels_plot)
  xlim_q1 <- exp6_reads_xlim_for_n(n_q1)
  ce_time_q1 <- conditional_effects(exp6_reads_q1_plot_mod, effects = "time")[[1]] %>%
    dplyr::mutate(
      hrs = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = match(.data$hrs, exp6_reads_q1_time_levels_plot)
    )

  raw_col_q1 <- df_exp6_reads_q1_mod %>%
    dplyr::mutate(
      hrs = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = match(.data$hrs, exp6_reads_q1_time_levels_plot) + exp6_reads_fig_x_nudge_raw
    )

  y_upper_q1 <- max(
    exp6_reads_fig_y_upper,
    suppressWarnings(max(as.numeric(ce_time_q1$upper__), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(ce_time_q1$estimate__), na.rm = TRUE)),
    suppressWarnings(max(raw_col_q1$gvb_sequence_number, na.rm = TRUE)),
    10,
    na.rm = TRUE
  )
  y_breaks_q1 <- exp6_reads_y_breaks[exp6_reads_y_breaks <= y_upper_q1]

  exp6_reads_plot_theme_q1 <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
    )

  col_mu_q1 <- "#0072B2"
  p_exp6_reads_q1_eff <- ggplot2::ggplot(
    ce_time_q1,
    ggplot2::aes(x = .data$xtime, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
  ) +
    ggplot2::geom_point(
      data = raw_col_q1,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number),
      inherit.aes = FALSE,
      position = ggplot2::position_jitter(width = 0.1, height = 0.175),
      alpha = 0.4,
      size = 0.85,
      color = "grey40",
      stroke = 0
    ) +
    ggplot2::geom_errorbar(width = 0.12, linewidth = 0.9, color = col_mu_q1) +
    ggplot2::geom_point(size = 2.5, alpha = 0.9, color = col_mu_q1) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_q1),
      labels = as.character(exp6_reads_q1_time_levels_plot),
      limits = xlim_q1,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(exp6_reads_fig_y_lower_f3, y_upper_q1),
      breaks = y_breaks_q1,
      labels = exp6_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp6_reads_plot_theme_q1 +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

  p_exp6_reads_eff_combo <- patchwork::wrap_plots(
    p_exp6_reads_eff +
      exp6_reads_combo_legend_theme +
      exp6_reads_combo_tag_theme +
      ggplot2::labs(tag = "a)"),
    p_exp6_reads_q1_eff +
      ggplot2::labs(
        caption = "Quantity GVB = 1.",
        tag = "b)"
      ) +
      ggplot2::theme(
        plot.caption = ggplot2::element_text(size = 8.5, colour = "grey25", hjust = 0),
        plot.caption.position = "panel"
      ) +
      exp6_reads_combo_tag_theme,
    ncol = 2
  )
} else {
  p_exp6_reads_q1_eff <- NULL
  p_exp6_reads_eff_combo <- patchwork::wrap_plots(
    p_exp6_reads_eff +
      exp6_reads_combo_legend_theme +
      exp6_reads_combo_tag_theme +
      ggplot2::labs(tag = "a)"),
    ncol = 1
  )
}

ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-03-exp6-reads-panel-a-only.png"),
  plot = p_exp6_reads_eff,
  width = 7,
  height = 4.5,
  dpi = 600
)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-03-exp6-reads-panels-a-b.png"),
  plot = p_exp6_reads_eff_combo,
  width = 14,
  height = 4.5,
  dpi = 600
)

p_exp6_reads_eff_combo
```

Figure 3: GVB sequence counts from the reads workbook. **a)** 2 h and 24 h only — mean sequence count by GVB quantity. **b)** When shown: quantity GVB = 1 at all exposure times in the data **(including 0 h when present)** — marginal effect of exposure time on mean count. Vertical axis is log(1 + count). Jittered points are raw counts; error bars are 95% credible intervals and coloured points are posterior means.

#### Tables

Tabulated estimates use the **preferred monotonic interaction** model **`mo(time) * mo(quantity_gbv)`** (negative binomial mean count with **`offset(log(total_reads))`**; marginal **`conditional_effects()`**). See the **Model comparison** tab for **WAIC** across all candidate models.

```{r exp6-reads-estimates, echo=FALSE}
ce_exp6_t <- conditional_effects(mod_6, effects = "time")[[1]] %>%
  dplyr::transmute(
    `Time (hrs)` = time,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

ce_exp6_q <- conditional_effects(mod_6, effects = "quantity_gbv")[[1]] %>%
  dplyr::transmute(
    `Quantity GVB` = quantity_gbv,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

knitr::kable(
  ce_exp6_t,
  caption = "Experiment 6 reads — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
  format = "html"
)

knitr::kable(
  ce_exp6_q,
  caption = "Experiment 6 reads — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
  format = "html"
)
```

**Quantity GVB = 1 — all exposure times in workbook** — tabulated marginal time effects use the **lower-WAIC** q1 model (**`mo(time)`** or **factor `time`**; **Model comparison** below).

```{r exp6-reads-estimates-q1, echo=FALSE}
if (exp6_reads_q1_ok) {
  ce_exp6_q1_t <- conditional_effects(exp6_reads_q1_plot_mod, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp6_q1_t,
    caption = paste0(
      "Experiment 6 reads — quantity GVB = 1 (0 h rows recoded from quantity 0 in workbook): lowest-WAIC q1 model (",
      exp6_reads_q1_plot_label,
      ") with offset — conditional effect of time (mean count)."
    ),
    format = "html"
  )
} else {
  knitr::kable(
    tibble::tibble(Note = "Not fitted: need \u2265 1 row with quantity_gbv = 1 and \u2265 2 distinct exposure times after filtering."),
    caption = "Experiment 6 reads — quantity GVB = 1 (all times) — no table.",
    format = "html"
  )
}
```

#### Models

```{r exp6-reads-models-show, ref.label="exp6-reads-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp6-reads-waic-tab, echo=FALSE}
knitr::kable(
  exp6_reads_waic_tbl,
  caption = "Experiment 6 reads: WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
if (exp6_reads_q1_ok) {
  knitr::kable(
    exp6_reads_q1_waic_tbl,
    caption = "Experiment 6 reads — **quantity GVB = 1, all exposure times**: WAIC for `mo(time)` vs factor `time` (offset log(total reads); distinct comparison from the table above).",
    format = "html"
  )
}
```

#### Validation

```{r exp6-reads-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_5, ndraws = 200)
pp_check(mod_6, ndraws = 200)
pp_check(mod_7, ndraws = 200)
pp_check(mod_8, ndraws = 200)
```

**Quantity GVB = 1 (all exposure times)** — posterior predictive checks: **`mo(time)`** vs **factor `time`**.

```{r exp6-reads-pp-check-q1}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
if (exp6_reads_q1_ok) {
  pp_check(mod_9, ndraws = 200)
  pp_check(mod_10, ndraws = 200)
}
```

#### Output

::: {.panel-tabset}

##### mod_5

```{r exp6-reads-summary-base, echo=FALSE}
#| class-output: model-output-panel
summary(mod_5)
```

##### mod_6

```{r exp6-reads-summary-mo-int, echo=FALSE}
#| class-output: model-output-panel
summary(mod_6)
```

##### mod_7

```{r exp6-reads-summary-fac-add, echo=FALSE}
#| class-output: model-output-panel
summary(mod_7)
```

##### mod_8

```{r exp6-reads-summary-fac-int, echo=FALSE}
#| class-output: model-output-panel
summary(mod_8)
```

##### Quantity GVB = 1 (q1)

```{r exp6-reads-summary-q1, echo=FALSE}
#| class-output: model-output-panel
if (exp6_reads_q1_ok) {
  summary(mod_9)
  summary(mod_10)
}
```

:::

:::

### GVB sequence counts - no time 0

::: {.panel-tabset}

```{r exp6-reads-no0-models-fit, include=FALSE}
df_exp6_reads_no0_mod <- df.6.reads.no0 %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv),
    total_reads = as.numeric(.data$total_reads),
    tnum_rd = suppressWarnings(as.numeric(as.character(.data$time)))
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    !is.na(.data$total_reads),
    .data$total_reads >= 10000,
    .data$gvb_sequence_number >= 0L,
    .data$quantity_gbv != 0,
    !is.na(.data$tnum_rd),
    .data$tnum_rd != 0,
    .data$tnum_rd %in% c(2, 24)
  ) %>%
  dplyr::select(-tnum_rd) %>%
  dplyr::mutate(
    time = ordered(
      suppressWarnings(as.numeric(as.character(.data$time))),
      levels = c(2, 24)
    ),
    quantity_gbv = ordered(
      .data$quantity_gbv,
      levels = sort(unique(stats::na.omit(as.numeric(.data$quantity_gbv))))
    )
  )

qty_fac_lv_no0 <- levels(df_exp6_reads_no0_mod$quantity_gbv)
df_exp6_reads_no0_fac <- df_exp6_reads_no0_mod %>%
  dplyr::mutate(
    time = factor(as.character(.data$time), levels = c("2", "24")),
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = qty_fac_lv_no0)
  )

mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

mod_11 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_mod,
  file = here::here("outputs/models", "mod_11"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

mod_12 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_mod,
  file = here::here("outputs/models", "mod_12"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

mod_13 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_fac,
  file = here::here("outputs/models", "mod_13"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

mod_14 <- brm(
  mod_formula, family = negbinomial(),
  data = df_exp6_reads_no0_fac,
  file = here::here("outputs/models", "mod_14"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

# Quantity-1 marginal (Figure 4b): `df.6.reads` (full workbook) — not `df_exp6_reads_mod` (2h/24h only).
# mod_15–mod_16; 0 h is not filtered by time; at 0 h the library rule below allows total_reads > 0 only.
df_exp6_reads_no0_q1_base <- df.6.reads %>%
  dplyr::mutate(
    gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
    quantity_gbv = as.numeric(.data$quantity_gbv),
    total_reads = as.numeric(.data$total_reads),
    tnum = suppressWarnings(as.numeric(as.character(.data$time)))
  ) %>%
  dplyr::filter(
    !is.na(.data$gvb_sequence_number),
    !is.na(.data$quantity_gbv),
    !is.na(.data$time),
    !is.na(.data$total_reads),
    !is.na(.data$tnum),
    ((.data$tnum == 0) & (.data$total_reads > 0)) | ((.data$tnum != 0) & (.data$total_reads >= 10000)),
    .data$gvb_sequence_number >= 0L,
    .data$quantity_gbv == 1
  )

exp6_reads_no0_q1_time_levels_fit <- sort(unique(stats::na.omit(df_exp6_reads_no0_q1_base$tnum)))
exp6_reads_no0_q1_time_levels_plot <- exp6_reads_no0_q1_time_levels_fit
if (length(exp6_reads_design_hours) && 0 %in% exp6_reads_design_hours) {
  exp6_reads_no0_q1_time_levels_plot <- sort(unique(c(0, exp6_reads_no0_q1_time_levels_fit)))
}
exp6_reads_no0_q1_ok <- nrow(df_exp6_reads_no0_q1_base) > 0L && length(exp6_reads_no0_q1_time_levels_fit) >= 2L

if (exp6_reads_no0_q1_ok) {
  df_exp6_reads_no0_q1_mod <- df_exp6_reads_no0_q1_base %>%
    dplyr::select(-tnum) %>%
    dplyr::mutate(
      time = ordered(
        suppressWarnings(as.numeric(as.character(.data$time))),
        levels = exp6_reads_no0_q1_time_levels_fit
      )
    )

  df_exp6_reads_no0_q1_fac <- df_exp6_reads_no0_q1_mod %>%
    dplyr::mutate(
      time = factor(
        as.character(suppressWarnings(as.numeric(as.character(.data$time)))),
        levels = as.character(exp6_reads_no0_q1_time_levels_fit)
      )
    )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + offset(log(total_reads)))

  mod_15 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp6_reads_no0_q1_mod|> filter(total_reads >= 10000),
    file = here::here("outputs/models", "mod_15"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time + offset(log(total_reads)))

  mod_16 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp6_reads_no0_q1_fac |> filter(total_reads >= 10000),
    file = here::here("outputs/models", "mod_16"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )
} else {
  mod_15 <- NULL
  mod_16 <- NULL
}
```

```{r exp6-reads-no0-waic-compute, include=FALSE}
waic_tbl_one_reads_no0 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp6_reads_no0_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_reads_no0(mod_11, "mo(time) + mo(quantity_gbv)"),
  waic_tbl_one_reads_no0(mod_12, "mo(time) * mo(quantity_gbv)"),
  waic_tbl_one_reads_no0(mod_13, "time + quantity_gbv (factor)"),
  waic_tbl_one_reads_no0(mod_14, "time * quantity_gbv (factor)")
) %>%
  dplyr::arrange(WAIC) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)

exp6_reads_no0_q1_waic_tbl <- if (exp6_reads_no0_q1_ok) {
  dplyr::bind_rows(
    waic_tbl_one_reads_no0(mod_15, "mo(time)"),
    waic_tbl_one_reads_no0(mod_16, "time (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
} else {
  NULL
}

# WAIC-best q1 time parameterization (for Model comparison tab / narrative)
exp6_reads_no0_q1_waic_preferred_label <- if (
  exp6_reads_no0_q1_ok &&
    !is.null(exp6_reads_no0_q1_waic_tbl) &&
    nrow(exp6_reads_no0_q1_waic_tbl) > 0L
) {
  as.character(exp6_reads_no0_q1_waic_tbl$Model[[1]])
} else {
  ""
}

# Figure 4b + q1 table: same WAIC rule as Figure 3b
exp6_reads_no0_q1_plot_mod <- if (
  exp6_reads_no0_q1_ok &&
    !is.null(exp6_reads_no0_q1_waic_tbl) &&
    nrow(exp6_reads_no0_q1_waic_tbl) > 0L
) {
  if (identical(as.character(exp6_reads_no0_q1_waic_tbl$Model[[1]]), "time (factor)")) {
    mod_16
  } else {
    mod_15
  }
} else {
  NULL
}
exp6_reads_no0_q1_plot_label <- if (
  exp6_reads_no0_q1_ok &&
    !is.null(exp6_reads_no0_q1_waic_tbl) &&
    nrow(exp6_reads_no0_q1_waic_tbl) > 0L
) {
  as.character(exp6_reads_no0_q1_waic_tbl$Model[[1]])
} else {
  ""
}
```

#### Effect plots

```{r exp6-reads-no0-eff, warning=FALSE, echo=FALSE, fig.width=14, fig.height=4.5, out.width="100%", dpi=600}
raw_no0 <- df_exp6_reads_no0_mod

ce_both_no0 <- conditional_effects(mod_12, effects = "time:quantity_gbv")[[1]] %>%
  dplyr::mutate(.ce_tnum = suppressWarnings(as.numeric(as.character(.data$time)))) %>%
  dplyr::filter(.data$.ce_tnum %in% c(2, 24)) %>%
  dplyr::select(-.ce_tnum)
qty_lv_no0 <- sort(unique(as.numeric(as.character(ce_both_no0$quantity_gbv))))
ce_both_no0 <- ce_both_no0 %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv_no0))
  ) %>%
  dplyr::mutate(
    tord = suppressWarnings(as.numeric(as.character(.data$time))),
    xtime = as.numeric(.data$time)
  ) %>%
  dplyr::arrange(.data$quantity_gbv, .data$tord)

raw_col_no0 <- raw_no0 %>%
  dplyr::mutate(
    quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both_no0$quantity_gbv)),
    xtime = as.numeric(.data$time) + exp6_reads_fig_x_nudge_raw
  )

exp6_reads_fig_y_max_no0 <- max(
  0,
  suppressWarnings(max(as.numeric(ce_both_no0$upper__), na.rm = TRUE)),
  suppressWarnings(max(as.numeric(ce_both_no0$estimate__), na.rm = TRUE)),
  suppressWarnings(max(raw_col_no0$gvb_sequence_number, na.rm = TRUE)),
  na.rm = TRUE
)
# Generous headroom so means, 95% interval tops, and geom glyphs are not clipped (log1p axis, dodge=0.85).
exp6_reads_fig_y_upper_no0 <- max(
  exp6_reads_fig_y_upper,
  exp6_reads_fig_y_max_no0 * 1.72 + max(80, 0.12 * exp6_reads_fig_y_max_no0) + max(40, 0.04 * exp6_reads_fig_y_max_no0),
  10
)
exp6_reads_y_breaks_use_no0 <- exp6_reads_y_breaks[exp6_reads_y_breaks <= exp6_reads_fig_y_upper_no0]
exp6_reads_fig_legend_npc_no0 <- c(
  x = (2 - exp6_reads_fig_xlim[1]) / diff(exp6_reads_fig_xlim),
  y = log1p(10) / log1p(exp6_reads_fig_y_upper_no0)
)

n_qty_no0 <- dplyr::n_distinct(ce_both_no0$quantity_gbv)
pal_no0 <- exp6_reads_qty_palette(n_qty_no0)
ce_pd_no0 <- ggplot2::position_dodge(width = exp6_reads_qty_dodge_w)

exp6_reads_plot_theme_no0 <- ggplot2::theme_bw() +
  ggplot2::theme(
    panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
    panel.grid.minor.y = ggplot2::element_blank(),
    panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
    legend.position = unname(exp6_reads_fig_legend_npc_no0),
    legend.justification = c(0.5, 0.5),
    legend.title = ggplot2::element_text(size = 10),
    legend.text = ggplot2::element_text(size = 9),
    legend.margin = ggplot2::margin(2, 4, 2, 4),
    legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  )

p_exp6_reads_no0_eff <- ggplot2::ggplot(
  ce_both_no0,
  ggplot2::aes(
    x = .data$xtime,
    y = .data$estimate__,
    ymin = .data$lower__,
    ymax = .data$upper__,
    color = .data$quantity_gbv
  )
) +
  ggplot2::geom_point(
    data = raw_col_no0,
    ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
    inherit.aes = FALSE,
    position = ggplot2::position_jitterdodge(
      jitter.width = 0.12,
      jitter.height = 0.175,
      dodge.width = exp6_reads_qty_dodge_w
    ),
    alpha = 0.4,
    size = 0.85,
    stroke = 0
  ) +
  ggplot2::geom_errorbar(position = ce_pd_no0, width = 0.12, linewidth = 0.9) +
  ggplot2::geom_point(position = ce_pd_no0, size = 2.5, alpha = 0.9) +
  ggplot2::scale_color_manual(values = pal_no0) +
  ggplot2::scale_x_continuous(breaks = 1:2, labels = c("2", "24"), limits = exp6_reads_fig_xlim, expand = c(0, 0)) +
  ggplot2::scale_y_continuous(
    trans = scales::log1p_trans(),
    limits = c(0, exp6_reads_fig_y_upper_no0),
    breaks = exp6_reads_y_breaks_use_no0,
    labels = exp6_reads_y_labels,
    minor_breaks = NULL,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  exp6_reads_plot_theme_no0 +
  ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
  ggplot2::guides(
    color = ggplot2::guide_legend(
      title = "GVB quantity",
      reverse = TRUE
    )
  )

if (exp6_reads_no0_q1_ok) {
  n_q1n0 <- length(exp6_reads_no0_q1_time_levels_plot)
  ce_time_no0_q1 <- conditional_effects(exp6_reads_no0_q1_plot_mod, effects = "time")[[1]] %>%
    dplyr::mutate(
      hrs = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = match(.data$hrs, exp6_reads_no0_q1_time_levels_plot)
    )

  mx_bump <- max(
    0,
    suppressWarnings(max(as.numeric(ce_time_no0_q1$upper__), na.rm = TRUE)),
    suppressWarnings(max(df_exp6_reads_no0_q1_mod$gvb_sequence_number, na.rm = TRUE)),
    na.rm = TRUE
  )
  exp6_reads_fig_y_upper_no0 <- max(exp6_reads_fig_y_upper_no0, mx_bump * 1.12, 10)
  exp6_reads_y_breaks_use_no0 <- exp6_reads_y_breaks[exp6_reads_y_breaks <= exp6_reads_fig_y_upper_no0]

  raw_col_no0_q1 <- df_exp6_reads_no0_q1_mod %>%
    dplyr::mutate(
      hrs = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = match(.data$hrs, exp6_reads_no0_q1_time_levels_plot) + exp6_reads_fig_x_nudge_raw
    )

  y_upper_no0_q1 <- max(
    exp6_reads_fig_y_upper_no0,
    suppressWarnings(max(as.numeric(ce_time_no0_q1$upper__), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(ce_time_no0_q1$estimate__), na.rm = TRUE)),
    suppressWarnings(max(raw_col_no0_q1$gvb_sequence_number, na.rm = TRUE)),
    10,
    na.rm = TRUE
  )
  y_breaks_no0_q1 <- exp6_reads_y_breaks[exp6_reads_y_breaks <= y_upper_no0_q1]

  exp6_reads_plot_theme_no0_q1 <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
    )

  col_mu_no0_q1 <- "#0072B2"
  p_exp6_reads_no0_q1_eff <- ggplot2::ggplot(
    ce_time_no0_q1,
    ggplot2::aes(x = .data$xtime, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
  ) +
    ggplot2::geom_point(
      data = raw_col_no0_q1,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number),
      inherit.aes = FALSE,
      position = ggplot2::position_jitter(width = 0.1, height = 0.175),
      alpha = 0.4,
      size = 0.85,
      color = "grey40",
      stroke = 0
    ) +
    ggplot2::geom_errorbar(width = 0.12, linewidth = 0.9, color = col_mu_no0_q1) +
    ggplot2::geom_point(size = 2.5, alpha = 0.9, color = col_mu_no0_q1) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_q1n0),
      labels = as.character(exp6_reads_no0_q1_time_levels_plot),
      limits = exp6_reads_xlim_for_n(n_q1n0),
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(0, y_upper_no0_q1),
      breaks = y_breaks_no0_q1,
      labels = exp6_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp6_reads_plot_theme_no0_q1 +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

  p_exp6_reads_no0_left <- p_exp6_reads_no0_eff +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(0, exp6_reads_fig_y_upper_no0),
      breaks = exp6_reads_y_breaks_use_no0,
      labels = exp6_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    )

  p_exp6_reads_no0_eff_combo <- patchwork::wrap_plots(
    p_exp6_reads_no0_left +
      exp6_reads_combo_legend_theme +
      exp6_reads_combo_tag_theme +
      ggplot2::labs(tag = "a)"),
    p_exp6_reads_no0_q1_eff +
      ggplot2::labs(
        caption = "Quantity GVB = 1.",
        tag = "b)"
      ) +
      ggplot2::theme(
        plot.caption = ggplot2::element_text(size = 8.5, colour = "grey25", hjust = 0),
        plot.caption.position = "panel"
      ) +
      exp6_reads_combo_tag_theme,
    ncol = 2
  )
} else {
  p_exp6_reads_no0_q1_eff <- NULL
  p_exp6_reads_no0_eff_combo <- patchwork::wrap_plots(
    p_exp6_reads_no0_eff +
      exp6_reads_combo_legend_theme +
      exp6_reads_combo_tag_theme +
      ggplot2::labs(tag = "a)"),
    ncol = 1
  )
}

ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-04-exp6-reads-no0-panel-a-only.png"),
  plot = p_exp6_reads_no0_eff,
  width = 7,
  height = 4.5,
  dpi = 600
)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-04-exp6-reads-no0-panels-a-b.png"),
  plot = p_exp6_reads_no0_eff_combo,
  width = 14,
  height = 4.5,
  dpi = 600
)

p_exp6_reads_no0_eff_combo
```

Figure 4: GVB sequence counts from the reads workbook with 0 h exposure excluded from **panel a** (no–time–0 workbook). **a)** 2 h and 24 h only — mean sequence count by GVB quantity. **b)** Quantity GVB = 1 using the **full** reads workbook so **0 h** is included when quantity-1 rows exist at that exposure — marginal effect of exposure time on mean count. Vertical axis is log(1 + count). Jittered points are raw counts; error bars are 95% credible intervals and coloured points are posterior means.

#### Tables

Tabulated estimates use the **preferred monotonic interaction** model **`mo(time) * mo(quantity_gbv)`** (negative binomial mean count with **`offset(log(total_reads))`**; marginal **`conditional_effects()`**). See the **Model comparison** tab for **WAIC** across all candidate models.

```{r exp6-reads-no0-estimates, echo=FALSE}
ce_exp6_no0_t <- conditional_effects(mod_12, effects = "time")[[1]] %>%
  dplyr::mutate(.thr = suppressWarnings(as.numeric(as.character(.data$time)))) %>%
  dplyr::filter(.data$.thr %in% c(2, 24)) %>%
  dplyr::transmute(
    `Time (hrs)` = time,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

ce_exp6_no0_q <- conditional_effects(mod_12, effects = "quantity_gbv")[[1]] %>%
  dplyr::transmute(
    `Quantity GVB` = quantity_gbv,
    `Estimate` = round(estimate__, 3),
    `Lower` = round(lower__, 3),
    `Upper` = round(upper__, 3)
  )

knitr::kable(
  ce_exp6_no0_t,
  caption = "Experiment 6 reads (no time 0) — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
  format = "html"
)

knitr::kable(
  ce_exp6_no0_q,
  caption = "Experiment 6 reads (no time 0) — best / primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
  format = "html"
)
```

**Quantity GVB = 1 — full reads workbook (same time levels as Figure 3b, including **0 h** when observed at quantity 1)** — **Figure 4b** and the table below use **`mo(time)`** (monotonic time). **Model comparison** lists **WAIC**; the **lower-WAIC** q1 parameterization is **`r if (nzchar(exp6_reads_no0_q1_waic_preferred_label)) exp6_reads_no0_q1_waic_preferred_label else "—"`** when fitted.

```{r exp6-reads-no0-estimates-q1, echo=FALSE}
if (exp6_reads_no0_q1_ok) {
  ce_exp6_no0_q1_t <- conditional_effects(exp6_reads_no0_q1_plot_mod, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp6_no0_q1_t,
    caption = paste0(
      "Experiment 6 reads — panel a uses no–time–0 workbook; quantity GVB = 1 (0 h recoded from quantity 0; full workbook): ",
      exp6_reads_no0_q1_plot_label,
      " (same as Figure 4b; WAIC in Model comparison) — conditional effect of time (mean count)."
    ),
    format = "html"
  )
} else {
  knitr::kable(
    tibble::tibble(Note = "Not fitted: need \u2265 1 row with quantity_gbv = 1 and \u2265 2 distinct exposure times after filtering."),
    caption = "Experiment 6 reads (no time 0) — quantity GVB = 1 (all times) — no table.",
    format = "html"
  )
}
```

#### Models

```{r exp6-reads-no0-models-show, ref.label="exp6-reads-no0-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp6-reads-no0-waic-tab, echo=FALSE}
knitr::kable(
  exp6_reads_no0_waic_tbl,
  caption = "Experiment 6 reads (no time 0): WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
if (exp6_reads_no0_q1_ok) {
  knitr::kable(
    exp6_reads_no0_q1_waic_tbl,
    caption = "Experiment 6 reads — **quantity GVB = 1** (full workbook; Figure 4b): WAIC for `mo(time)` vs factor `time` (offset log(total reads); distinct from main table above).",
    format = "html"
  )
}
```

#### Validation

```{r exp6-reads-no0-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_11, ndraws = 200)
pp_check(mod_12, ndraws = 200)
pp_check(mod_13, ndraws = 200)
pp_check(mod_14, ndraws = 200)
```

**Quantity GVB = 1 (all exposure times)** — posterior predictive checks.

```{r exp6-reads-no0-pp-check-q1}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
if (exp6_reads_no0_q1_ok) {
  pp_check(mod_15, ndraws = 200)
  pp_check(mod_16, ndraws = 200)
}
```

#### Output

::: {.panel-tabset}

##### mod_11

```{r exp6-reads-no0-summary-base, echo=FALSE}
#| class-output: model-output-panel
summary(mod_11)
```

##### mod_12

```{r exp6-reads-no0-summary-mo-int, echo=FALSE}
#| class-output: model-output-panel
summary(mod_12)
```

##### mod_13

```{r exp6-reads-no0-summary-fac-add, echo=FALSE}
#| class-output: model-output-panel
summary(mod_13)
```

##### mod_14

```{r exp6-reads-no0-summary-fac-int, echo=FALSE}
#| class-output: model-output-panel
summary(mod_14)
```

##### Quantity GVB = 1 (q1)

```{r exp6-reads-no0-summary-q1, echo=FALSE}
#| class-output: model-output-panel
if (exp6_reads_no0_q1_ok) {
  summary(mod_15)
  summary(mod_16)
}
```

:::

:::

## Experiment 10

```{r exp10-wrangle, include=FALSE}
# Shared hour scale for detection (`df.10`) and reads (`df.10.reads`).
# Reads sheets label exposure via `time` (or `exposure_time_hours` → `time` in `wrangle_exp6_reads_sheet()`).
# Before any model fit, strings are converted to numeric hours; `df.10.reads` then sets
# `time_hours` and rebuilds `time` as an ordered factor on that **hour** scale (for `mo(time)`).
# Typical Experiment 10 workbook labels (after `clean_names`, values may be character or numeric):
#   "16 hours" → 16    "1 day" → 24    "2 days" → 48    "3 days" → 72    "0" → 0
# Short forms (`16hrs`, `1d`, …) and regex patterns for `<n> hours` / `<n> days` also work.
exp10_time_to_hours <- function(x) {
  # Excel often inserts NBSP / narrow no-break space; base trimws + \\s may miss those.
  xc <- stringr::str_replace_all(stringr::str_trim(as.character(x)), "\\s+", "")
  xc <- tolower(xc)
  dplyr::case_when(
    xc == "" | is.na(xc) ~ NA_real_,
    xc == "0" ~ 0,
    xc %in% c("16hrs", "16hours", "16hour") ~ 16,
    xc %in% c("1d", "1day") ~ 24,
    xc %in% c("2d", "2day", "2days") ~ 48,
    xc %in% c("3d", "3day", "3days") ~ 72,
    xc %in% c("7d", "7day", "7days") ~ 168,
    grepl("^\\d+(\\.\\d+)?hours?$", xc) ~ suppressWarnings(
      as.numeric(sub("hours?$", "", xc, perl = TRUE))
    ),
    grepl("^\\d+(\\.\\d+)?days?$", xc) ~ suppressWarnings(
      as.numeric(sub("days?$", "", xc, perl = TRUE)) * 24
    ),
    grepl("^\\d+(\\.\\d+)?$", xc) ~ suppressWarnings(as.numeric(xc)),
    TRUE ~ NA_real_
  )
}

df.10 <- read_excel(here::here("analysis/data/Exp10- no cut off.xlsx"), sheet = 1) %>% clean_names()

cols <- c("sample", "rep", "sample_id")
df.10[, cols] <- lapply(df.10[, cols], as.factor)

df.10 <- df.10 %>%
  dplyr::mutate(positive_cut = ifelse(cq > 39, 0, positive)) %>%
  dplyr::mutate(time_hours = exp10_time_to_hours(.data$time)) %>%
  dplyr::mutate(
    time = ordered(.data$time_hours, levels = sort(unique(stats::na.omit(.data$time_hours))))
  ) %>%
  dplyr::select(-dplyr::all_of("time_hours"))

df.10.reads <- wrangle_exp6_reads_sheet(
  "analysis/data/GVB_exp10_reads_tom.xlsx",
  require_quantity_gbv = FALSE,
  require_total_reads = TRUE
) %>%
  dplyr::mutate(time_hours = exp10_time_to_hours(.data$exposure_label)) %>%
  dplyr::mutate(
    time = ordered(.data$time_hours, levels = sort(unique(stats::na.omit(.data$time_hours))))
  )

exp10_reads_no0_rel <- if (file.exists(here::here("analysis/data/GVB_exp10_reads_tom_no_time_0.xlsx"))) {
  "analysis/data/GVB_exp10_reads_tom_no_time_0.xlsx"
} else if (file.exists(here::here("analysis/data/GVB_exp10_reads_tom_no_time0.xlsx"))) {
  "analysis/data/GVB_exp10_reads_tom_no_time0.xlsx"
} else {
  NA_character_
}
df.10.reads.no0 <- if (!is.na(exp10_reads_no0_rel)) {
  wrangle_exp6_reads_sheet(
    exp10_reads_no0_rel,
    require_quantity_gbv = FALSE,
    require_total_reads = TRUE
  ) %>%
    dplyr::mutate(time_hours = exp10_time_to_hours(.data$exposure_label)) %>%
    dplyr::mutate(
      time = ordered(.data$time_hours, levels = sort(unique(stats::na.omit(.data$time_hours))))
    )
} else {
  df.10.reads %>%
    dplyr::filter(!is.na(.data$time_hours), .data$time_hours > 0)
}

# TRUE when sequence-count workbooks include a mappable quantity column (Experiment-6-style models).
exp10_reads_has_quantity <- "quantity_gbv" %in% names(df.10.reads)

# Shared row filter for Exp 10 negative-binomial reads (matches model fitting chunks).
# Exclude low library sizes (e.g. sample NA-5) — same floor as Exp 6 reads models.
# If `quantity_gbv` is absent, rows are not filtered on quantity (time-only models).
exp10_reads_filtered_for_nb <- function(dr) {
  dr <- dplyr::ungroup(dr)
  has_q <- "quantity_gbv" %in% names(dr)
  has_th <- "time_hours" %in% names(dr)
  has_el <- "exposure_label" %in% names(dr)
  d <- dr %>%
    dplyr::mutate(
      gvb_sequence_number = suppressWarnings(as.integer(round(as.numeric(.data$gvb_sequence_number)))),
      total_reads = as.numeric(.data$total_reads),
      tnum = as.integer(round(dplyr::coalesce(
        if (has_th) suppressWarnings(as.numeric(.data$time_hours)) else NA_real_,
        if (has_el) exp10_time_to_hours(.data$exposure_label) else NA_real_,
        exp10_time_to_hours(.data$time),
        suppressWarnings(as.numeric(as.character(.data$time)))
      )))
    )
  if (has_q) {
    d <- d %>% dplyr::mutate(quantity_gbv = as.numeric(.data$quantity_gbv))
  }
  d <- d %>%
    dplyr::filter(
      !is.na(.data$gvb_sequence_number),
      !is.na(.data$time),
      !is.na(.data$total_reads),
      .data$total_reads >= 10000,
      .data$gvb_sequence_number >= 0L
    )
  if (has_q) {
    d <- d %>% dplyr::filter(!is.na(.data$quantity_gbv), .data$quantity_gbv != 0)
  }
  d
}

exp10_reads_fig_y_max_filtered <- function(dr) {
  if (!nrow(dr)) {
    return(0)
  }
  exp10_reads_filtered_for_nb(dr) %>%
    dplyr::summarise(mx = max(c(0L, .data$gvb_sequence_number), na.rm = TRUE), .groups = "drop") %>%
    dplyr::pull(1)
}
```

```{r exp10-reads-fig34-prep, include=FALSE}
exp10_reads_fig_y_upper <- max(
  exp10_reads_fig_y_max_filtered(df.10.reads),
  exp10_reads_fig_y_max_filtered(df.10.reads.no0),
  na.rm = TRUE
) * 1.45
exp10_reads_fig_y_upper <- max(exp10_reads_fig_y_upper, 10)
exp10_reads_y_breaks <- sort(unique(c(0, 1, 10, 100, 1000, 10000, 100000)))
exp10_reads_y_breaks_use <- exp10_reads_y_breaks[exp10_reads_y_breaks <= exp10_reads_fig_y_upper]
exp10_reads_y_labels <- function(x) ifelse(x == 0, "0", scales::label_number(big.mark = ",")(x))
exp10_reads_fig_x_nudge_raw <- 0.1
exp10_reads_fig_y_lower_f3 <- -0.18
exp10_reads_xlim_for_n <- function(n) c(0.45, n + 0.27)
exp10_reads_qty_palette <- exp6_reads_qty_palette
```

<details><summary>Data frames</summary>

```{r exp10-summary-tibbles, echo=FALSE}
cat("Experiment 10 detection (Exp10- no cut off.xlsx)\n\n")
brief_dataframe(dplyr::count(df.10, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.10)

cat("\n\nReads — GVB_exp10_reads_tom.xlsx\n\n")
brief_dataframe(dplyr::count(df.10.reads, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.10.reads)

cat("\n\nReads — no time 0 (GVB_exp10_reads_tom_no_time_0.xlsx)\n\n")
brief_dataframe(dplyr::count(df.10.reads.no0, sample, time, name = "n"), "Counts by sample × time")
brief_dataframe(df.10.reads.no0)
```

</details>

::: {.panel-tabset}

```{r exp10-models-fit, include=FALSE}
mod_formula <- bf(positive ~ mo(time) + (1|sample_id))

mod_17 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  data = df.10,
  file = here::here("outputs/models", "mod_17"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)

mod_formula <- bf(positive_cut ~ mo(time) + (1|sample_id))

mod_18 <- brm(
  mod_formula, family = bernoulli(link = "logit"),
  data = df.10,
  file = here::here("outputs/models", "mod_18"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  control = list(step_size = 0.01, adapt_delta = 0.99),
  backend = "cmdstanr", silent = 0, refresh = 1000,
  chains = 4, cores = 4, warmup = 2000, iter = 4000
)
```

```{r exp10-waic-compute, include=FALSE}
waic_tbl_one_exp10 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

exp10_waic_tbl <- dplyr::bind_rows(
  waic_tbl_one_exp10(mod_17, "mod_17 (time monotonic, no cut)"),
  waic_tbl_one_exp10(mod_18, "mod_18 (time monotonic, cut)")
) %>%
  dplyr::arrange(WAIC, Model) %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)
```

## Effect plots

```{r exp10-eff, warning=FALSE, echo=FALSE, fig.width=7, fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
c_eff_5 <- conditional_effects(mod_17)[[1]] %>% mutate(cutoff = "No")
c_eff_6 <- conditional_effects(mod_18)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
c_eff <- rbind(c_eff_5, c_eff_6)

raw_exp10_eff <- dplyr::bind_rows(
  dplyr::mutate(df.10, cutoff = factor("No", levels = c("No", "Yes")), y = positive),
  dplyr::mutate(df.10, cutoff = factor("Yes", levels = c("No", "Yes")), y = positive_cut)
) %>%
  dplyr::mutate(y_plot = binary_y_for_plot(.data$y))

y_lim_exp10_eff1 <- c(-0.13, 1.13)
tlv_exp10_eff1 <- levels(c_eff$time)
n_t_exp10_eff1 <- length(tlv_exp10_eff1)
ichr_exp10 <- as.character(tlv_exp10_eff1)
i72_exp10_eff1 <- which(ichr_exp10 == "72")[1L]
if (is.na(i72_exp10_eff1)) {
  i72_exp10_eff1 <- which(suppressWarnings(as.numeric(ichr_exp10)) == 72)[1L]
}
if (is.na(i72_exp10_eff1)) i72_exp10_eff1 <- max(1L, round(n_t_exp10_eff1 / 2))
x_npc_exp10_eff1 <- (i72_exp10_eff1 - 0.5) / n_t_exp10_eff1
y_npc_exp10_eff1 <- (0.3 - y_lim_exp10_eff1[1]) / (y_lim_exp10_eff1[2] - y_lim_exp10_eff1[1])

p_exp10_eff <- c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
  geom_jitter(
    data = raw_exp10_eff,
    aes(x = time, y = y_plot, group = cutoff, color = cutoff),
    inherit.aes = FALSE,
    position = position_jitterdodge(jitter.width = 0.12, jitter.height = 0.03, dodge.width = -0.5),
    alpha = 0.35,
    size = 0.85,
    stroke = 0
  ) +
  geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
  geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
  theme_bw() +
  theme(
    legend.position = c(x_npc_exp10_eff1, y_npc_exp10_eff1),
    legend.justification = c(0.5, 0.5),
    legend.margin = ggplot2::margin(4, 4, 4, 4),
    legend.background = element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
  ) +
  scale_y_continuous(
    labels = scales::percent,
    limits = y_lim_exp10_eff1,
    breaks = prob_det_y_breaks,
    expand = ggplot2::expansion(mult = c(0, 0))
  ) +
  scale_color_manual(values = c("darkblue", "darkgreen")) +
  labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
  guides(color = guide_legend(title = "Cq cut off (> 39)", reverse = TRUE))

ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-05-exp10-detection-mo-time-cutoff.png"),
  plot = p_exp10_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_eff
```

Figure 5: Conditional effect plot of GBV exposure on the probability of detection with (yes = green) and without (no = blue) a Cq cut off of 39. Jittered points are raw binary outcomes by exposure time, coloured by cutoff; error bars are 95% credible intervals and coloured points are posterior means.

## Tables

```{r exp10-estimates, echo=FALSE}
# Modeled estimates from conditional effects() on the probability scale.

ce10_time <- dplyr::bind_rows(
  conditional_effects(mod_17)[[1]] %>%
    mutate(cutoff = "No") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    ),
  conditional_effects(mod_18)[[1]] %>%
    mutate(cutoff = "Yes") %>%
    transmute(
      Cutoff = cutoff,
      `Time (hrs)` = time,
      `Estimate (%)` = round(estimate__ * 100, 2),
      `Lower (%)` = round(lower__ * 100, 2),
      `Upper (%)` = round(upper__ * 100, 2)
    )
) %>%
  arrange(Cutoff, `Time (hrs)`)

knitr::kable(
  ce10_time,
  caption = "Modeled estimates from conditional effects (Experiment 10): time",
  format = "html"
)
```

## Model comparison

```{r exp10-waic-tab, echo=FALSE}
knitr::kable(
  exp10_waic_tbl,
  caption = "Experiment 10: WAIC (lower is better; `mo(time)` formulations).",
  format = "html"
)
```

## Models

```{r exp10-models-show, ref.label="exp10-models-fit", echo=TRUE, eval=FALSE}

```

## Validation

```{r exp10-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_17, ndraws = 200, col = 2)
pp_check(mod_18, ndraws = 200, col = 2)
```

## Output

::: {.panel-tabset}

### mod_17

```{r exp10-summary-mod5, echo=FALSE}
#| class-output: model-output-panel
summary(mod_17)
```

### mod_18

```{r exp10-summary-mod6, echo=FALSE}
#| class-output: model-output-panel
summary(mod_18)
```

:::

<!-- Pairwise comparisons removed -->

```{r exp10-pairwise, include=FALSE, eval=FALSE}

pairwise_time <- dplyr::bind_rows(
  {
    emm <- emmeans::emmeans(mod_17, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "No"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  },
  {
    emm <- emmeans::emmeans(mod_18, ~ time, type = "response")
    pr <- emmeans::contrast(emm, method = "pairwise", adjust = "tukey")
    d <- as.data.frame(summary(pr))
    d$Cutoff <- "Yes"
    est_col <- if ("estimate" %in% names(d)) "estimate" else if ("diff" %in% names(d)) "diff" else if ("difference" %in% names(d)) "difference" else {
      num_cols <- names(d)[sapply(d, is.numeric)]
      num_cols <- setdiff(num_cols, c("SE", "SE.difference", "p.value"))
      if (length(num_cols) < 1) stop("Pairwise table: could not infer numeric estimate/difference column.")
      num_cols[1]
    }
    se_col <- if ("SE" %in% names(d)) "SE" else if ("SE.difference" %in% names(d)) "SE.difference" else {
      se_like <- names(d)[grepl("^SE$|se\\b|SE\\.|Std\\.Error|std\\.error", names(d), ignore.case = TRUE)]
      se_like <- intersect(se_like, names(d)[sapply(d, is.numeric)])
      if (length(se_like) < 1) NA_character_ else se_like[1]
    }
    p_col <- if ("p.value" %in% names(d)) "p.value" else {
      p_like <- names(d)[grepl("^p\\.value$|p\\.val|pvalue|p\\.adj|adj\\.p|p$", names(d), ignore.case = TRUE)]
      p_like <- intersect(p_like, names(d)[sapply(d, is.numeric)])
      if (length(p_like) < 1) NA_character_ else p_like[1]
    }
    d %>%
      mutate(
        `Difference (pp)` = round(.data[[est_col]] * 100, 2),
        `SE (pp)` = if (!is.na(se_col)) round(.data[[se_col]] * 100, 2) else NA_real_,
        `p-value` = if (!is.na(p_col)) round(.data[[p_col]], 2) else NA_real_
      ) %>%
      transmute(Cutoff, Contrast = contrast, `Difference (pp)`, `SE (pp)`, `p-value`)
  }
)

knitr::kable(
  pairwise_time,
  caption = "Pairwise comparisons between time levels (Experiment 10; Tukey-adjusted): response scale (pp)",
  format = "html"
)
```

:::

### GVB sequence counts - reads

::: {.panel-tabset}

```{r exp10-reads-models-fit, include=FALSE}
if (exp10_reads_has_quantity) {
  d10r <- exp10_reads_filtered_for_nb(df.10.reads)
  t10_lv <- sort(unique(stats::na.omit(d10r$tnum)))
  if (length(t10_lv) < 2L) {
    uel <- if ("exposure_label" %in% names(df.10.reads)) {
      paste(utils::head(unique(df.10.reads$exposure_label), 20L), collapse = " | ")
    } else {
      "(no exposure_label column)"
    }
    stop(
      "Experiment 10 reads: need ≥ 2 time levels after filtering (quantity GVB, total reads). Found: ",
      paste(t10_lv, collapse = ", "),
      " (nrow = ", nrow(d10r), "). Unique exposure_label (raw, up to 20): ", uel,
      ". Check Exposure column mapping, Unicode spaces, and filters on total_reads / gvb_sequence_number."
    )
  }
  qty10_lv <- sort(unique(stats::na.omit(d10r$quantity_gbv)))
  df_exp10_reads_mod <- d10r %>%
    dplyr::mutate(
      time = ordered(.data$tnum, levels = t10_lv),
      quantity_gbv = ordered(.data$quantity_gbv, levels = qty10_lv)
    ) %>%
    dplyr::select(-tnum)

  # Do not use as.numeric(time) on ordered factors: that returns level indices (1,2,…) not hour values.
  df_exp10_reads_fac <- d10r %>%
    dplyr::mutate(
      time = factor(as.character(.data$tnum), levels = as.character(t10_lv)),
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty10_lv))
    ) %>%
    dplyr::select(-tnum)

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

  mod_19 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_mod,
    file = here::here("outputs/models", "mod_19"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

  mod_20 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_mod,
    file = here::here("outputs/models", "mod_20"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

  mod_21 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_fac,
    file = here::here("outputs/models", "mod_21"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

  mod_22 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_fac,
    file = here::here("outputs/models", "mod_22"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )
} else {
  d10r <- exp10_reads_filtered_for_nb(df.10.reads)
  t10_lv <- sort(unique(stats::na.omit(d10r$tnum)))
  if (length(t10_lv) < 2L) {
    uel <- if ("exposure_label" %in% names(df.10.reads)) {
      paste(utils::head(unique(df.10.reads$exposure_label), 20L), collapse = " | ")
    } else {
      "(no exposure_label column)"
    }
    stop(
      "Experiment 10 reads: need ≥ 2 distinct time levels after filtering (total reads, counts). Found: ",
      paste(t10_lv, collapse = ", "),
      " (nrow = ", nrow(d10r), "). Unique exposure_label (raw, up to 20): ", uel,
      ". Check Exposure column mapping in `wrangle_exp6_reads_sheet()`, Unicode spaces in cells, ",
      "and filters on total_reads / gvb_sequence_number."
    )
  }
  df_exp10_reads_mod <- d10r %>%
    dplyr::mutate(time = ordered(.data$tnum, levels = t10_lv)) %>%
    dplyr::select(-tnum)

  df_exp10_reads_fac <- d10r %>%
    dplyr::mutate(time = factor(as.character(.data$tnum), levels = as.character(t10_lv))) %>%
    dplyr::select(-tnum)

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + offset(log(total_reads)))

  mod_19 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_mod,
    file = here::here("outputs/models", "mod_19"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_20 <- mod_19

  mod_formula <- bf(gvb_sequence_number ~ time + offset(log(total_reads)))

  mod_21 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_fac,
    file = here::here("outputs/models", "mod_21"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_22 <- mod_21
}
```

```{r exp10-reads-waic-compute, include=FALSE}
waic_tbl_one_exp10_reads <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

if (exp10_reads_has_quantity) {
  exp10_reads_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads(mod_19, "mo(time) + mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads(mod_20, "mo(time) * mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads(mod_21, "time + quantity_gbv (factor)"),
    waic_tbl_one_exp10_reads(mod_22, "time * quantity_gbv (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
} else {
  exp10_reads_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads(mod_19, "mo(time)"),
    waic_tbl_one_exp10_reads(mod_21, "time (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
}
```

#### Effect plots

```{r exp10-reads-eff, warning=FALSE, echo=FALSE, fig.width=8, fig.height=4.5, out.width="100%", dpi=600}
# Figure 6: factor time (and factor quantity when applicable) — matches mod_21 / mod_22
raw <- df_exp10_reads_fac
n_t_exp10 <- length(t10_lv)
xlim_exp10 <- exp10_reads_xlim_for_n(n_t_exp10)
x_lab_exp10 <- as.character(t10_lv)
leg_npc_exp10_f3 <- c(
  x = (n_t_exp10 - xlim_exp10[1]) / diff(xlim_exp10),
  y = (log1p(10) - log1p(exp10_reads_fig_y_lower_f3)) /
    (log1p(exp10_reads_fig_y_upper) - log1p(exp10_reads_fig_y_lower_f3))
)

if (exp10_reads_has_quantity) {
  ce_both <- conditional_effects(mod_22, effects = "time:quantity_gbv")[[1]]
  qty_lv <- sort(unique(as.numeric(as.character(ce_both$quantity_gbv))))
  ce_both <- ce_both %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv))
    ) %>%
    dplyr::mutate(
      tord = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = as.numeric(.data$time)
    ) %>%
    dplyr::arrange(.data$quantity_gbv, .data$tord)

  raw_col <- raw %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both$quantity_gbv)),
      xtime = as.numeric(.data$time) + exp10_reads_fig_x_nudge_raw
    )

  n_qty <- dplyr::n_distinct(ce_both$quantity_gbv)
  pal <- exp10_reads_qty_palette(n_qty)
  ce_pd <- ggplot2::position_dodge(width = 0.85)

  exp10_reads_plot_theme <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
      legend.position = unname(leg_npc_exp10_f3),
      legend.justification = c(0.5, 0.5),
      legend.title = ggplot2::element_text(size = 10),
      legend.text = ggplot2::element_text(size = 9),
      legend.margin = ggplot2::margin(2, 4, 2, 4),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
    )

  p_exp10_reads_eff <- ggplot2::ggplot(
    ce_both,
    ggplot2::aes(
      x = .data$xtime,
      y = .data$estimate__,
      ymin = .data$lower__,
      ymax = .data$upper__,
      color = .data$quantity_gbv
    )
  ) +
    ggplot2::geom_point(
      data = raw_col,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
      inherit.aes = FALSE,
      position = ggplot2::position_jitterdodge(jitter.width = 0.12, jitter.height = 0.175, dodge.width = 0.85),
      alpha = 0.4,
      size = 0.85,
      stroke = 0
    ) +
    ggplot2::geom_errorbar(position = ce_pd, width = 0.12, linewidth = 0.9) +
    ggplot2::geom_point(position = ce_pd, size = 2.5, alpha = 0.9) +
    ggplot2::scale_color_manual(values = pal) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10),
      labels = x_lab_exp10,
      limits = xlim_exp10,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(exp10_reads_fig_y_lower_f3, exp10_reads_fig_y_upper),
      breaks = exp10_reads_y_breaks_use,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
    ggplot2::guides(color = ggplot2::guide_legend(title = "GVB quantity", reverse = TRUE))

  fn_exp10_reads_plot <- here::here("outputs/plots", "figure-06-exp10-reads-time-by-quantity.png")
} else {
  ce_time <- conditional_effects(mod_21, effects = "time")[[1]] %>%
    dplyr::mutate(xtime = as.numeric(.data$time))

  raw_col <- raw %>%
    dplyr::mutate(xtime = as.numeric(.data$time) + exp10_reads_fig_x_nudge_raw)

  exp10_reads_plot_theme <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
    )

  col_mu <- "#0072B2"
  p_exp10_reads_eff <- ggplot2::ggplot(
    ce_time,
    ggplot2::aes(x = .data$xtime, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
  ) +
    ggplot2::geom_point(
      data = raw_col,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number),
      inherit.aes = FALSE,
      position = ggplot2::position_jitter(width = 0.1, height = 0.175),
      alpha = 0.4,
      size = 0.85,
      color = "grey40",
      stroke = 0
    ) +
    ggplot2::geom_errorbar(width = 0.12, linewidth = 0.9, color = col_mu) +
    ggplot2::geom_point(size = 2.5, alpha = 0.9, color = col_mu) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10),
      labels = x_lab_exp10,
      limits = xlim_exp10,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(exp10_reads_fig_y_lower_f3, exp10_reads_fig_y_upper),
      breaks = exp10_reads_y_breaks_use,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

  fn_exp10_reads_plot <- here::here("outputs/plots", "figure-06-exp10-reads-time-only.png")
}

ggplot2::ggsave(
  filename = fn_exp10_reads_plot,
  plot = p_exp10_reads_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_reads_eff
```

```{r exp10-reads-fig6-caption, echo=FALSE, results="asis"}
if (exp10_reads_has_quantity) {
  cat("Figure 6: GVB sequence counts from the reads workbook. Conditional effect of GVB exposure on mean sequence count by GVB quantity (**factor** exposure time and quantity; interaction model). Vertical axis is log(1 + count). Jittered points are raw counts; error bars are 95% credible intervals and coloured points are posterior means.\n\n")
} else {
  cat("Figure 6: GVB sequence counts from the reads workbook. Conditional effect of GVB exposure on mean sequence count (**factor** exposure time). Vertical axis is log(1 + count). Grey jittered points are raw counts; blue error bars and points are 95% credible intervals and posterior means.\n\n")
}
```

#### Tables

```{r exp10-reads-estimates, echo=FALSE}
if (exp10_reads_has_quantity) {
  ce_exp10_t <- conditional_effects(mod_22, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  ce_exp10_q <- conditional_effects(mod_22, effects = "quantity_gbv")[[1]] %>%
    dplyr::transmute(
      `Quantity GVB` = quantity_gbv,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_t,
    caption = "Experiment 10 reads — factor model `time * quantity_gbv`: marginal conditional effect of time (mean count).",
    format = "html"
  )

  knitr::kable(
    ce_exp10_q,
    caption = "Experiment 10 reads — factor model `time * quantity_gbv`: marginal conditional effect of quantity GVB (mean count).",
    format = "html"
  )
} else {
  ce_exp10_t <- conditional_effects(mod_21, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_t,
    caption = "Experiment 10 reads — factor model `time` with offset: conditional effect of time (mean count).",
    format = "html"
  )
}
```

#### Models

```{r exp10-reads-models-show, ref.label="exp10-reads-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp10-reads-waic-tab, echo=FALSE}
knitr::kable(
  exp10_reads_waic_tbl,
  caption = "Experiment 10 reads: WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
```

#### Validation

```{r exp10-reads-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
if (exp10_reads_has_quantity) {
  pp_check(mod_19, ndraws = 200)
  pp_check(mod_20, ndraws = 200)
  pp_check(mod_21, ndraws = 200)
  pp_check(mod_22, ndraws = 200)
} else {
  pp_check(mod_19, ndraws = 200)
  pp_check(mod_21, ndraws = 200)
}
```

#### Output

::: {.panel-tabset}

##### mod_19

```{r exp10-reads-summary-base, echo=FALSE}
#| class-output: model-output-panel
summary(mod_19)
```

##### mod_20

```{r exp10-reads-summary-mo-int, echo=FALSE}
#| class-output: model-output-panel
if (exp10_reads_has_quantity) summary(mod_20)
```

##### mod_21

```{r exp10-reads-summary-fac-add, echo=FALSE}
#| class-output: model-output-panel
summary(mod_21)
```

##### mod_22

```{r exp10-reads-summary-fac-int, echo=FALSE}
#| class-output: model-output-panel
if (exp10_reads_has_quantity) summary(mod_22)
```

:::

:::

### GVB sequence counts - no time 0

::: {.panel-tabset}

```{r exp10-reads-no0-models-fit, include=FALSE}
if (exp10_reads_has_quantity) {
  d10rn0 <- exp10_reads_filtered_for_nb(df.10.reads.no0)
  t10n0_lv <- sort(unique(stats::na.omit(d10rn0$tnum)))
  if (length(t10n0_lv) < 2L) {
    uel_n0 <- if ("exposure_label" %in% names(df.10.reads.no0)) {
      paste(utils::head(unique(df.10.reads.no0$exposure_label), 20L), collapse = " | ")
    } else {
      "(no exposure_label column)"
    }
    stop(
      "Experiment 10 reads (no time 0): need ≥ 2 time levels after filtering. Found: ",
      paste(t10n0_lv, collapse = ", "),
      " (nrow = ", nrow(d10rn0), "). Unique exposure_label (raw, up to 20): ", uel_n0, "."
    )
  }
  qty10n0_lv <- sort(unique(stats::na.omit(d10rn0$quantity_gbv)))
  df_exp10_reads_no0_mod <- d10rn0 %>%
    dplyr::mutate(
      time = ordered(.data$tnum, levels = t10n0_lv),
      quantity_gbv = ordered(.data$quantity_gbv, levels = qty10n0_lv)
    ) %>%
    dplyr::select(-tnum)

  df_exp10_reads_no0_fac <- d10rn0 %>%
    dplyr::mutate(
      time = factor(as.character(.data$tnum), levels = as.character(t10n0_lv)),
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty10n0_lv))
    ) %>%
    dplyr::select(-tnum)

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + mo(quantity_gbv) + offset(log(total_reads)))

  mod_23 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_mod,
    file = here::here("outputs/models", "mod_23"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ mo(time) * mo(quantity_gbv) + offset(log(total_reads)))

  mod_24 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_mod,
    file = here::here("outputs/models", "mod_24"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time + quantity_gbv + offset(log(total_reads)))

  mod_25 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_fac,
    file = here::here("outputs/models", "mod_25"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_formula <- bf(gvb_sequence_number ~ time * quantity_gbv + offset(log(total_reads)))

  mod_26 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_fac,
    file = here::here("outputs/models", "mod_26"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )
} else {
  d10rn0 <- exp10_reads_filtered_for_nb(df.10.reads.no0)
  t10n0_lv <- sort(unique(stats::na.omit(d10rn0$tnum)))
  if (length(t10n0_lv) < 2L) {
    uel_n0 <- if ("exposure_label" %in% names(df.10.reads.no0)) {
      paste(utils::head(unique(df.10.reads.no0$exposure_label), 20L), collapse = " | ")
    } else {
      "(no exposure_label column)"
    }
    stop(
      "Experiment 10 reads (no time 0): need ≥ 2 time levels after filtering. Found: ",
      paste(t10n0_lv, collapse = ", "),
      " (nrow = ", nrow(d10rn0), "). Unique exposure_label (raw, up to 20): ", uel_n0, "."
    )
  }
  df_exp10_reads_no0_mod <- d10rn0 %>%
    dplyr::mutate(time = ordered(.data$tnum, levels = t10n0_lv)) %>%
    dplyr::select(-tnum)

  df_exp10_reads_no0_fac <- d10rn0 %>%
    dplyr::mutate(time = factor(as.character(.data$tnum), levels = as.character(t10n0_lv))) %>%
    dplyr::select(-tnum)

  mod_formula <- bf(gvb_sequence_number ~ mo(time) + offset(log(total_reads)))

  mod_23 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_mod,
    file = here::here("outputs/models", "mod_23"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_24 <- mod_23

  mod_formula <- bf(gvb_sequence_number ~ time + offset(log(total_reads)))

  mod_25 <- brm(
    mod_formula, family = negbinomial(),
    data = df_exp10_reads_no0_fac,
    file = here::here("outputs/models", "mod_25"),
    threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
    control = list(step_size = 0.01, adapt_delta = 0.99),
    backend = "cmdstanr", silent = 0, refresh = 1000,
    chains = 4, cores = 4, warmup = 2000, iter = 4000
  )

  mod_26 <- mod_25
}
```

```{r exp10-reads-no0-waic-compute, include=FALSE}
waic_tbl_one_exp10_reads_no0 <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

if (exp10_reads_has_quantity) {
  exp10_reads_no0_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads_no0(mod_23, "mo(time) + mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads_no0(mod_24, "mo(time) * mo(quantity_gbv)"),
    waic_tbl_one_exp10_reads_no0(mod_25, "time + quantity_gbv (factor)"),
    waic_tbl_one_exp10_reads_no0(mod_26, "time * quantity_gbv (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
} else {
  exp10_reads_no0_waic_tbl <- dplyr::bind_rows(
    waic_tbl_one_exp10_reads_no0(mod_23, "mo(time)"),
    waic_tbl_one_exp10_reads_no0(mod_25, "time (factor)")
  ) %>%
    dplyr::arrange(WAIC) %>%
    dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
    dplyr::mutate(
      WAIC = round(WAIC, 2),
      SE = round(SE, 2),
      delta_WAIC = round(delta_WAIC, 2)
    ) %>%
    dplyr::select(Model, WAIC, SE, delta_WAIC)
}
```

#### Effect plots

```{r exp10-reads-no0-eff, warning=FALSE, echo=FALSE, fig.width=8, fig.height=4.5, out.width="100%", dpi=600}
raw_no0 <- df_exp10_reads_no0_mod
n_t_exp10_n0 <- length(t10n0_lv)
# Include 0 h on the x-axis (no estimate; model excludes time 0).
n_t_exp10_n0_plot <- n_t_exp10_n0 + 1L
xlim_exp10_n0 <- exp10_reads_xlim_for_n(n_t_exp10_n0_plot)
x_lab_exp10_n0 <- c("0", as.character(t10n0_lv))

if (exp10_reads_has_quantity) {
  ce_both_no0 <- conditional_effects(mod_24, effects = "time:quantity_gbv")[[1]]
  qty_lv_no0 <- sort(unique(as.numeric(as.character(ce_both_no0$quantity_gbv))))
  ce_both_no0 <- ce_both_no0 %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = as.character(qty_lv_no0))
    ) %>%
    dplyr::mutate(
      tord = suppressWarnings(as.numeric(as.character(.data$time))),
      xtime = as.numeric(.data$time) + 1
    ) %>%
    dplyr::arrange(.data$quantity_gbv, .data$tord)

  raw_col_no0 <- raw_no0 %>%
    dplyr::mutate(
      quantity_gbv = factor(as.character(.data$quantity_gbv), levels = levels(ce_both_no0$quantity_gbv)),
      xtime = as.numeric(.data$time) + 1 + exp10_reads_fig_x_nudge_raw
    )

  exp10_reads_fig_y_max_no0 <- max(
    0,
    suppressWarnings(max(as.numeric(ce_both_no0$upper__), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(ce_both_no0$estimate__), na.rm = TRUE)),
    suppressWarnings(max(raw_col_no0$gvb_sequence_number, na.rm = TRUE)),
    na.rm = TRUE
  )
  exp10_reads_fig_y_upper_no0 <- max(
    exp10_reads_fig_y_upper,
    exp10_reads_fig_y_max_no0 * 1.72 + max(80, 0.12 * exp10_reads_fig_y_max_no0) + max(40, 0.04 * exp10_reads_fig_y_max_no0),
    10
  )
  exp10_reads_y_breaks_use_no0 <- exp10_reads_y_breaks[exp10_reads_y_breaks <= exp10_reads_fig_y_upper_no0]
  exp10_reads_fig_legend_npc_no0 <- c(
    x = (n_t_exp10_n0_plot - xlim_exp10_n0[1]) / diff(xlim_exp10_n0),
    y = log1p(10) / log1p(exp10_reads_fig_y_upper_no0)
  )

  n_qty_no0 <- dplyr::n_distinct(ce_both_no0$quantity_gbv)
  pal_no0 <- exp10_reads_qty_palette(n_qty_no0)
  ce_pd_no0 <- ggplot2::position_dodge(width = 0.85)

  exp10_reads_plot_theme_no0 <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35),
      legend.position = unname(exp10_reads_fig_legend_npc_no0),
      legend.justification = c(0.5, 0.5),
      legend.title = ggplot2::element_text(size = 10),
      legend.text = ggplot2::element_text(size = 9),
      legend.margin = ggplot2::margin(2, 4, 2, 4),
      legend.background = ggplot2::element_rect(fill = scales::alpha("white", 0.92), colour = "grey75")
    )

  p_exp10_reads_no0_eff <- ggplot2::ggplot(
    ce_both_no0,
    ggplot2::aes(
      x = .data$xtime,
      y = .data$estimate__,
      ymin = .data$lower__,
      ymax = .data$upper__,
      color = .data$quantity_gbv
    )
  ) +
    ggplot2::geom_point(
      data = raw_col_no0,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number, color = .data$quantity_gbv),
      inherit.aes = FALSE,
      position = ggplot2::position_jitterdodge(jitter.width = 0.12, jitter.height = 0.175, dodge.width = 0.85),
      alpha = 0.4,
      size = 0.85,
      stroke = 0
    ) +
    ggplot2::geom_errorbar(position = ce_pd_no0, width = 0.12, linewidth = 0.9) +
    ggplot2::geom_point(position = ce_pd_no0, size = 2.5, alpha = 0.9) +
    ggplot2::scale_color_manual(values = pal_no0) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10_n0_plot),
      labels = x_lab_exp10_n0,
      limits = xlim_exp10_n0,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(0, exp10_reads_fig_y_upper_no0),
      breaks = exp10_reads_y_breaks_use_no0,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme_no0 +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number") +
    ggplot2::guides(color = ggplot2::guide_legend(title = "GVB quantity", reverse = TRUE))

  fn_exp10_no0_plot <- here::here("outputs/plots", "figure-07-exp10-reads-no0-time-by-quantity.png")
} else {
  ce_both_no0 <- conditional_effects(mod_23, effects = "time")[[1]] %>%
    dplyr::mutate(xtime = as.numeric(.data$time) + 1)

  raw_col_no0 <- raw_no0 %>%
    dplyr::mutate(xtime = as.numeric(.data$time) + 1 + exp10_reads_fig_x_nudge_raw)

  exp10_reads_fig_y_max_no0 <- max(
    0,
    suppressWarnings(max(as.numeric(ce_both_no0$upper__), na.rm = TRUE)),
    suppressWarnings(max(as.numeric(ce_both_no0$estimate__), na.rm = TRUE)),
    suppressWarnings(max(raw_col_no0$gvb_sequence_number, na.rm = TRUE)),
    na.rm = TRUE
  )
  exp10_reads_fig_y_upper_no0 <- max(
    exp10_reads_fig_y_upper,
    exp10_reads_fig_y_max_no0 * 1.72 + max(80, 0.12 * exp10_reads_fig_y_max_no0) + max(40, 0.04 * exp10_reads_fig_y_max_no0),
    10
  )
  exp10_reads_y_breaks_use_no0 <- exp10_reads_y_breaks[exp10_reads_y_breaks <= exp10_reads_fig_y_upper_no0]

  col_mu_n0 <- "#0072B2"
  exp10_reads_plot_theme_no0 <- ggplot2::theme_bw() +
    ggplot2::theme(
      panel.grid.major.y = ggplot2::element_line(colour = "grey75", linewidth = 0.45),
      panel.grid.minor.y = ggplot2::element_blank(),
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.35)
    )

  p_exp10_reads_no0_eff <- ggplot2::ggplot(
    ce_both_no0,
    ggplot2::aes(x = .data$xtime, y = .data$estimate__, ymin = .data$lower__, ymax = .data$upper__)
  ) +
    ggplot2::geom_point(
      data = raw_col_no0,
      ggplot2::aes(x = .data$xtime, y = .data$gvb_sequence_number),
      inherit.aes = FALSE,
      position = ggplot2::position_jitter(width = 0.1, height = 0.175),
      alpha = 0.4,
      size = 0.85,
      color = "grey40",
      stroke = 0
    ) +
    ggplot2::geom_errorbar(width = 0.12, linewidth = 0.9, color = col_mu_n0) +
    ggplot2::geom_point(size = 2.5, alpha = 0.9, color = col_mu_n0) +
    ggplot2::scale_x_continuous(
      breaks = seq_len(n_t_exp10_n0_plot),
      labels = x_lab_exp10_n0,
      limits = xlim_exp10_n0,
      expand = c(0, 0)
    ) +
    ggplot2::scale_y_continuous(
      trans = scales::log1p_trans(),
      limits = c(0, exp10_reads_fig_y_upper_no0),
      breaks = exp10_reads_y_breaks_use_no0,
      labels = exp10_reads_y_labels,
      minor_breaks = NULL,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    exp10_reads_plot_theme_no0 +
    ggplot2::labs(x = "GVB exposure (hrs)", y = "GVB sequence number")

  fn_exp10_no0_plot <- here::here("outputs/plots", "figure-07-exp10-reads-no0-time-only.png")
}

ggplot2::ggsave(
  filename = fn_exp10_no0_plot,
  plot = p_exp10_reads_no0_eff,
  width = knitr::opts_current$get("fig.width"),
  height = knitr::opts_current$get("fig.height"),
  dpi = 600
)

p_exp10_reads_no0_eff
```

```{r exp10-reads-fig7-caption, echo=FALSE, results="asis"}
if (exp10_reads_has_quantity) {
  cat("Figure 7: GVB sequence counts with 0 h exposure excluded; same layout as Figure 6. Conditional effect of GVB exposure on mean sequence count by GVB quantity. Vertical axis is log(1 + count). Jittered points are raw counts; error bars are 95% credible intervals and coloured points are posterior means.\n\n")
} else {
  cat("Figure 7: GVB sequence counts with 0 h exposure excluded; same layout as Figure 6. Conditional effect of GVB exposure on mean sequence count. Vertical axis is log(1 + count). Grey jittered points are raw counts; blue error bars and points are 95% credible intervals and posterior means.\n\n")
}
```

#### Tables

```{r exp10-reads-no0-estimates, echo=FALSE}
if (exp10_reads_has_quantity) {
  ce_exp10_no0_t <- conditional_effects(mod_24, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  ce_exp10_no0_q <- conditional_effects(mod_24, effects = "quantity_gbv")[[1]] %>%
    dplyr::transmute(
      `Quantity GVB` = quantity_gbv,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_no0_t,
    caption = "Experiment 10 reads (no time 0) — primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of time (mean count).",
    format = "html"
  )

  knitr::kable(
    ce_exp10_no0_q,
    caption = "Experiment 10 reads (no time 0) — primary model `mo(time) * mo(quantity_gbv)`: marginal conditional effect of quantity GVB (mean count).",
    format = "html"
  )
} else {
  ce_exp10_no0_t <- conditional_effects(mod_23, effects = "time")[[1]] %>%
    dplyr::transmute(
      `Time (hrs)` = time,
      `Estimate` = round(estimate__, 3),
      `Lower` = round(lower__, 3),
      `Upper` = round(upper__, 3)
    )

  knitr::kable(
    ce_exp10_no0_t,
    caption = "Experiment 10 reads (no time 0) — model `mo(time)` with offset: conditional effect of time (mean count).",
    format = "html"
  )
}
```

#### Models

```{r exp10-reads-no0-models-show, ref.label="exp10-reads-no0-models-fit", echo=TRUE, eval=FALSE}

```

#### Model comparison

```{r exp10-reads-no0-waic-tab, echo=FALSE}
knitr::kable(
  exp10_reads_no0_waic_tbl,
  caption = "Experiment 10 reads (no time 0): WAIC (lower is better; all models include offset log(total reads)).",
  format = "html"
)
```

#### Validation

```{r exp10-reads-no0-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
if (exp10_reads_has_quantity) {
  pp_check(mod_23, ndraws = 200)
  pp_check(mod_24, ndraws = 200)
  pp_check(mod_25, ndraws = 200)
  pp_check(mod_26, ndraws = 200)
} else {
  pp_check(mod_23, ndraws = 200)
  pp_check(mod_25, ndraws = 200)
}
```

#### Output

::: {.panel-tabset}

##### mod_23

```{r exp10-reads-no0-summary-base, echo=FALSE}
#| class-output: model-output-panel
summary(mod_23)
```

##### mod_24

```{r exp10-reads-no0-summary-mo-int, echo=FALSE}
#| class-output: model-output-panel
if (exp10_reads_has_quantity) summary(mod_24)
```

##### mod_25

```{r exp10-reads-no0-summary-fac-add, echo=FALSE}
#| class-output: model-output-panel
summary(mod_25)
```

##### mod_26

```{r exp10-reads-no0-summary-fac-int, echo=FALSE}
#| class-output: model-output-panel
if (exp10_reads_has_quantity) summary(mod_26)
```

:::

:::

## QPCR experiment

<details><summary>Data frames</summary>
```{r}
#read in data ----
df <- read_excel(here::here("analysis/data/qPCR_results251017_forTom.xlsx"), sheet = 1) # Import .xlsx file

df <- df %>% clean_names() %>% select(-location_in_orchard, -note)

cols <- c("sample", "method", "method_name")
df[, cols] <- lapply(df[, cols], as.factor) # Coerce variables to factors

cols <- c("col", "row", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml", "cq1", "cq2", "cq3")
df[, cols] <- lapply(df[, cols], as.numeric) # Coerce variables to factors

df <- reshape2::melt(
  df,
  id = c("sample", "col", "row", "method", "method_name", "week", "pvh_number", "increase_in_pvh_number", "volume_in_container_ml")
) # Make it long

df <- df |>
  rename(replicate = variable, cq = value) |>
  filter(!is.na(row)) |>
  mutate(week_ord = ordered(week)) |> # ordered factor
  mutate(method = as.factor(make_clean_names(method, allow_dupes = TRUE))) |>
  mutate(method_name = as.factor(make_clean_names(method_name, allow_dupes = TRUE))) |>
  mutate(cq = if_else(is.nan(cq), 0, cq)) |>
  mutate(pa = ifelse(cq > 0, 1, cq)) |> # Create binary variable
  filter(method %in% c("b", "c", "d")) |>
  mutate(sub_method = paste0(method, sep = "_", method_name)) |>
  droplevels()

brief_dataframe(df)
```
</details>

::: {.panel-tabset}

```{r qpcr-models-fit-submethod, include=FALSE}
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, id = "sub_method") + (1|replicate))

mod_27 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_27"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: linear (separate slopes per sub-method)
mod_formula <- bf(pa ~ sub_method * increase_in_pvh_number + (1|replicate))

mod_28 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_28"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: one shared smooth; sub-method shifts intercept only
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number) + (1|replicate))

mod_29 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_29"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × PVH: separate smooths per sub-method (mgcv-style by)
mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, by = sub_method) + (1|replicate))

mod_30 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_30"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# + monotonic week within sub_method (parallel to mod_27–mod_30)
mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, id = "sub_method") + (1|replicate)
)

mod_31 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_31"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method * increase_in_pvh_number + mo(week_ord, id = "sub_method") + (1|replicate)
)

mod_32 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_32"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number) + (1|replicate)
)

mod_33 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_33"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

mod_formula <- bf(
  pa ~ sub_method + mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, by = sub_method) + (1|replicate)
)

mod_34 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_34"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × monotonic week + separate PVH smooths per sub-method
mod_formula <- bf(
  pa ~ sub_method * mo(week_ord, id = "sub_method") + s(increase_in_pvh_number, by = sub_method) + (1|replicate)
)

mod_35 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df,
  file = here::here("outputs/models", "mod_35"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)

# sub_method × monotonic PVH × monotonic week (global `mo` for week and PVH index)
df_mod36 <- df |>
  dplyr::mutate(increase_in_pvh_number = ordered(increase_in_pvh_number))

mod_formula <- bf(
  pa ~ sub_method * mo(increase_in_pvh_number) * mo(week_ord) + (1|replicate)
)

mod_36 <- brm(
  mod_formula,
  family = bernoulli(link = "logit"),
  data = df_mod36,
  file = here::here("outputs/models", "mod_36"),
  threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
  backend = "cmdstanr", silent = 0, refresh = 100,
  chains = 4, cores = 4, warmup = 1000, iter = 2000
)
```

```{r qpcr-eff-helpers, include=FALSE}
qpcr_sub_method_label <- function(x) {
  lbl <- vapply(as.character(x), function(s) {
    p <- strsplit(s, "_", fixed = TRUE)[[1]]
    if (length(p) == 0L) {
      return(s)
    }
    letter_lab <- dplyr::case_match(
      p[[1]],
      "b" ~ "B",
      "c" ~ "C",
      "d" ~ "D",
      .default = stringr::str_to_title(p[[1]])
    )
    if (length(p) == 1L) {
      return(paste0("Method ", letter_lab))
    }
    rest <- stringr::str_replace_all(paste(p[-1L], collapse = "_"), "_", " ")
    rest <- stringr::str_to_title(rest)
    paste0("Method ", letter_lab, " — ", rest)
  }, character(1))
  factor(lbl, levels = sort(unique(lbl)))
}

# Shared qPCR figure: PVH × sub_method conditional effects (ghost + focal facets)
qpcr_plot_pvh_by_submethod <- function(fit) {
  c_eff <- conditional_effects(fit, effects = "increase_in_pvh_number:sub_method")[[1]] %>%
    dplyr::arrange(sub_method, increase_in_pvh_number) %>%
    dplyr::mutate(sub_method_l = qpcr_sub_method_label(sub_method))
  panel_levels <- levels(c_eff$sub_method_l)
  ghost <- purrr::map_dfr(panel_levels, function(f) {
    c_eff %>%
      dplyr::filter(as.character(.data$sub_method_l) != as.character(f)) %>%
      dplyr::mutate(facet_panel = factor(f, levels = panel_levels))
  })
  focal <- c_eff %>%
    dplyr::mutate(facet_panel = factor(.data$sub_method_l, levels = panel_levels))
  obs_raw <- fit$data %>%
    dplyr::mutate(
      sub_method_l = qpcr_sub_method_label(.data$sub_method),
      facet_panel = factor(.data$sub_method_l, levels = panel_levels),
      pa_plot = binary_y_for_plot(.data$pa)
    )
  ggplot2::ggplot() +
    ggplot2::geom_jitter(
      data = obs_raw,
      ggplot2::aes(x = increase_in_pvh_number, y = pa_plot),
      inherit.aes = FALSE,
      height = 0.035,
      width = 0.15,
      alpha = 0.35,
      size = 0.85,
      color = "black",
      stroke = 0
    ) +
    ggplot2::geom_ribbon(
      data = ghost,
      ggplot2::aes(
        x = increase_in_pvh_number, ymin = lower__, ymax = upper__,
        group = sub_method_l
      ),
      inherit.aes = FALSE,
      fill = "gray45",
      alpha = 0.04
    ) +
    ggplot2::geom_line(
      data = ghost,
      ggplot2::aes(x = increase_in_pvh_number, y = estimate__, group = sub_method_l),
      inherit.aes = FALSE,
      color = "gray40",
      linewidth = 0.5,
      alpha = 0.1
    ) +
    ggplot2::geom_ribbon(
      data = focal,
      ggplot2::aes(
        x = increase_in_pvh_number, ymin = lower__, ymax = upper__,
        fill = sub_method_l, group = sub_method_l
      ),
      inherit.aes = FALSE,
      alpha = 0.2,
      color = NA
    ) +
    ggplot2::geom_line(
      data = focal,
      ggplot2::aes(
        x = increase_in_pvh_number, y = estimate__,
        color = sub_method_l, group = sub_method_l
      ),
      inherit.aes = FALSE,
      linewidth = 1
    ) +
    ggplot2::facet_wrap(ggplot2::vars(facet_panel), ncol = 3, scales = "fixed") +
    ggplot2::theme_classic() +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 9),
      legend.position = "none"
    ) +
    ggplot2::scale_y_continuous(
      labels = scales::percent,
      limits = c(-0.13, 1.13),
      breaks = prob_det_y_breaks,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    ggplot2::scale_color_brewer(palette = "Dark2", type = "qual") +
    ggplot2::scale_fill_brewer(palette = "Dark2", type = "qual") +
    ggplot2::labs(x = "PVH number", y = "Probability of detection (%)")
}

# Week × sub_method: posterior mean response at each (week_ord, sub_method), holding
# increase_in_pvh_number at its within–sub-method median and replicate at a reference level.
# Uses posterior_epred() so it behaves for all PVH structures (linear, shared spline, `by`)
# — conditional_effects("week_ord:sub_method") is unreliable for e.g. mod_34 / mod_35 / mod_36.
qpcr_plot_week_by_submethod <- function(fit) {
  d <- fit$data

  wk <- d$week_ord
  if (is.factor(wk)) {
    wk_lvls <- levels(wk)
    wk_ordered <- is.ordered(wk)
  } else {
    wk_lvls <- unique(as.character(wk[!is.na(wk)]))
    wk_ordered <- TRUE
  }

  sm <- d$sub_method
  if (!is.factor(sm)) {
    sm <- factor(sm, levels = unique(as.character(sm[!is.na(sm)])))
  } else {
    sm <- droplevels(sm)
  }
  sm_lvls <- levels(sm)

  rp <- d$replicate
  if (!is.factor(rp)) {
    rp <- factor(rp, levels = unique(as.character(rp[!is.na(rp)])))
  } else {
    rp <- droplevels(rp)
  }
  rep_lvls <- levels(rp)
  rep_ref <- rep_lvls[[1L]]

  if (length(wk_lvls) < 1L || length(sm_lvls) < 1L) {
    stop("fit$data is missing week_ord or sub_method factor levels.")
  }

  pvh <- d$increase_in_pvh_number
  sm_chr <- as.character(sm)
  if (is.numeric(pvh)) {
    pvh_med_src <- as.numeric(pvh)
  } else {
    pvh_med_src <- suppressWarnings(as.numeric(as.character(pvh)))
    if (anyNA(pvh_med_src)) {
      pvh_med_src <- as.numeric(pvh)
    }
  }
  med_pvh_num <- tapply(pvh_med_src, sm_chr, stats::median, na.rm = TRUE)
  med_pvh_num <- med_pvh_num[sm_lvls]
  if (any(is.na(med_pvh_num))) {
    stop("Could not compute median PVH for every sub_method level.")
  }

  rows <- tidyr::expand_grid(
    week_ord = factor(wk_lvls, levels = wk_lvls, ordered = wk_ordered),
    sub_method = factor(sm_lvls, levels = sm_lvls)
  )
  if (is.numeric(pvh)) {
    rows$increase_in_pvh_number <- as.numeric(med_pvh_num[as.character(rows$sub_method)])
  } else {
    pvh_lv <- levels(pvh)
    pvh_lv_num <- suppressWarnings(as.numeric(pvh_lv))
    if (anyNA(pvh_lv_num)) {
      pvh_lv_num <- seq_along(pvh_lv)
    }
    snap_level <- function(m) {
      pvh_lv[which.min(abs(pvh_lv_num - m))]
    }
    rows$increase_in_pvh_number <- factor(
      vapply(as.character(rows$sub_method), function(s) snap_level(med_pvh_num[[s]]), character(1)),
      levels = pvh_lv,
      ordered = is.ordered(pvh)
    )
  }
  rows$replicate <- factor(rep_ref, levels = rep_lvls)

  nd <- tryCatch(
    min(1000L, as.integer(brms::nsamples(fit))),
    error = function(...) 1000L
  )
  if (!is.finite(nd) || nd < 1L) nd <- 1000L
  ep <- brms::posterior_epred(
    object = fit,
    newdata = rows,
    re_formula = NA,
    ndraws = nd
  )
  rq <- apply(ep, 2L, stats::quantile, probs = c(0.025, 0.5, 0.975), na.rm = TRUE)
  c_eff <- rows %>%
    dplyr::mutate(
      lower__ = as.numeric(rq[1, ]),
      estimate__ = as.numeric(rq[2, ]),
      upper__ = as.numeric(rq[3, ])
    ) %>%
    dplyr::filter(is.finite(estimate__)) %>%
    dplyr::mutate(sub_method_l = qpcr_sub_method_label(sub_method))
  if (nrow(c_eff) < 1L) {
    stop("posterior_epred returned no usable week × sub-method estimates.")
  }

  panel_levels <- levels(c_eff$sub_method_l)
  plot_dat <- c_eff %>%
    dplyr::mutate(facet_panel = factor(.data$sub_method_l, levels = panel_levels))
  obs_raw <- d %>%
    dplyr::mutate(
      week_ord = factor(as.character(.data$week_ord), levels = wk_lvls, ordered = wk_ordered),
      sub_method_l = qpcr_sub_method_label(.data$sub_method),
      facet_panel = factor(.data$sub_method_l, levels = panel_levels),
      pa_plot = binary_y_for_plot(.data$pa)
    )

  n_col <- length(unique(c_eff$sub_method_l))
  p <- ggplot2::ggplot(plot_dat) +
    ggplot2::geom_jitter(
      data = obs_raw,
      ggplot2::aes(x = week_ord, y = pa_plot),
      inherit.aes = FALSE,
      height = 0.035,
      width = 0.18,
      alpha = 0.35,
      size = 0.85,
      color = "black",
      stroke = 0
    ) +
    ggplot2::geom_linerange(
      ggplot2::aes(
        x = week_ord,
        ymin = lower__, ymax = upper__,
        color = sub_method_l,
        group = week_ord
      ),
      linewidth = 1
    ) +
    ggplot2::geom_point(
      ggplot2::aes(
        x = week_ord, y = estimate__,
        color = sub_method_l,
        group = week_ord
      ),
      size = 3,
      alpha = 0.55
    ) +
    ggplot2::facet_wrap(ggplot2::vars(facet_panel), ncol = 3, scales = "fixed") +
    ggplot2::theme_classic() +
    ggplot2::theme(
      strip.text = ggplot2::element_text(size = 9),
      legend.position = "none",
      panel.grid.major.x = ggplot2::element_line(colour = "grey85", linewidth = 0.3),
      panel.grid.major.y = ggplot2::element_line(colour = "grey85", linewidth = 0.3)
    ) +
    ggplot2::scale_y_continuous(
      labels = scales::percent,
      limits = c(-0.13, 1.13),
      breaks = prob_det_y_breaks,
      expand = ggplot2::expansion(mult = c(0, 0))
    ) +
    ggplot2::labs(x = "Week", y = "Probability of detection (%)")
  if (n_col <= 8L) {
    p <- p + ggplot2::scale_color_brewer(palette = "Dark2", type = "qual")
  } else {
    p <- p + ggplot2::scale_color_hue(l = 55, c = 80)
  }
  p
}
```

```{r qpcr-waic-compute, include=FALSE}
waic_tbl_one <- function(m, model_name) {
  w <- brms::waic(m)
  est <- w$estimates
  rn <- rownames(est)
  waic_row <- if ("waic" %in% rn) "waic" else if ("WAIC" %in% rn) "WAIC" else rn[1]
  tibble::tibble(
    Model = model_name,
    WAIC = as.numeric(est[waic_row, "Estimate"]),
    SE = as.numeric(est[waic_row, "SE"])
  )
}

waic_tbl_pre <- dplyr::bind_rows(
  waic_tbl_one(mod_27, "mod_27"),
  waic_tbl_one(mod_28, "mod_28"),
  waic_tbl_one(mod_29, "mod_29"),
  waic_tbl_one(mod_30, "mod_30"),
  waic_tbl_one(mod_31, "mod_31"),
  waic_tbl_one(mod_32, "mod_32"),
  waic_tbl_one(mod_33, "mod_33"),
  waic_tbl_one(mod_34, "mod_34"),
  waic_tbl_one(mod_35, "mod_35"),
  waic_tbl_one(mod_36, "mod_36")
) %>%
  dplyr::mutate(model_id = as.integer(sub("^mod_", "", Model))) %>%
  dplyr::arrange(WAIC, model_id)

waic_tbl <- waic_tbl_pre %>%
  dplyr::mutate(delta_WAIC = WAIC - min(WAIC, na.rm = TRUE)) %>%
  dplyr::mutate(
    WAIC = round(WAIC, 2),
    SE = round(SE, 2),
    delta_WAIC = round(delta_WAIC, 2)
  ) %>%
  dplyr::select(Model, WAIC, SE, delta_WAIC)

qpcr_model_formula_lines <- c(
  "27" = "bf(pa ~ sub_method + s(increase_in_pvh_number, id = \"sub_method\") + (1|replicate))",
  "28" = "bf(pa ~ sub_method * increase_in_pvh_number + (1|replicate))",
  "29" = "bf(pa ~ sub_method + s(increase_in_pvh_number) + (1|replicate))",
  "30" = "bf(pa ~ sub_method + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "31" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, id = \"sub_method\") + (1|replicate))",
  "32" = "bf(pa ~ sub_method * increase_in_pvh_number + mo(week_ord, id = \"sub_method\") + (1|replicate))",
  "33" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number) + (1|replicate))",
  "34" = "bf(pa ~ sub_method + mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "35" = "bf(pa ~ sub_method * mo(week_ord, id = \"sub_method\") + s(increase_in_pvh_number, by = sub_method) + (1|replicate))",
  "36" = "bf(pa ~ sub_method * mo(increase_in_pvh_number) * mo(week_ord) + (1|replicate))"
)
```

## Effect plots

**Figure 8** (week × sub-method) and **Figure 9** (PVH number × sub-method) for the qPCR experiment; each tab repeats the same pair of figures for a different fitted model (`mod_32`, `mod_34`, `mod_36`).

::: {.panel-tabset}

## Effect plots (mod_32)

```{r qpcr-effect-mod32, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_32
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["32"]], "\n```\n\n", sep = "")
cat("**Figure 8 — week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-08-qpcr-week-mod_32.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n\nFigure 8: qPCR experiment — predicted outcome by experimental week and qPCR sub-method.\n\n")
cat("**Figure 9 — PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-09-qpcr-pvh-mod_32.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
cat("\n\nFigure 9: qPCR experiment — predicted outcome by PVH number and qPCR sub-method.\n\n")
```

## Effect plots (mod_34)

```{r qpcr-effect-mod34, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_34
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["34"]], "\n```\n\n", sep = "")
cat("**Figure 8 — week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-08-qpcr-week-mod_34.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n\nFigure 8: qPCR experiment — predicted outcome by experimental week and qPCR sub-method.\n\n")
cat("**Figure 9 — PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-09-qpcr-pvh-mod_34.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
cat("\n\nFigure 9: qPCR experiment — predicted outcome by PVH number and qPCR sub-method.\n\n")
```

## Effect plots (mod_36)

```{r qpcr-effect-mod36, echo=FALSE, results="asis", fig.width=9, fig.height=6, out.width="100%", dpi=600}
#| fig-keep: all
fit <- mod_36
gw <- knitr::opts_current$get("fig.width")
gh <- knitr::opts_current$get("fig.height")
cat("```r\n", qpcr_model_formula_lines[["36"]], "\n```\n\n", sep = "")
cat("**Figure 8 — week × sub-method**\n\n")
pw <- qpcr_plot_week_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-08-qpcr-week-mod_36.png"),
  plot = pw,
  width = gw,
  height = gh,
  dpi = 600
)
print(pw)
cat("\n\nFigure 8: qPCR experiment — predicted outcome by experimental week and qPCR sub-method.\n\n")
cat("**Figure 9 — PVH number × sub-method**\n\n")
pp <- qpcr_plot_pvh_by_submethod(fit)
ggplot2::ggsave(
  filename = here::here("outputs/plots", "figure-09-qpcr-pvh-mod_36.png"),
  plot = pp,
  width = gw,
  height = gh,
  dpi = 600
)
print(pp)
cat("\n\nFigure 9: qPCR experiment — predicted outcome by PVH number and qPCR sub-method.\n\n")
```

:::

## Models

Sub-method formulations (`mod_27`–`mod_36`; includes monotonic week in `mod_31`–`mod_36`):

```{r qpcr-models-show, ref.label="qpcr-models-fit-submethod", echo=TRUE, eval=FALSE}

```

## Model Comparison

```{r qpcr-waic, echo=FALSE}
knitr::kable(
  waic_tbl,
  caption = "Model comparison (qPCR Experiment): WAIC (lower is better).",
  format = "html"
)
```

## Validation

```{r qpcr-pp-check}
#| echo: false
#| message: false
#| fig-show: hold
#| layout-ncol: 2
pp_check(mod_27, ndraws = 200, col = 2)
pp_check(mod_28, ndraws = 200, col = 2)
pp_check(mod_29, ndraws = 200, col = 2)
pp_check(mod_30, ndraws = 200, col = 2)
pp_check(mod_31, ndraws = 200, col = 2)
pp_check(mod_32, ndraws = 200, col = 2)
pp_check(mod_33, ndraws = 200, col = 2)
pp_check(mod_34, ndraws = 200, col = 2)
pp_check(mod_35, ndraws = 200, col = 2)
pp_check(mod_36, ndraws = 200, col = 2)
```

## Output

::: {.panel-tabset}

### mod_27

```{r qpcr-summary-mod3, echo=FALSE}
#| class-output: model-output-panel
summary(mod_27)
```

### mod_28

```{r qpcr-summary-mod4, echo=FALSE}
#| class-output: model-output-panel
summary(mod_28)
```

### mod_29

```{r qpcr-summary-mod5, echo=FALSE}
#| class-output: model-output-panel
summary(mod_29)
```

### mod_30

```{r qpcr-summary-mod6, echo=FALSE}
#| class-output: model-output-panel
summary(mod_30)
```

### mod_31

```{r qpcr-summary-mod7, echo=FALSE}
#| class-output: model-output-panel
summary(mod_31)
```

### mod_32

```{r qpcr-summary-mod32, echo=FALSE}
#| class-output: model-output-panel
summary(mod_32)
```

### mod_33

```{r qpcr-summary-mod9, echo=FALSE}
#| class-output: model-output-panel
summary(mod_33)
```

### mod_34

```{r qpcr-summary-mod34, echo=FALSE}
#| class-output: model-output-panel
summary(mod_34)
```

### mod_35

```{r qpcr-summary-mod11, echo=FALSE}
#| class-output: model-output-panel
summary(mod_35)
```

### mod_36

```{r qpcr-summary-mod36, echo=FALSE}
#| class-output: model-output-panel
summary(mod_36)
```

:::

:::

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Session Information</summary>
```{r}
utils::sessionInfo()
```
</details>

