---
title: "Justine - GVB (2026)"
author:  "Dr Tom Moore - Statistical Scientist"
date:  '`r Sys.Date()`'
knit: (function(inputFile, encoding) { out_dir <- "Reports";
      rmarkdown::render(inputFile, encoding=encoding, output_file=file.path(dirname(inputFile),
      out_dir, 'Analysis_Report.html'))})
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: true
    toc_float: true
    toc_depth: 5
    theme: united
    highlight: tango
    fig_width: 9
    fig_height: 6
---
<details><summary>Setup</summary>
```{r}
#| label: setup
#| results: false

# Install packages and functions
if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak")
}

packages <- c(
  "tidyverse", "here", "janitor", "readxl", "modelr",
  "reshape2", "arrow", "pins", "usethis",
  "ggtext", "forcats", "cowplot", "ggdendro",
  "gplots", "GGally", "dagitty", "ggdag", "simDAG",
  "future", "furrr", "rstan", "bnlearn", "Rgraphviz",
  "brms", "bayesplot", "tidybayes", "cmdstanr",
  "nadiv", "graph4lg", "marginaleffects", "bayesplot", "rethinking"
)

# Install missing packages with pak
missing_pkgs <- packages[!sapply(packages, requireNamespace, quietly = TRUE)]
if (length(missing_pkgs) > 0) {
  pak::pkg_install(missing_pkgs, ask = FALSE)
  cat("Installed:", paste(missing_pkgs, collapse = ", "), "\n")
} else {
  cat("All Packages Installed :) \n")
}

# Load all packages
invisible(sapply(packages, library, character.only = TRUE))

bayesplot::color_scheme_set("blue")                                    # Set color scheme
options(mc.cores = parallel::detectCores())                            # Parallel Chains

# Production Sever Settings
cmdstanr::set_cmdstan_path("/powerplant/workspace/hrltxm/workbench-k8s/stan/vanilla_stan/cmdstan-2.36.0")
```
</details>

<details><summary>Wrangling & Summary</summary>
```{r}
#read in data ----
df <- read_excel(here("Analysis/Data/qPCR_results251017_forTom.xlsx"), sheet = 1) # Import .xlsx file

df <- df %>% clean_names() %>% select(-location_in_orchard, -note)

cols <- c('sample', 'method','method_name')
df[,cols] <- lapply(df[,cols],as.factor) # Coerce variables to factors 

cols <- c('col','row','week','pvh_number', 'increase_in_pvh_number', 'volume_in_container_ml', 'cq1', 'cq2', 'cq3')
df[,cols] <- lapply(df[,cols],as.numeric) # Coerce variables to factors 

df <- reshape2::melt(df, id=c('sample', 'col', 'row', 'method', 'method_name', 'week', 'pvh_number', 'increase_in_pvh_number', 'volume_in_container_ml'))  # Make it long

df <- df |>
  rename(replicate = variable,
         cq = value) |>
  filter(!is.na(row)) |>
  mutate(week_ord = ordered(week)) |> # ordered factor
  mutate(method = as.factor(make_clean_names(method, allow_dupes = TRUE))) |>
  mutate(method_name = as.factor(make_clean_names(method_name, allow_dupes = TRUE))) |>
  mutate(cq = if_else(is.nan(cq), 0, cq)) |>
  mutate(pa = ifelse(cq > 0, 1, cq)) |> # Create binary variable
  filter(method %in% c("b", "c","d")) |>
  mutate(sub_method = paste0(method, sep = "_", method_name)) |>
  droplevels()

df |> group_by(method, week_ord) |> tally()

str(df)
```

</details>


```{r}
df |> ggplot(
  aes(x = week_ord, y = pa , color = method)
) +
  geom_jitter()
```


# Methods:

To compare the probability of detection (%) of GVB at different exposure levels (0, 2, 6, 16, 24, 48 hrs), with and without a Cq cut off (i.e., if a Cq value was greater than 39 then classify as no detection), and GVB density (where data was available: at 2hr and 24hrs), generalized linear mixed effect (GLMM) models were performed in R (R Core Team 2022) using the brms package (Bürkner, 2017). 

For all models a Bernoulli distribution was selected with a logit link function to model the binary detection response (yes = 1, no = 0). The fixed effects of time and number of GBV were treated as monotonic predictors since they are ordered categories. An interaction term was included between time and number of GVB when this information was available. A random effect of sample was used for all models since measures were collected in triplicates. A weakly informative prior was used for the fixed effects to aid convergence (normal distribution with a mean of 0 and standard deviation of 1.5). 

Model validation was performed using prior and posterior predictive checks, as well as effective samples sizes and r-hats (McElreath, 2020).

## Experiment 6
### Analyses
<details><summary>Models</summary>
```{r}
mod_formula <- bf(pa ~ method + mo(week_ord, id = "method") + gp(row,col) + s(increase_in_pvh_number) + s(volume_in_container_ml) + (1|replicate))

priors <- prior(normal(0,1.5), class = b)

mod_1 <- brm(mod_formula,  
             family = bernoulli(link = "logit"),
             prior = priors,
             data = df, 
             file = paste0(here("Analysis/Output/Models/mod_1_2025.Rmd")),
             threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
             # control = list(step_size = 0.01, adapt_delta = 0.99),
             backend = "cmdstanr", silent = 0, refresh = 100,
             chains = 4, cores = 4, warmup = 1000, iter = 2000)



mod_formula <- bf(pa ~ method + s(increase_in_pvh_number, id = "method") + (1|replicate))

priors <- prior(normal(0,1.5), class = b)

mod_2 <- brm(mod_formula,  
             family = bernoulli(link = "logit"),
             prior = priors,
             data = df, 
             file = paste0(here("Analysis/Output/Models/mod_2_2026.Rmd")),
             threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
             # control = list(step_size = 0.01, adapt_delta = 0.99),
             backend = "cmdstanr", silent = 0, refresh = 100,
             chains = 4, cores = 4, warmup = 1000, iter = 2000)

mod_formula <- bf(pa ~ sub_method + s(increase_in_pvh_number, id = "sub_method") + (1|replicate))

priors <- prior(normal(0,1.5), class = b)

mod_3 <- brm(mod_formula,  
             family = bernoulli(link = "logit"),
             prior = priors,
             data = df, 
             file = paste0(here("Analysis/Output/Models/mod_3_2026.Rmd")),
             threads = threading(threads = NULL, grainsize = 1250, static = FALSE),
             # control = list(step_size = 0.01, adapt_delta = 0.99),
             backend = "cmdstanr", silent = 0, refresh = 100,
             chains = 4, cores = 4, warmup = 1000, iter = 2000)
```
</details><details><summary>Validation</summary>
```{r, echo=FALSE, message=FALSE, fig.show='hold',out.width='50%'}
pp_check(mod_1, ndraws = 200)
pp_check(mod_2, ndraws = 200)
pp_check(mod_3, ndraws =200)
```
</details><details><summary>Output</summary>
```{r,echo=FALSE}
summary(mod_1)
summary(mod_2)
summary(mod_3)
```
</details> 

### Effect Plots
## Model 1 (all effects)
```{r,warning=FALSE,echo=FALSE,fig.width=7,fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
# c_eff_1 <- conditional_effects(mod.1)[[1]] %>% mutate(cutoff = "No")
# c_eff_2 <- conditional_effects(mod.2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
# c_eff <- rbind(c_eff_1, c_eff_2)

# c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
#   geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
#     geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
#     theme_bw() +
#   scale_y_continuous(labels = percent) +
#   scale_color_manual(values = c("darkblue", "darkgreen")) +
#   labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
#   guides(color=guide_legend(title="Cq cut off (> 39)", reverse = TRUE))


conditional_effects(mod_1, effects = "week_ord:method")
conditional_effects(mod_1)
```

Figure 1: Conditional effect plot of GBV exposure on the probability of detection. Error bars represent 95% credible intervals with points representing the mean estimate.


## Model 2 (increase in pvh and method only)

```{r,warning=FALSE,echo=FALSE,fig.width=7,fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
# c_eff_1 <- conditional_effects(mod.1)[[1]] %>% mutate(cutoff = "No")
# c_eff_2 <- conditional_effects(mod.2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
# c_eff <- rbind(c_eff_1, c_eff_2)

# c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
#   geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
#     geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
#     theme_bw() +
#   scale_y_continuous(labels = percent) +
#   scale_color_manual(values = c("darkblue", "darkgreen")) +
#   labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
#   guides(color=guide_legend(title="Cq cut off (> 39)", reverse = TRUE))


conditional_effects(mod_2, effects = "increase_in_pvh_number:method")
conditional_effects(mod_2)
```

Figure 2: Conditional effect plot of GBV exposure on the probability of detection. Error bars represent 95% credible intervals with points representing the mean estimate.

## Model 3 (increase in pvh and sub method only)

```{r,warning=FALSE,echo=FALSE,fig.width=7,fig.height=4, out.width='100%', dpi = 600}
# Custom conditional effect plots
# c_eff_1 <- conditional_effects(mod.1)[[1]] %>% mutate(cutoff = "No")
# c_eff_2 <- conditional_effects(mod.2)[[1]] %>% mutate(cutoff = "Yes") %>% rename(positive = positive_cut)
# c_eff <- rbind(c_eff_1, c_eff_2)

# c_eff %>% ggplot(aes(x = time, y = estimate__, ymin = lower__, ymax = upper__, color = cutoff)) +
#   geom_errorbar(position = position_dodge(width = -0.5), width = 0.3, size = 1) +
#     geom_point(position = position_dodge(width = -0.5), size = 3, alpha = 0.5) +
#     theme_bw() +
#   scale_y_continuous(labels = percent) +
#   scale_color_manual(values = c("darkblue", "darkgreen")) +
#   labs(x = "GVB exposure (hrs)", y = "Probability of detection (%)") +
#   guides(color=guide_legend(title="Cq cut off (> 39)", reverse = TRUE))


conditional_effects(mod_3, effects = "increase_in_pvh_number:sub_method")
conditional_effects(mod_3)
```


## Results

In Bayesian analysis, explanatory variables are considered “important” (significant) if their 95% credible interval (95% CI) is exclusive of zero.

# References

Bürkner, P.C., (2017). brms: An R Package for Bayesian Multilevel Models Using Stan. Journal of Statistical Software, 80(1), 1-28. <https://econpapers.repec.org/article/jssjstsof/v_3a080_3ai01.htm>

McElreath, R. (2020). Statistical rethinking: A Bayesian course with examples in R and Stan. Chapman and Hall/CRC.

R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

<details><summary>Session Information</summary>
```{r}
utils::sessionInfo()
```
</details>
